### 2016-10-31 analysis/0_field_work_proposal.Rmd
add fieldwork proposal

---
title: "Field work"
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: fill
    orientation: rows
---

```{r setup, include=FALSE}
library('flexdashboard')
```

```{r pacakges}
library('rgdal')
library("leaflet") 
library("sp")
library("raster")

library('dplyr')

library('ggplot2')

```

```{r readData}
# mymachine <- '/Users/ajpeluLap/' 
mymachine <- '/Users/ajpelu/' 
di <- paste0(mymachine, 'Dropbox/phd/phd_repos/qpyr_dendro/')

# Data 
ot <- read.csv(paste0(di,'data_raw/geoinfo/other_projects.csv'), 
               header = TRUE, sep = ',')
mig <- readOGR(paste0(di, 'data_raw/geoinfo/'), layer = 'migrame_transectos', verbose = FALSE)

# Spatial info
ots <- SpatialPointsDataFrame(coords = ot[,c('x','y')], data = ot,
                               proj4string = CRS("+proj=utm +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))

mig <- spTransform(mig, CRS("+init=epsg:4236"))
ots <- spTransform(ots, CRS("+init=epsg:4236"))

# Select only transect of forest (migrame)
mig_forest <- mig[mig$tipo == "ROBLEDAL",]
mig_forest_ca <- mig_forest[mig_forest$localidad == "CANAR",]
mig_forest_sj <- mig_forest[mig_forest$localidad == "SANJUAN",]

ots_phe <- ots[ots$notes == 'Fenologia_Obsnev',]
ots_gea <- ots[ots$notes == 'GeaIzquierdo2014',]

```

```{r}
mymap <- leaflet() %>% 
  # BaseMaps 
  addWMSTiles('http://www.ideandalucia.es/wms/ortofoto2009?',
              layers = 'oca10_2008-09',
              options = WMSTileOptions(format = "image/png", transparent = FALSE),
              attribution = '<a href="http://www.juntadeandalucia.es/institutodeestadisticaycartografia" target="_blank">Instituto de Estadística y Cartografía de Andalucía</a>',
              group = 'Ortophotography 2009') %>%
   addWMSTiles('http://www.ideandalucia.es/services/toporaster10/wms?',
              layers = 'toporaster10',
              options = WMSTileOptions(format = "image/png", transparent = FALSE),
              attribution = '<a href="http://www.juntadeandalucia.es/institutodeestadisticaycartografia" target="_blank">Instituto de Estadística y Cartografía de Andalucía</a>',
              group = 'Topographical') %>%

  # Layers control
  addLayersControl(position = 'bottomright',
                   # baseGroups = c("Topographical","Ortophotography 2009","PNOA"),
                   baseGroups = c("Ortophotography 2009", "Topographical"),
                   overlayGroups = c("Migrame", "OBSNEV Fen", "Gea Izquierdo 2014"),
                   options = layersControlOptions(collapsed = FALSE))

```

San Juan
=======================================================================

```{r}
myext <- extent(mig_forest_sj)

mymap %>% 
  fitBounds(myext@xmin, myext@ymin, myext@xmax, myext@ymax) %>% 
  addPolylines(data=mig_forest_sj, color='blue', group='Migrame', popup=mig_forest_sj$nombre) %>% 
  addCircles(ots_phe, 
             lng=coordinates(ots_phe)[,'x'], 
             lat=coordinates(ots_phe)[,'y'], 
             group='OBSNEV Fen', radius=2, color = 'red', popup=ots_phe$name) %>% 
  addCircles(ots_gea, 
             lng=coordinates(ots_gea)[,'x'], 
             lat=coordinates(ots_gea)[,'y'], 
             group='Gea Izquierdo 2014', radius=2, color = 'green', popup=ots_gea$name)

```

Caniar
=======================================================================

```{r}
myext <- extent(mig_forest_ca)

mymap %>% 
  fitBounds(myext@xmin, myext@ymin, myext@xmax, myext@ymax) %>% 
  addPolylines(data=mig_forest_ca, color='blue', group='Migrame', popup=mig_forest_ca$nombre) %>% 
  addCircles(ots_phe, 
             lng=coordinates(ots_phe)[,'x'], 
             lat=coordinates(ots_phe)[,'y'], 
             group='OBSNEV Fen', radius=2, color = 'red', popup=ots_phe$name) %>% 
  addCircles(ots_gea, 
             lng=coordinates(ots_gea)[,'x'], 
             lat=coordinates(ots_gea)[,'y'], 
             group='Gea Izquierdo 2014', radius=2, color = 'green', popup=ots_gea$name)

```