### 2016-11-30 analysis/geo.Rmd
updating

---
title: "Get topographic info of gps points"
author: "AJ Perez-Luque (@ajpelu)"
date: "2016 Nov"
output:  
  md_document:
    variant: markdown_github
---

```{r packages, warning=FALSE, message=FALSE}}
library("rgdal")
library("sp")
library("raster")
library("dplyr")
```

# Read data 
```{r wd, echo=FALSE}
machine <- 'ajpelu'
# machine <- 'ajpeluLap'
di <- paste('/Users/', machine, '/Dropbox/phd/phd_repos/qpyr_dendro/', sep='')

# Connect with NAS and copy the mde file located at 
# '/cartografia/Informacion_Ambiental/BASES_DE_REFERENCIA_TERRITORIAL/MDE_SN_AreaInfluecia_10m/mde_sn_area_influencia.asc' 
#---------------------------------
```

# Extract data
```{r} 
ditemp <- paste('/Users/', machine, '/Documents/', sep='')
setwd(ditemp)

# read DEM 
demsn <- raster("mde_sn_area_influencia.asc")

setwd(di)

## Projection 
# Get projection
projection(demsn)

# Set projection
crs(demsn) <- "+proj=utm +zone=30 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

# Read points data

gps_point <- readOGR(paste0(di, 'data_raw/geoinfo/'), layer = 'dendro_garmin', verbose = FALSE)

# Read pixels data 
eviresi <- read.csv(file=paste(di, "/data/evi_resilience.csv", sep=""), header = TRUE, sep = ',')

## convert to spatial df 
eviresi_sp <- SpatialPointsDataFrame(coords = eviresi[,c('lng','lat')], data = eviresi,
                               proj4string = CRS("+init=epsg:4326"))
## reproject 
eviresi_sp <- spTransform(eviresi_sp, crs(demsn))

evi_elev <- extract(demsn, eviresi_sp, method='simple', sp=TRUE)

elev <- evi_elev@data %>% 
  dplyr::select(iv_malla_modi_id, mde_sn_area_influencia) %>%
  transmute(iv_malla_modi_id, elev = mde_sn_area_influencia)

# Export data
write.csv(elev, file=paste(di, "/data/elev.csv", sep=""), row.names = FALSE)
```
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.