### 2016-12-02 analysis/get_elevation_points.Rmd
get elevation points

---
title: "Get topographic info of gps points"
author: "AJ Perez-Luque (@ajpelu)"
date: "2016 Nov"
output:  
  md_document:
    variant: markdown_github
---

```{r packages, warning=FALSE, message=FALSE}}
library("rgdal")
library("sp")
library("raster")
library("dplyr")
```

# Read data 
```{r wd, echo=FALSE}
# machine <- 'ajpelu'
machine <- 'ajpeluLap'
di <- paste('/Users/', machine, '/Dropbox/phd/phd_repos/qpyr_dendro/', sep='')

# Connect with NAS and copy the mde file located at 
# '/cartografia/Informacion_Ambiental/BASES_DE_REFERENCIA_TERRITORIAL/MDE_SN_AreaInfluecia_10m/mde_sn_area_influencia.asc' 
#---------------------------------
```

# Extract data
```{r} 
ditemp <- paste('/Users/', machine, '/Documents/', sep='')
setwd(ditemp)

# read DEM 
demsn <- raster("mde_sn_area_influencia.asc")

setwd(di)

## Projection 
# Get projection
projection(demsn)

# Set projection
crs(demsn) <- "+proj=utm +zone=30 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

# Read points data 
gps_point <- readOGR(paste0(di, 'data_raw/geoinfo/'), layer = 'dendro_garmin', verbose = FALSE)

## reproject 
gps_point_re <- spTransform(gps_point, crs(demsn))

gps_elev <- extract(demsn, gps_point_re, method='simple', sp=TRUE)

elev <- gps_elev@data %>% 
  dplyr::select(lat, lon, elev, name, date, elev_mde = mde_sn_area_influencia)

# Export data
write.csv(elev, file=paste(di, "/data/elev.csv", sep=""), row.names = FALSE)
```