### 2016-12-07 man/notes_dendro.Rmd

---
title: "Dendrocronology notes"
author: "AJ Perez-Luque (@ajpelu)"
date: "2016 Nov"
output:  
  md_document:
    toc: true
    variant: markdown_github
bibliography: references.bib
csl: ecology.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Ring Width measurement

## Notes from [@GeaIzquierdo2012Tree]

* Two cores were taken at 1.30 m from each selected tree
* The stem diamenter and tree height were measured
* Cores were air dried, fixed in woodend mounts, sanded and RW measured 
* Ring-width series were visually cross-dated and the dating was checked using COFECHA `$TODO$` Check COFECHA
* Ring-width inidces were obtained by removing the age trend using 30-year individual splines and autocorrelation before computing the final indices using the ration between the observed and expected annual mean growth

# Features of vascular system of *Q. pyrenaica*

* Ring-porous tress are assumed to be particularly sensitive to climate warming, because they operate with narrow hydraulic safety margins (see Choat et al. )[@PerezdeLis2016]

## References