### 2017-03-01 analysis/spatial_focal_tree.Rmd
change point apparence

---
title: "Field work 2016"
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: fill
    orientation: rows
---

```{r setup, include=FALSE}
library('flexdashboard')
```

```{r pacakges, warning=FALSE, message=FALSE}
library('rgdal')
library("leaflet") 
library("sp")
library("raster")
library("tidyverse")
```

```{r readData}
# mymachine <- '/Users/ajpeluLap/' 
mymachine <- '/Users/ajpelu/' 
di <- paste0(mymachine, 'Dropbox/phd/phd_repos/qpyr_dendro/')

# Read spatial data 
# Centroid of focal tree 
## filename 
mygpx_ca <- paste0(di, 'data_raw/geoinfo/dendro_ca.GPX')
mygpx_sj <- paste0(di, 'data_raw/geoinfo/dendro_sj.GPX')

## reading  
field_work_ca <- readOGR(mygpx_ca, layer="waypoints", verbose = FALSE)
field_work_sj <- readOGR(mygpx_sj, layer="waypoints", verbose = FALSE)
```

```{r prepareBaseMap}
nombre_capas_base <- c("Ortophotography 2009", "Topographical")
nombre_capas_overlay <- c("Field Work 2016")

mymap <- leaflet() %>% 
  # BaseMaps 
  addWMSTiles('http://www.ideandalucia.es/wms/ortofoto2009?',
              layers = 'oca10_2008-09',
              options = WMSTileOptions(format = "image/png", transparent = FALSE),
              attribution = '<a href="http://www.juntadeandalucia.es/institutodeestadisticaycartografia" target="_blank">Instituto de Estadística y Cartografía de Andalucía</a>',
              group = 'Ortophotography 2009') %>%
   addWMSTiles('http://www.ideandalucia.es/services/toporaster10/wms?',
              layers = 'toporaster10',
              options = WMSTileOptions(format = "image/png", transparent = FALSE),
              attribution = '<a href="http://www.juntadeandalucia.es/institutodeestadisticaycartografia" target="_blank">Instituto de Estadística y Cartografía de Andalucía</a>',
              group = 'Topographical') %>%

  # Layers control
  addLayersControl(position = 'bottomright',
                   baseGroups = nombre_capas_base,
                   overlayGroups = nombre_capas_overlay,
                   options = layersControlOptions(collapsed = FALSE))
```

San Juan
=======================================================================

```{r}
# Spatial object 
fw <- field_work_sj

# Set spatial extension 
myext <- extent(fw)

mymap %>% 
  fitBounds(myext@xmin, myext@ymin, myext@xmax, myext@ymax) %>% 
  addCircles(fw, 
             lng=coordinates(fw)[,'coords.x1'], 
             lat=coordinates(fw)[,'coords.x2'], 
             group='Field Work 2016', radius=5, 
             color = 'red', popup=fw$name, 
             stroke = TRUE, fillOpacity = 0.5)

```

Caniar
=======================================================================

```{r}
# Spatial object 
fw <- field_work_ca

# Set spatial extension 
myext <- extent(fw)

mymap %>% 
  fitBounds(myext@xmin, myext@ymin, myext@xmax, myext@ymax) %>% 
  addCircles(fw, 
             lng=coordinates(fw)[,'coords.x1'], 
             lat=coordinates(fw)[,'coords.x2'], 
             group='Field Work 2016', radius=5, 
             color = 'red', popup=fw$name, 
             stroke = TRUE, fillOpacity = 0.5)

```

```{r eval=FALSE, echo=FALSE}

# Data 
ot <- read.csv(paste0(di,'data_raw/geoinfo/other_projects.csv'), 
               header = TRUE, sep = ',')
mig <- readOGR(paste0(di, 'data_raw/geoinfo/'), layer = 'migrame_transectos', verbose = FALSE)

# field wok points 
fw <- readOGR(paste0(di, 'data_raw/geoinfo/'), layer = 'dendro_garmin', verbose = FALSE)

mygpx <- paste0(di, 'data_raw/geoinfo/points_dendro_rep.GPX')
my_points <- readOGR(mygpx, layer = 'waypoints', verbose = TRUE)

# Spatial info
ots <- SpatialPointsDataFrame(coords = ot[,c('x','y')], data = ot,
                               proj4string = CRS("+proj=utm +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))

# Reproject 
mig <- spTransform(mig, CRS("+init=epsg:4236"))
ots <- spTransform(ots, CRS("+init=epsg:4236"))
fw <- spTransform(fw, CRS("+init=epsg:4236"))

# Select only transect of forest (migrame)
mig_forest <- mig[mig$tipo == "ROBLEDAL",]
mig_forest_ca <- mig_forest[mig_forest$localidad == "CANAR",]
mig_forest_sj <- mig_forest[mig_forest$localidad == "SANJUAN",]

ots_phe <- ots[ots$notes == 'Fenologia_Obsnev',]
ots_gea <- ots[ots$notes == 'GeaIzquierdo2014',]
myext <- extent(mig_forest_ca)

mymap %>%
  fitBounds(myext@xmin, myext@ymin, myext@xmax, myext@ymax) %>%
  addPolylines(data=mig_forest_ca, color='blue', group='Migrame', popup=mig_forest_ca$nombre) %>%
  addCircles(ots_phe,
             lng=coordinates(ots_phe)[,'x'],
             lat=coordinates(ots_phe)[,'y'],
             group='OBSNEV Fen', radius=2, color = 'red', popup=ots_phe$name) %>%
  addCircles(ots_gea,
             lng=coordinates(ots_gea)[,'x'],
             lat=coordinates(ots_gea)[,'y'],
             group='Gea Izquierdo 2014', radius=2, color = 'green', popup=ots_gea$name) %>%
  addCircles(fw,
             lng=coordinates(fw)[,'coords.x1'],
             lat=coordinates(fw)[,'coords.x2'],
group='Field Work 2016', radius=2, color = 'yellow', popup=fw$name) %>%
     addCircles(my_points,
             lng=coordinates(my_points)[,'coords.x1'],
             lat=coordinates(my_points)[,'coords.x2'],
             group='Field Work rep ', radius=2, color = 'purple', popup=my_points$name)

```