### 2017-05-10 analysis/analysis_competence.Rmd
update competence analysis

---
title: "Competence distance"
author: "AJ Perez-Luque (@ajpelu)"
date: "2017 Feb"
output:  
  md_document:
    variant: markdown_github
---

```{r, warning=FALSE, message=FALSE}}
library("tidyverse")
library("stringr")
```

# Read data 
```{r, echo=FALSE}
#machine <- 'ajpelu'
machine <- 'ajpeluLap'
di <- paste('/Users/', machine, '/Dropbox/phd/phd_repos/qpyr_dendro/', sep='')

# density data 
compete <- read.csv(file=paste0(di, '/data_raw/dendro_competence.csv'), header=TRUE, sep=',')
```

```{r}
# Compute diameter normal (in meters)
compete <- compete %>% 
  mutate(dn = (perim_mm/1000) /pi) 
```

## Compute competence indexes

* Get unique identifier of focal tree 
* Competence indexes: 
    *

```{r}
# Get name of focal trees
focal_trees <- compete %>% filter(id_focal != 'Fresno') 
focal_trees <- unique(focal_trees$id_focal)

# Empty object to store results 
output<- vector("list", length(focal_trees))

for (i in 1:length(focal_trees)){

  # Create df with all trees inside plot (10 m radius)
  df <- compete %>% filter(id_focal == focal_trees[[i]])

  # Compute fixed variables: 
  radius_plot <- 10 # radius plot (10 m)
  sf_plot <- pi * radius_plot^2 # surface plot (m2)
  dnft <- filter(df, sp== 'Focal')$dn # normal diameter of focal tree    
  # ----------

  # Competence indices 
  ## Number of competitors inside plot wihtout focal tree (r = 10)
  n_competitors <- filter(df, sp != 'Focal') %>% count()

  ## Number of competitors with dn_j > dn_i (focal tree)
  n_competitors_hihger <- df %>%
    filter(sp != 'Focal') %>%
    filter(dn > dnft) %>% 
    count()

  ## plot density (tree/plot)
  plot_density <- n_competitors / sf_plot

  ## Stand density (tree / ha) 
  stand_density <- (n_competitors * 10000) / sf_plot

  ## Sum sizes 
  sum_sizes <- df %>% 
    filter(sp != 'Focal') %>%
    summarise(sum_sizes=sum(dn))

  ## Size ratio 
  size_ratio <- dnft / (dnft + sum_sizes$sum_sizes)

  ## Distance nearest tree
  dist_nn <- df %>% 
    filter(sp != 'Focal') %>% 
    slice(which.min(distance_cm))

  dist_nn <- (dist_nn$distance_cm / 100) #in meters

  ## Various 
  ### Basal area
  ### Size ratio proportional to distance 
  ### Size difference proportional to distance 
  ### Negative exponential size ratio
  ### Negative exponential weighted size ratio
  ### Lorimer 
  ### Crowding

  df_more_indices <- df %>%
    filter(sp != 'Focal') %>%
    mutate(basal_area = ((pi/4) * dn * dn), 
           size_ratio_distance = (dn / dnft) * (1 / ((distance_cm/100) + 1)),
           size_diff_distance = (dn - dnft) / ((distance_cm/100) + 1),
           neg_exp_size_ratio = (dn / dnft) * (1 / exp((distance_cm/100) + 1)),
           neg_exp_w_size_ratio = (dn / dnft) * exp((-(distance_cm/100) + 1) / (dn + dnft)),
           lorimer = (dn / dnft) / (sqrt((distance_cm/100) / radius_plot)),
           crowding = dn / (distance_cm/100))

  df_indices <- df_more_indices %>% 
    summarise(ba = sum(basal_area),
              srd = sum(size_ratio_distance),
              sdd = sum(size_diff_distance),
              nesr = sum(neg_exp_size_ratio),
              newsr = sum(neg_exp_w_size_ratio),
              lorimer = sum(lorimer),
              crowding = sum(crowding)) %>% 
    mutate(nc = n_competitors,
           nch = n_competitors_hihger,
           pd = plot_density,
           std = stand_density,
           ss = sum_sizes,
           sr = size_ratio,
           dnn = dist_nn,
           id_focal = focal_trees[[i]])

  output[[i]] <- df_indices

}

df_indices_plot <- bind_rows(output)

rm(n_competitors, n_competitors_hihger, plot_density, stand_density, sum_sizes, size_ratio, dist_nn, df, df_indices, output, df_more_indices, dnft)

# Export data

```

## Plot

```{r}
source(paste0(getwd(), '/script/R/get_coords_trees.R'))

# Add variables with coordinates of each tree. Use get_coords_trees() function 
compete <- compete %>% 
  mutate(xcord = get_coords_trees(angle= rumbo_degree, 
                           distance = distance_cm/100,
                           x0 = 0, y0 = 0)$x,
         ycord = get_coords_trees(angle = rumbo_degree, 
                           distance = distance_cm/100,
                           x0 = 0, y0 = 0)$y)
```

## Plot by site 
```{r}
# Set levels of eleveation 
sj_lowcode  <- paste0('A', str_pad(1:10, 2, pad='0'))
sj_highcode <- paste0('A', 11:20)
ca_lowcode <- c(paste0('B', str_pad(1:10, 2, pad='0')),
            paste0('B', 26:30))
ca_highcode <- paste0('B', 11:25)

compete <- compete %>% 
  mutate(elev = ifelse(id_focal %in% sj_lowcode, 'Low',
                       ifelse(id_focal %in% sj_highcode, 'High',
                              ifelse(id_focal %in% ca_lowcode, 'Low', 'High')))) %>%
  mutate(site = paste0(loc, '_', elev))
```

# see stackoverflow questions 6862742
gg_circle <- function(radius, xcenter, ycenter, color='black', fill=NA, ...){
  x <- xcenter + radius*cos(seq(0,pi, length.out = 100))
  ymax <- ycenter + radius*sin(seq(0, pi, length.out = 100))
  ymin <- ycenter + radius*sin(seq(0, -pi, length.out = 100))
  annotate("ribbon", x=x, ymin=ymin, ymax=ymax, color=color, fill=fill, ...)
}
```

compete %>%  
  filter(sp != 'Focal') %>%
  ggplot(aes(x=xcord, y=ycord)) + 
  facet_wrap(~id_focal) + 
  geom_point(aes(size=perim_mm), shape=21, fill='transparent') + 
  xlim(-11,11) + ylim(-11,11) + 
  geom_point(aes(x=0, y=0), color='green') +
  theme_bw() +
  gg_circle(radius = 10, xcenter = 0, ycenter = 0) +
  theme(panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank()) 
```

```{r}

```