### 2017-05-10 analysis/analysis_competence.Rmd
update references and include equations

---
title: "Analysis of Competence Index"
author: "AJ Perez-Luque (@ajpelu)"
date: "2017 Feb"
output:  
  md_document:
    variant: markdown_github
bibliography: references.bib 
csl: ecology.csl
---

```{r, warning=FALSE, message=FALSE}
library("tidyverse")
library("stringr")
```

# Read data 
```{r, echo=FALSE}
machine <- 'ajpelu'
# machine <- 'ajpeluLap'
di <- paste('/Users/', machine, '/Dropbox/phd/phd_repos/qpyr_dendro/', sep='')

# density data 
compete <- read.csv(file=paste0(di, '/data_raw/dendro_competence.csv'), header=TRUE, sep=',')
```

```{r}
# Compute diameter normal (in meters)
compete <- compete %>% 
  mutate(dn = (perim_mm/1000) /pi) 
```

## Compute competition indices

The following competition indices were computed. See @GeaIzquierdo2009 for more info.

### Distance-Independet Indices

* Basal Area (**ba**)

$$ BA = \sum_{i=1}^{n}pi\cdot (dbh_{i} / 2)^2 $$

* Stand Density (**std**)

$$ std = n_{trees} / \mathrm{plot\, area(ha)} $$

* Plot Density (**pd**)

$$ pd = n_{trees} / \mathrm{plot\, area}$$

* Number of competitors (**n_competitors**): number of competitors within *r* meters (10 m)

* Number of competitors higher (**n_competitors_higher**): number of competitors within *r* meters (10 m) such that $ dbh_j > dbh_i $

* Sum of size (**sum_sizes**) of trees within *r* meters (10 m): $ \sum dbh_{j} $

* Size ratio (**sr**)

$$ sr = dbh_i / \sum_{i=1}^{n}dbh_i $$

### Distance-Dependet Indices

* Distance to nearest tree (**dnn**)

* Crowding

$$ \mathrm{crowding} = \sum_{i=1}^{n}(dbh_i / dist_{ij}) $$

* Lorimer

$$ \mathrm{lorimer} = \sum_{i=1}^{n}\left [ (dbh_i / dbh_i) / \sqrt{dist_{ij}/r} \right ] $$

* Negative Exponential size ratio (**nesr**)

$$ \mathrm{nesr} = \sum_{i=1}^{n} (dbh_j / dbh_i) \cdot \left[ 1 / \exp(dist_{ij} + 1) \right ] $$

* Negative Exponential Weighted size ratio (**newsr**)

$$ \mathrm{newsr} = \sum_{i=1}^{n} (dbh_j / dbh_i) \cdot \exp\left[ (-dist_{ij} + 1) / (dbh_i + dbh_j) \right ] $$

* Size ratio proportional to distance (**srd**)

$$ \mathrm{srd} = \sum_{i=1}^{n} (dbh_j / dbh_i) \cdot \left[1/ (dist_{ij} + 1) \right ] $$

* Size difference proportional to distance (**sdd**)

$$ \mathrm{sdd} = \sum_{i=1}^{n} \left[ (dbh_j  - dbh_i) / (dist_{ij} + 1) \right ] $$

```{r}
# 1 Get name of focal trees
focal_trees <- compete %>% filter(id_focal != 'Fresno') 
focal_trees <- unique(focal_trees$id_focal)

# 2 Compute Fixed variables 
radius_plot <- 10 # radius plot (10 m)
sf_plot <- pi * radius_plot^2 # surface plot (m2)

# 3 Empty object to store results 
output<- vector("list", length(focal_trees))

# 4 Loop 
for (i in 1:length(focal_trees)){

  # Create df with all trees inside plot (10 m radius)
  df <- compete %>% filter(id_focal == focal_trees[[i]])

  # Focal tree
  focal_t <- filter(df, sp== 'Focal') 
  ## Normal diameter of focal tree
  dnft <- focal_t$dn

  # Competence indices 
  ## Number of competitors inside plot wihtout focal tree (r = 10)
  n_competitors <- filter(df, sp != 'Focal') %>% count() %>% transmute(n_competitors = n)

  ## Number of competitors with dn_j > dn_i (focal tree)
  n_competitors_higher <- df %>%
    filter(sp != 'Focal') %>%
    filter(dn > dnft) %>% 
    count() %>% 
    transmute(n_competitors_higher = n)

  ## plot density (tree/plot)
  plot_density <- n_competitors$n_competitors / sf_plot

  ## Stand density (tree / ha) 
  stand_density <- (n_competitors$n_competitors * 10000) / sf_plot

  ## Sum sizes 
  sum_sizes <- df %>% 
    filter(sp != 'Focal') %>%
    summarise(sum_sizes=sum(dn))

  ## Size ratio 
  size_ratio <- dnft / (dnft + sum_sizes$sum_sizes)

  ## Distance nearest tree
  dist_nn <- df %>% 
    filter(sp != 'Focal') %>% 
    slice(which.min(distance_cm))

  dist_nn <- (dist_nn$distance_cm / 100) #in meters

  ## Various 
    ### Basal area
    ### Size ratio proportional to distance 
    ### Size difference proportional to distance 
    ### Negative exponential size ratio
    ### Negative exponential weighted size ratio
    ### Lorimer 
    ### Crowding

  df_more_indices <- df %>%
    filter(sp != 'Focal') %>%
    mutate(basal_area = ((pi/4) * dn * dn), 
           size_ratio_distance = (dn / dnft) * (1 / ((distance_cm/100) + 1)),
           size_diff_distance = (dn - dnft) / ((distance_cm/100) + 1),
           neg_exp_size_ratio = (dn / dnft) * (1 / exp((distance_cm/100) + 1)),
           neg_exp_w_size_ratio = (dn / dnft) * exp((-(distance_cm/100) + 1) / (dn + dnft)),
           lorimer = (dn / dnft) / (sqrt((distance_cm/100) / radius_plot)),
           crowding = dn / (distance_cm/100))

  df_indices <- df_more_indices %>% 
    summarise(ba = sum(basal_area),
              srd = sum(size_ratio_distance),
              sdd = sum(size_diff_distance),
              nesr = sum(neg_exp_size_ratio),
              newsr = sum(neg_exp_w_size_ratio),
              lorimer = sum(lorimer),
              crowding = sum(crowding)) %>% 
    mutate(n_competitors,
           n_competitors_higher,
           pd = plot_density,
           std = stand_density,
           sum_sizes,
           sr = size_ratio,
           dnn = dist_nn,
           id_focal = focal_trees[[i]])

    output[[i]] <- df_indices

}

# df_indices_plot <- bind_rows(output)
# As the output is a nested list I used this solution (http://stackoverflow.com/questions/27930883/converting-elements-in-a-nested-list-to-dataframe) to convert to a dataframe

df_indices_plot <- bind_rows(lapply(output, as.data.frame.list, stringsAsFactors=FALSE))

rm(n_competitors, n_competitors_higher, plot_density, stand_density, sum_sizes, size_ratio, dist_nn, df, df_indices, output, df_more_indices, dnft)

# Export data 
write.csv(df_indices_plot, file=paste(di, "data/competence/competence_indexes.csv", sep=""), row.names = FALSE)
```

## Plot

```{r}
source(paste0(di, '/script/R/get_coords_trees.R'))

# Add variables with coordinates of each tree. Use get_coords_trees() function 
compete <- compete %>% 
  mutate(xcord = get_coords_trees(angle= rumbo_degree, 
                           distance = distance_cm/100,
                           x0 = 0, y0 = 0)$x,
         ycord = get_coords_trees(angle = rumbo_degree, 
                           distance = distance_cm/100,
                           x0 = 0, y0 = 0)$y)
```

## Plot by site

```{r}
# Set levels of eleveation 
sj_lowcode  <- paste0('A', str_pad(1:10, 2, pad='0'))
sj_highcode <- paste0('A', 11:20)
ca_lowcode <- c(paste0('B', str_pad(1:10, 2, pad='0')),
            paste0('B', 26:30))
ca_highcode <- paste0('B', 11:25)

compete <- compete %>% 
  mutate(elev = ifelse(id_focal %in% sj_lowcode, 'Low',
                       ifelse(id_focal %in% sj_highcode, 'High',
                              ifelse(id_focal %in% ca_lowcode, 'Low', 'High')))) %>%
  mutate(site = paste0(loc, '_', elev))
```

#### San Juan 
```{r, echo=FALSE, fig.height=8, fig.width=10}
# see stackoverflow questions 6862742
gg_circle <- function(radius, xcenter, ycenter, color='black', fill=NA, ...){
  x <- xcenter + radius*cos(seq(0,pi, length.out = 100))
  ymax <- ycenter + radius*sin(seq(0, pi, length.out = 100))
  ymin <- ycenter + radius*sin(seq(0, -pi, length.out = 100))
  annotate("ribbon", x=x, ymin=ymin, ymax=ymax, color=color, fill=fill, ...)
}

compete %>%  
  filter(sp != 'Focal') %>%
  filter(loc == 'SJ') %>% 
  ggplot(aes(x=xcord, y=ycord)) + 
  facet_wrap(~id_focal) + 
  geom_point(aes(size=dn), shape=21, fill='transparent') +
  xlim(-11,11) + ylim(-11,11) + 
  geom_point(aes(x=0, y=0), color='green') +
  theme_bw() +
  gg_circle(radius = 10, xcenter = 0, ycenter = 0) +
  theme(panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank(), 
        strip.background = element_rect(colour = "black", fill ='transparent')) 
```

#### Canar 
```{r, echo=FALSE, fig.height=9, fig.width=12}
compete %>%  
  filter(sp != 'Focal') %>%
  filter(loc == 'CA') %>% 
  ggplot(aes(x=xcord, y=ycord)) + 
  facet_wrap(~id_focal) + 
  geom_point(aes(size=dn), shape=21, fill='transparent') +
  xlim(-11,11) + ylim(-11,11) + 
  geom_point(aes(x=0, y=0), color='green') +
  theme_bw() +
  gg_circle(radius = 10, xcenter = 0, ycenter = 0) +
  theme(panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank(), 
        strip.background = element_rect(colour = "black", fill ='transparent')) 
```

## References