### 2017-05-15 analysis/analysis_stand_features.Rmd
update function; add min max; loop for topo variables; close output

---
title: "Analysis Stand Features "
author: "AJ Perez-Luque (@ajpelu)"
date: "2017 May"
output:  
  md_document:
    toc: true
    variant: markdown_github
---

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library("tidyverse")
library("raster")
library("stringr")
library('rgdal')
library("sp")
library("knitr")
library("broom")
library("multcomp")
library("pander")
library("devtools")
source_gist("https://gist.github.com/ajpelu/194e721077ec045a2b864088908e7aca", filename = 'table_glht.r') # multicomp table
```

# Prepare Data
```{r, echo=FALSE}
machine <- 'ajpelu'
# machine <- 'ajpeluLap'
di <- paste0('/Users/', machine, '/Dropbox/phd/phd_repos/qpyr_dendro/', sep='')

# Read tree data
tree <- read.csv(file=paste0(di, '/data_raw/dendro_competence.csv'), header=TRUE, sep=',')

# Read spatial data 
# Centroid of focal tree 
## filename 
mygpx_ca <- paste0(di, 'data_raw/geoinfo/dendro_ca.GPX')
mygpx_sj <- paste0(di, 'data_raw/geoinfo/dendro_sj.GPX')

## reading  
field_work_ca <- readOGR(mygpx_ca, layer="waypoints", verbose = FALSE)
field_work_sj <- readOGR(mygpx_sj, layer="waypoints", verbose = FALSE)
```

* Two datasets: focal tree and competence

```{r}
# Compute diameter (mm)
tree <- tree %>% 
  mutate(dn_mm = (perim_mm / pi))

# Set levels of eleveation 
sj_lowcode  <- paste0('A', str_pad(1:10, 2, pad='0'))
sj_highcode <- paste0('A', 11:20)
ca_lowcode <- c(paste0('B', str_pad(1:10, 2, pad='0')),
            paste0('B', 26:30))
ca_highcode <- paste0('B', 11:25)

tree <- tree %>% 
  mutate(elevF = ifelse(id_focal %in% sj_lowcode, 'Low',
                       ifelse(id_focal %in% sj_highcode, 'High',
                              ifelse(id_focal %in% ca_lowcode, 'Low', 'High')))) %>%
  mutate(site = paste0(loc, '_', elevF))

# Get only focal trees  
ft <- tree %>% 
  filter(sp=='Focal') %>% 
  filter(id_focal!='Fresno')

# Get only no focal trees 
nft <- tree %>% 
  filter(sp!='Focal') 
```

# General Variables 
Numbers of focal trees by site

```{r}
general_var <- ft %>% group_by(loc, site) %>% count()

general_var %>% kable
```

# Spatial Info
Coordinates of the centroid for each site

```{r, spatialData} 
## Get coordinates of spatial data  
sp_ca <- as.data.frame(field_work_ca) %>% 
  dplyr::select(ele, name, lat = coords.x2, long = coords.x1)

sp_sj <- as.data.frame(field_work_sj) %>% 
  dplyr::select(ele, name, lat = coords.x2, long = coords.x1)

## Add code site and get centroid (see # http://rspatial.org/analysis/rst/8-pointpat.html)
sp_ca <- sp_ca %>% 
  mutate(loc = 'CA', 
         elevF = ifelse(name %in% ca_lowcode, 'Low', 'High'),
         site = paste0(loc, '_', elevF))

coord_ca <- sp_ca %>% 
  group_by(site) %>%
  summarise(lat_m = mean(lat),
            long_m = mean(long))

# plot(sp_ca$long, sp_ca$lat, pch=19, col='gray')
# points(coord_ca$long_m, coord_ca$lat_m, pch=19, col='blue')

sp_sj <- sp_sj %>% 
  mutate(loc = 'SJ', 
         elevF = ifelse(name %in% sj_lowcode, 'Low', 'High'),
         site = paste0(loc, '_', elevF))        
coord_sj <- sp_sj %>% 
  group_by(site) %>%
  summarise(lat_m = mean(lat),
            long_m = mean(long))

coords_sites <- coord_sj %>% rbind(coord_ca)

#plot(sp_sj$long, sp_sj$lat, pch=19, col='gray')
#points(coord_sj$long_m, coord_sj$lat_m, pch=19, col='blue')

coords_sites %>% kable()
```

# Competition data

* Read data 
```{r, echo=FALSE}
# Read competition
compe <- read.csv(file=paste(di, "/data/competence/competence_indices.csv", sep=""), header=TRUE, sep=',')

compe <- compe %>% 
  mutate(name = id_focal) %>% 
  mutate(loc = ifelse(str_detect(name, "A"), 'SJ', 'CA'),
         elevF = ifelse(name %in% sj_lowcode, 'Low',
                        ifelse(name %in% sj_highcode, 'High',
                              ifelse(name %in% ca_lowcode, 'Low', 'High')))) %>%
  mutate(site = paste0(loc, '_', elevF)) %>% 
  mutate(site = as.factor(site),
         loc = as.factor(loc),
         elevF = as.factor(elevF))

```

* Create a custom function to compare between sites (aov & post hoc)

```{r, echo=FALSE}
#### Function to comparison (aov)
compara <- function(df, mivariable){ 
  require('dplyr')
  require('multcomp')
  require('broom') 
  # mivariable: variabe of interest

  output <- list()

  # Model formulation
  myformula <- as.formula(paste0(mivariable, " ~ site"))

  # AOV 
  mymodel <- aov(myformula, data=df)

  # Summary AOV (broom style)
  tm <- broom::tidy(mymodel)

  # Multiple comparison 
  tuk <- glht(mymodel, linfct = mcp(site = "Tukey"))

  # Convert comparisons into letters 
  df_letter <- fortify(cld(tuk)) %>% 
    transmute(site = as.factor(lhs),
         tukey = letters) %>%
    mutate(variable = mivariable)

  aux_name <- paste0('df_tuk_', mivariable)
  assign(aux_name, df_letter)

  mymult <- as.data.frame(table_glht(tuk))

  output$mymodel <- mymodel
  output$tm <- tm
  output$aux_name <- df_letter
  output$mymult <- mymult

  # Return also summary values 
  summ_values <- df %>% 
    dplyr::select_(mivariable, 'site') %>%
    group_by(site) %>% 
    summarise_each(funs(mean, sd, se=sd(.)/sqrt(n()), min, max))

  # Join letters and summary values
  output$summ_values <- summ_values

  summ_comparison <- summ_values %>% left_join(df_letter, by='site')

  output$summ_comparison <- summ_comparison

  return(output)
}

```

* Export data into text files (see `/out/anovas_competition/`)

```{r, echo=FALSE}
## Comparison 
# Select only variables to compare 
compe_sel <- compe %>% dplyr::select(ba:dnn,site)

# Get vector with variables 
variables <- compe %>% dplyr::select(ba:dnn) %>% names()

for (i in variables){

  # apply comparison
  out_compara <- compara(df=compe_sel, mivariable = i)

  out_name  <- paste0('aov_', i)
  assign(out_name, out_compara)

}

# Loop to export into txt files (see ./out/anovas_competition ... )
for (i in variables){

  out <- get(paste0('aov_', i))

  file_out <- file(paste0(di,'out/anovas_competition/aov_', i, '.txt'))
  sink(file_out)
  cat("MODEL \n")
  print(out$mymodel) 
  cat("\n")

  cat("MODEL pretty \n")
  print(out$tm) 
  cat("\n")

  cat("POST HOC \n")
  print(out$mymult) 
  cat("\n")

  cat("SUMMARY VALUES \n")
  print(out$summ_comparison) 
  cat("\n")

  sink()
  close(file_out)
}
```

## Distance-Independet Indices 
### Basal Area
```{r, echo=FALSE}
mivariable <- 'ba'
s <- get(paste0('aov_', mivariable))

s$summ_comparison %>% 
  pander(round=4, caption=paste0('Mean values (', mivariable,')'))

tm <- s$tm

pander(tm, round=5,
       caption=paste0('ANOVA table (', mivariable,')'), 
       missing = '', emphasize.strong.cells = 
               which(tm < 0.1 & tm == tm$p.value, arr.ind = T))
```

### Stand Density
```{r, echo=FALSE}
mivariable <- 'std'
s <- get(paste0('aov_', mivariable))

s$summ_comparison %>% 
  pander(round=4, caption=paste0('Mean values (', mivariable,')'))

tm <- s$tm

pander(tm, round=5,
       caption=paste0('ANOVA table (', mivariable,')'), 
       missing = '', emphasize.strong.cells = 
               which(tm < 0.1 & tm == tm$p.value, arr.ind = T))
```

### Plot Density
```{r, echo=FALSE}
mivariable <- 'pd'
s <- get(paste0('aov_', mivariable))

s$summ_comparison %>% 
  pander(round=4, caption=paste0('Mean values (', mivariable,')'))

tm <- s$tm

pander(tm, round=5,
       caption=paste0('ANOVA table (', mivariable,')'), 
       missing = '', emphasize.strong.cells = 
               which(tm < 0.1 & tm == tm$p.value, arr.ind = T))
```

### Number of competitors within *r* meters (10 m)
```{r, echo=FALSE}
mivariable <- 'n_competitors'
s <- get(paste0('aov_', mivariable))

s$summ_comparison %>% 
  pander(round=4, caption=paste0('Mean values (', mivariable,')'))

tm <- s$tm

pander(tm, round=5,
       caption=paste0('ANOVA table (', mivariable,')'), 
       missing = '', emphasize.strong.cells = 
               which(tm < 0.1 & tm == tm$p.value, arr.ind = T))
```

### Number of competitors within *r* meters (10 m) such that $ dbh_j > dbh_i $
```{r, echo=FALSE}
mivariable <- 'n_competitors_higher'
s <- get(paste0('aov_', mivariable))

s$summ_comparison %>% 
  pander(round=4, caption=paste0('Mean values (', mivariable,')'))

tm <- s$tm

pander(tm, round=5,
       caption=paste0('ANOVA table (', mivariable,')'), 
       missing = '', emphasize.strong.cells = 
               which(tm < 0.1 & tm == tm$p.value, arr.ind = T))
```

### Sum of size of trees within *r* meters (10 m)

```{r, echo=FALSE}
mivariable <- 'sum_sizes'
s <- get(paste0('aov_', mivariable))

s$summ_comparison %>% 
  pander(round=4, caption=paste0('Mean values (', mivariable,')'))

tm <- s$tm

pander(tm, round=5,
       caption=paste0('ANOVA table (', mivariable,')'), 
       missing = '', emphasize.strong.cells = 
               which(tm < 0.1 & tm == tm$p.value, arr.ind = T))
```

## Size ratio

```{r, echo=FALSE}
mivariable <- 'sr'
s <- get(paste0('aov_', mivariable))

s$summ_comparison %>% 
  pander(round=4, caption=paste0('Mean values (', mivariable,')'))

tm <- s$tm

pander(tm, round=5,
       caption=paste0('ANOVA table (', mivariable,')'), 
       missing = '', emphasize.strong.cells = 
               which(tm < 0.1 & tm == tm$p.value, arr.ind = T))
```

## Distance-Dependet Indices

### Distance to nearest tree       
```{r, echo=FALSE}
mivariable <- 'dnn'
s <- get(paste0('aov_', mivariable))

s$summ_comparison %>% 
  pander(round=4, caption=paste0('Mean values (', mivariable,')'))

tm <- s$tm

pander(tm, round=5,
       caption=paste0('ANOVA table (', mivariable,')'), 
       missing = '', emphasize.strong.cells = 
               which(tm < 0.1 & tm == tm$p.value, arr.ind = T))
```

### Crowding       
```{r, echo=FALSE}
mivariable <- 'crowding'
s <- get(paste0('aov_', mivariable))

s$summ_comparison %>% 
  pander(round=4, caption=paste0('Mean values (', mivariable,')'))

tm <- s$tm

pander(tm, round=5,
       caption=paste0('ANOVA table (', mivariable,')'), 
       missing = '', emphasize.strong.cells = 
               which(tm < 0.1 & tm == tm$p.value, arr.ind = T))
```

### Lorimer

```{r, echo=FALSE}
mivariable <- 'lorimer'
s <- get(paste0('aov_', mivariable))

s$summ_comparison %>% 
  pander(round=4, caption=paste0('Mean values (', mivariable,')'))

tm <- s$tm

pander(tm, round=5,
       caption=paste0('ANOVA table (', mivariable,')'), 
       missing = '', emphasize.strong.cells = 
               which(tm < 0.1 & tm == tm$p.value, arr.ind = T))
```

### Negative Exponential size ratio

```{r, echo=FALSE}
mivariable <- 'nesr'
s <- get(paste0('aov_', mivariable))

s$summ_comparison %>% 
  pander(round=4, caption=paste0('Mean values (', mivariable,')'))

tm <- s$tm

pander(tm, round=5,
       caption=paste0('ANOVA table (', mivariable,')'), 
       missing = '', emphasize.strong.cells = 
               which(tm < 0.1 & tm == tm$p.value, arr.ind = T))
```

### Negative Exponential Weighted size ratio
```{r, echo=FALSE}
mivariable <- 'newsr'
s <- get(paste0('aov_', mivariable))

s$summ_comparison %>% 
  pander(round=4, caption=paste0('Mean values (', mivariable,')'))

tm <- s$tm

pander(tm, round=5,
       caption=paste0('ANOVA table (', mivariable,')'), 
       missing = '', emphasize.strong.cells = 
               which(tm < 0.1 & tm == tm$p.value, arr.ind = T))
```

### Size ratio proportional to distance

```{r, echo=FALSE}
mivariable <- 'srd'
s <- get(paste0('aov_', mivariable))

s$summ_comparison %>% 
  pander(round=4, caption=paste0('Mean values (', mivariable,')'))

tm <- s$tm

pander(tm, round=5,
       caption=paste0('ANOVA table (', mivariable,')'), 
       missing = '', emphasize.strong.cells = 
               which(tm < 0.1 & tm == tm$p.value, arr.ind = T))
```

### Size difference proportional to distance

```{r, echo=FALSE}
mivariable <- 'sdd'
s <- get(paste0('aov_', mivariable))

s$summ_comparison %>% 
  pander(round=4, caption=paste0('Mean values (', mivariable,')'))

tm <- s$tm

pander(tm, round=5,
       caption=paste0('ANOVA table (', mivariable,')'), 
       missing = '', emphasize.strong.cells = 
               which(tm < 0.1 & tm == tm$p.value, arr.ind = T))
```

# Topographic data

* Read data 
```{r}
# Read topo data 
topo <- read.csv(file=paste(di, "/data/topo/topo.csv", sep=""), header=TRUE, sep=',')

topo <- topo %>% 
  mutate(loc = ifelse(str_detect(name, "A"), 'SJ', 'CA'),
         elevF = ifelse(name %in% sj_lowcode, 'Low',
                        ifelse(name %in% sj_highcode, 'High',
                              ifelse(name %in% ca_lowcode, 'Low', 'High')))) %>%
  mutate(site = paste0(loc, '_', elevF)) %>% 
  mutate(site = as.factor(site),
         loc = as.factor(loc),
         elevF = as.factor(elevF))

topo_summary <- topo %>%
  group_by(site) %>%
  summarise(mde_m = mean(mde),
            mde_sd = sd(mde),
            mde_min = min(mde),
            mde_max = max(mde),
            slope_m = mean(slope),
            slope_sd = sd(slope))

# Another way to obtain summary values 
variables <- c('mde','slope','aspect')
auxdf <- data.frame()

for (i in variables){ 
aux <- topo %>% 
  dplyr::group_by(site) %>% 
  summarise_each_(funs(mean, sd, se=sd(.)/sqrt(n())), i) %>% mutate(variable=i)

auxdf <- rbind(auxdf, aux) }
```

* Compare data and export data into text files (see `/out/anovas_topo/`)

```{r, echo=FALSE}
## Comparison 
# Select only variables to compare

topo_sel <- topo %>% dplyr::select(mde, slope, site)

# Get vector with variables 
variables <- c('mde','slope')

for (i in variables){

  # apply comparison
  out_compara <- compara(df=topo_sel, mivariable = i)

  out_name  <- paste0('aov_', i)
  assign(out_name, out_compara)

}

# Loop to export into txt files (see ./out/anovas_topo ... )
for (i in variables){

  out <- get(paste0('aov_', i))

  file_out <- file(paste0(di,'out/anovas_topo/aov_', i, '.txt'))
  sink(file_out)
  cat("MODEL \n")
  print(out$mymodel) 
  cat("\n")

  cat("MODEL pretty \n")
  print(out$tm) 
  cat("\n")

  cat("POST HOC \n")
  print(out$mymult) 
  cat("\n")

  cat("SUMMARY VALUES \n")
  print(out$summ_comparison) 
  cat("\n")

  sink()
  close(file_out)
}
```

### Elevation

```{r, echo=FALSE}
mivariable <- 'mde'
s <- get(paste0('aov_', mivariable))

s$summ_comparison %>% 
  pander(round=4, caption=paste0('Mean values (', mivariable,')'))

tm <- s$tm

pander(tm, round=5,
       caption=paste0('ANOVA table (', mivariable,')'), 
       missing = '', emphasize.strong.cells = 
               which(tm < 0.1 & tm == tm$p.value, arr.ind = T))
```

### Slope

```{r, echo=FALSE}
mivariable <- 'slope'
s <- get(paste0('aov_', mivariable))

s$summ_comparison %>% 
  pander(round=4, caption=paste0('Mean values (', mivariable,')'))

tm <- s$tm

pander(tm, round=5,
       caption=paste0('ANOVA table (', mivariable,')'), 
       missing = '', emphasize.strong.cells = 
               which(tm < 0.1 & tm == tm$p.value, arr.ind = T))
```

### Min & max mde