### 2017-05-30 man/material_methods.Rmd
update material methods

---
title: "material and methods"
author: "AJ Perez-Luque (@ajpelu); G. Gea-Izquierdo"
date: "2017"
output:  
  md_document:
    toc: true
    variant: markdown_github
bibliography: references.bib
csl: ecology.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Materials and methods

## Dendrochronological methods
... dendro
For each focal tree we measured diameter at breast height (DBH) and total height. 
A total of xx trees were sampled. 
...

We built chronologies for each site (two)

Site chronologies were built by averaging all tree BAI measurement of the same site. To explore similarity within locality, each site chronology was smoothed using centred moving averages with different window sizes, and then Pearson's correlation coefficient between the two chronologies of the same locality (higher and lower elevation) were calculated. Significance was tested using 1000 boostrap replicates and with 95 % confidence intervals built using the R packgae `boot` (:red_circle: cite boot).