### 2017-09-15 analysis/tables_resiliences.Rmd
update tables

---
title: "tabla_mean_values Mhuber"
author: "AJ Perez-Luque (@ajpelu)"
date: "2017"
output:  
  md_document:
    variant: markdown_github
bibliography: ../references.bib 
csl: ../ecology.csl
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,warning=FALSE, error=TRUE, message=FALSE)
```

```{r, warning=FALSE, message=FALSE}
library("tidyverse")
library("pander")
```

```{r, echo=FALSE}
machine <- 'ajpelu'
# machine <- 'ajpeluLap'
di <- paste0('/Users/', machine, '/Dropbox/phd/phd_repos/qpyr_dendro/', sep='')
```

# Read data from EVI

```{r}
# Estos datos vienen del repo de modis_resilience (/analysis/explore_Resilience.Rmd)
mhuber_evi <- read.csv(file=paste0(di, '/out/anovas_resilience/huber_evi/robust_mhuber.csv'), header = TRUE, sep=",")
mhuber_eviA <- read.csv(file=paste0(di, '/out/anovas_resilience/huber_evi/robust_mhuber_a.csv'), header = TRUE, sep=",")
mhuber_eviB <- read.csv(file=paste0(di, '/out/anovas_resilience/huber_evi/robust_mhuber_b.csv'), header = TRUE, sep=",")
```

```{r}
# Estos datos vienen del repo de modis_resilience (/analysis/explore_Resilience.Rmd)
mhuber_bai <- read.csv(file=paste0(di, '/out/anovas_resilience/robust_mhuber.csv'), header = TRUE, sep=",")
mhuber_baiA <- read.csv(file=paste0(di, '/out/anovas_resilience/robust_mhuber_a.csv'), header = TRUE, sep=",")
mhuber_baiB <- read.csv(file=paste0(di, '/out/anovas_resilience/robust_mhuber_b.csv'), header = TRUE, sep=",")
```

### Tablas 
```{r}
evi_a <- mhuber_eviA %>% 
  mutate(M.Huber = round(M.Huber, 4),
         lower.ci = round(lower.ci, 4),
         upper.ci = round(upper.ci, 4)) %>% 
  unite("ci", c("lower.ci", "upper.ci"), sep=", ") %>% 
  mutate(ci = paste0('(', ci, ')')) %>% 
  unite("value", c("M.Huber", "ci"), sep= " ") %>%
  dplyr::select(-MonoLetter, -n) %>% 
  unite("value", c("value", "Letter"), sep= "^") %>% 
  dplyr::select(var, disturb_year, value) %>% as.data.frame()

evi_a %>% pander() 
```