### 2017-09-15 man/report.Rmd
update ms and add compile document

---
title: "ms_borrador"
author: "AJ Perez-Luque; G. Gea-Izquierdo; R. Zamora; FJ. Bonet"
date: "2017"
output:  word_document
bibliography: ../references.bib
csl: ../ecology.csl
---

```{r child = 'ms_0_intro.Rmd'}
```

```{r child = 'ms_1_material_methods.Rmd'}
```

```{r child = 'ms_2_results.Rmd'}
```