### 2017-11-07 analysis/drought/explore_drought_spei_regional.Rmd
regional drought analysis from old repo

---
title: "SPEI drought SN"
author: "AJ Perez-Luque (@ajpelu)"
date: "2016 March"
output:  
    md_document:
      variant: markdown_github

---
# Drought data

* Obtain data of drought index SPEI from [SPEI Global Drought Monitor](http://sac.csic.es/spei/map/maps.html)
* Query data from Sierra Nevada [36.75, -3.75], [37.25, -3.25]

```{r metadata, echo=FALSE}
# Set working directory 
machine <- 'ajpelu'
# machine <- 'ajpeluLap'
di <- paste0('/Users/', machine, '/Dropbox/phd/phd_repos/qpyr_dendro/', sep='')

```

```{r packages, warning=FALSE, echo=FALSE, message=FALSE}
# Load packages 
library("dplyr")
library("stringr")
library("lubridate")
library("reshape2")
library("ggplot2")
source(paste0(di,"/script/R/exportpdf.R"))
```

## Prepare data
```{r}

# Get file names
myfiles <- list.files(path = paste(di, "/data_raw/spei_all", sep=""), pattern = "\\.csv$")

mydf <- data.frame()

# Loop to read files and prepare data
for (j in myfiles){ 
  aux <- read.csv(file=paste(di, "/data_raw/spei_all/", j, sep= ""),
              header = TRUE,
              sep = ',')

  # Remove csv and get name 
  name_aux <- str_replace(j, ".csv", "")

  # Get lat long
  latlong <- str_replace(name_aux, "SPEI_", "")

  mylat <- as.numeric(str_split_fixed(latlong, "_", 2))[1]
  mylong <- as.numeric(str_split_fixed(latlong, "_", 2))[2]

  aux$lat <- mylat
  aux$long <- mylong

  # Split time 
  # Get year 
  aux$year <- as.numeric(str_extract(aux$DATA, "[0-9]+"))

  # Get month 
  aux$month <- str_extract(aux$DATA, "[aA-zZ]+")

  # Variable month in number 
  aux$months <- match(aux$month, month.abb)

  # assign(name_aux, aux)

  mydf <- rbind(mydf, aux)
}
```

Select data for SPEI_1, SPEI_3, SPEI_6, SPEI_12, SPEI_24, and SPEI_48 and compute the mean of the 6 cells (0.5º size)

```{r}
spei <- mydf %>% 
  select(SPEI_1,SPEI_3,SPEI_6,SPEI_12,SPEI_24, SPEI_48, lat, long, year, months)

# Get mean values of the all cells (4 celss ) 
# and create a date format variable 
spei_mean <- spei %>% 
  select(-lat, -long) %>% 
  group_by(year,months) %>% 
  summarise(spei1 = mean(SPEI_1),
            spei3 = mean(SPEI_3),
            spei6 = mean(SPEI_6),
            spei12 = mean(SPEI_12),
            spei24 = mean(SPEI_24),
            spei48 = mean(SPEI_48)) %>%
  mutate(date = lubridate::ymd(paste(year, months, "01")))

# Melting data and compute a sing auxiliar column to colour bars 
spei_mean_melt <- melt(spei_mean, id.vars = c("year", "months", "date"))
spei_mean_melt$sign <- ifelse(spei_mean_melt$value > 0, "pos", "neg")

spei_mean_melt2000 <- spei_mean_melt %>% 
  filter(year >= 2000)
```

### Plot 
```{r}
label_variables <- c('spei1' = '1-month',
                     'spei3' = '3-month',
                     'spei6' = '6-month',
                     'spei12' = '12-month',
                     'spei24' = '24-month',
                     'spei48' = '48-month')

d <- ggplot(spei_mean_melt, aes(x=date, y=value, fill=sign)) +  
  geom_bar(stat = "identity") + 
  facet_wrap(~variable, ncol=1, labeller = as_labeller(label_variables)) + 
  scale_fill_manual(values = c("pos" = "darkblue", "neg" = "red")) +
  theme_bw()  + 
  labs(title= 'Drought evolution for Sierra Nevada',
    x= '',
    y='SPEI') +
  theme(strip.background = element_rect(fill = "white"),
        legend.position = "none")

d

# >2000
d2000 <- ggplot(spei_mean_melt2000, aes(x=date, y=value, fill=sign)) +  
  geom_bar(stat = "identity") + 
  facet_wrap(~variable, ncol=1, labeller = as_labeller(label_variables)) + 
  scale_fill_manual(values = c("pos" = "darkblue", "neg" = "red")) +
  theme_bw()  + 
  labs(title= 'Drought evolution for Sierra Nevada (From year 2000)',
    x= '',
    y='SPEI') +
  theme(strip.background = element_rect(fill = "white"),
        legend.position = "none")

d2000
```

# Explore 3, 6 months scales
```{r}
spei_mean_melt1970 <- spei_mean_melt %>% 
  filter(year >= 1970) %>%
  filter(variable %in% c('spei3', 'spei6'))

d1970 <- ggplot(spei_mean_melt1970, aes(x=date, y=value, fill=sign)) +  
  geom_bar(stat = "identity") + 
  facet_wrap(~variable, ncol=1, labeller = as_labeller(label_variables)) + 
  scale_fill_manual(values = c("pos" = "darkblue", "neg" = "red")) +
  theme_bw()  + 
  labs(title= 'Drought evolution for Sierra Nevada (From year 1995)',
    x= '',
    y='SPEI') +
  theme(strip.background = element_rect(fill = "white"),
        legend.position = "none") +
  scale_x_date(date_breaks = "2 year", date_labels = "%Y")

d1970
```

```{r}
### Export all pdfs
di_export <- paste0(di, "/out/fig/drought")

pdf(file=paste0(di_export, "/spei_sn_all.pdf"), height = 10, width =9)
d 
dev.off()

pdf(file=paste0(di_export, "/spei_sn_all_2000.pdf"), height = 10, width =9)
d2000
dev.off()

pdf(file=paste0(di_export, "/spei_sn_all_1970.pdf"), height = 5, width =9)
d1970
dev.off()
```