### 2017-11-08 man/notes_drought.Rmd
update drought notes

---
title: "notes Drought"
author: "AJ Perez-Luque (@ajpelu)"
date: "2016 Aug"
output:  
  md_document:
    toc: true
    variant: markdown_github
bibliography: ../references.bib
csl: ../ecology.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Drought Characterization

* Three most severe drought events (**1950–2012**) [@Trigo2013, @Gouveia2015]:
    * **1998-1999**
    * **2004-2005**
    * **2011-2012**

* The two drought episodes of 2004/2005 and 2011/2012 were extreme in both its magnitude and spatial extent [@Gouveia2014]

<center>![](/out/images/trigo_2012.jpg)</center>

<br>

Figure. Accumulated monthly precipitation averaged over the IP during the hydrological year 2011–12 (black line) and during the other two most severe drought events in the period 1950–2012 (colored lines). Light, dark, and medium gray shaded areas show the ensemble spread from the corresponding evolutions of all hydrological years between 2002–03 and 2011–12, 1950–51 and 1959–60, and the overlapping between them, respectively. The gray thick line indicates the 1950 –2010 mean evolution, with boxes and whiskers representing the ±0.5 sigma level and the 10th–90th percentile interval, respectively.

## Summer of 2005

* 2004/2005 hydrological year is considered one of the worst drooughts ever recorded in the Iberian Peninsula, particularly in the central and southern sectors [@GarciaHerrera2007].

* The southern half of Iberia received less than 45 % of the usual precipitation between October 2004 and June 2005 [@GarciaHerrera2007].

* The hydrological year from October 2004 to September 2005 was the driest on record at several locations throughout Iberia [@GarciaHerrera2007].

## Summer of 2012

<center>![](/man/images/trigo_2012.jpg)</center>

## Long-term perspective

* @GarciaHerrera2007 analyzed three longest dialy series available (1941-2005). 
The following figure shows the daily accumulated precipitation anomaly compared to the 1961–90 normal for Granada station. The colored lines display the six greater droughts ever recorded in each station (1941–2005 period) during the corresponding October (year - 1) to June (year) hydrological year.

<center>![](/man/images/garcia_herrera_granada_zoom.jpg)</center>

* From a long-term perspective, the decades of 40s and the early 50s along with the 80s and 90s are described as particularly dry throughout the whole Iberian territory (Vicente-Serrano 2006a,b)

* An analysis of the long term series from meteorological stations (n=54) of Iberian Peninsula (1961-2011) reveals that major drought episodes in the Iberian Peninsula were recorded in 1981, 1995, 2000 and 2005 [@VicenteSerrano2014].

# Drought characterization

Althought drought events are characteristic of the Mediterranean area (`TODO`: `CITA`), an increase in the drought severity in the Iberian Peninsula has been observed in the last decades [@VicenteSerrano2014]. The 2004/2005 and 2011/2012 hydrological years are considered two of the worst drought periods recorded in the Iberian Peninsula, particularly in the southeren sector [@GarciaHerrera2007 @Gouveia2015 @Trigo2013]. These events were extreme in both its magnitude and spatial extent [@Gouveia2014].

Althought several works have reported these two years as some of the worst drought events, we characterised the drought at several spatio-temporal scales in the study area. From a long-term perspective, we compare the accumulated monthly precipitation at a meteorological station (Granada, Base Aérea) during the hydrological years 2004-2005 and 2011-2012 with the average of accumulated monthly precipitation for the period 1950-2015.

Figure. Accumulated monthly precipitation averaged over the IP during the hydrological year 2011–12 (black line) and during the other two most severe drought events in the period 1950–2012 (colored lines). Light, dark, and medium gray shaded areas show the ensemble spread from the corresponding evolutions of all hydrological years between 2002–03 and 2011–12, 1950–51 and 1959–60, and the overlapping between them, respectively. The gray thick line indicates the 1950 –2010 mean evolution, with boxes and whiskers representing the ±0.5 sigma level and the 10th–90th percentile interval, respectively.

# References