### 2017-12-01 analysis/plots_correlations_resiliences.Rmd
add plots of resiliences

---
title: "Resilience Correlations beetwen drought events"
author: "AJ Perez-Luque (@ajpelu)"
date: "2017"
output: 
  md_document:
    variant: markdown_github
bibliography: ../references.bib 
csl: ../ecology.csl
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, error=TRUE, message=FALSE)
```

```{r, warning=FALSE, message=FALSE}
library("tidyverse")
library("pander")
library('gridExtra')
```

```{r, echo=FALSE}
machine <- 'ajpelu'
#machine <- 'ajpeluLap'
di <- paste0('/Users/', machine, '/Dropbox/phd/phd_repos/qpyr_dendro/', sep='')
```

# Read data from EVI

```{r}
r.bai <- read.csv(file=paste0(di, '/data/resilience/resilience_bai.csv'), header=T, sep = ",")
r.evi <- read.csv(file=paste0(di, '/data/resilience/resilience_evi.csv'), header=T, sep = ",")

# add data of pop
anomalias <- read.csv(file=paste(di, "/data/anomalies/anomalias_evimean.csv", sep=""), header = TRUE, sep = ',')

attr_iv_malla_modis_id <- anomalias %>% dplyr::select(iv_malla_modi_id,long,lat,pop) %>% unique()
r.evi <- r.evi %>% inner_join(attr_iv_malla_modis_id, by='iv_malla_modi_id')

r.evi <- r.evi %>% 
  mutate(
    clu_pop = as.factor(case_when(
      pop == 1 ~ "Camarate",
      pop %in% c(2,3,4,5) ~ 'Northern slope',
      pop %in% c(6,7,8) ~ 'Southern slope',
      pop == 9 ~ 'out')),
    clu_pop2 = as.factor(case_when(
      pop %in% c(1,2,3,4,5) ~ 'N',
      pop %in% c(6,7,8) ~ 'S',
      pop == 9 ~ 'out'))) %>% 
  filter(clu_pop != 'out')

# Change name of clu_pop2 and disturb_year 
r.evi <- r.evi %>% dplyr::rename(site = clu_pop2) %>% 
  mutate(disturb_year = as.factor(disturb_year))

```

# BAI

```{r}
df2005 <- r.bai %>% 
  filter(disturb_year == 2005) %>% 
  rename(rs.2005 = rs, 
         rc.2005 = rc,
         rt.2005 = rt, 
         rrs.2005 = rrs) %>% 
  dplyr::select(-disturb_year)

df2012 <- r.bai %>% 
  filter(disturb_year == 2012) %>% 
  rename(rs.2012 = rs, 
         rc.2012 = rc,
         rt.2012 = rt, 
         rrs.2012 = rrs) %>% 
  dplyr::select(-disturb_year) %>% 
  dplyr::select(-site)

d <- inner_join(df2005, df2012, by = 'tree')
```

# Plots BAI

```{r bai_rs}
# Resilience
bai_rs <- d %>% ggplot(aes(x=rs.2005, y=rs.2012, fill=site)) + 
  stat_ellipse(aes(fill=site), geom = 'polygon', alpha=1/6, type='norm', level=.95) +
  geom_point(aes(colour=site)) + 
  xlab('Resilience to 2005 drouht event') + 
  ylab('Resilience to 2012 drouht event') + 
  xlim(0,1.4) + ylim(0.4,1.8) + 
  theme_bw()+
  geom_hline(yintercept = 1) +
  geom_vline(xintercept = 1) 
```

```{r bai_rt}
# Resistance
bai_rt <- d %>% ggplot(aes(x=rt.2005, y=rt.2012, fill=site)) + 
  stat_ellipse(aes(fill=site), geom = 'polygon', alpha=1/6, type='norm', level=.95) +
  geom_point(aes(colour=site)) + 
  xlab('Resistance to 2005 drouht event') + 
  ylab('Resistance to 2012 drouht event') + 
  xlim(0,1.3) + ylim(0,1.3) + 
  theme_bw()+
  geom_hline(yintercept = 1) +
  geom_vline(xintercept = 1)

```

```{r bai_rc}
# Recovery
bai_rc <- d %>% ggplot(aes(x=rc.2005, y=rc.2012, fill=site)) + 
  stat_ellipse(aes(fill=site), geom = 'polygon', alpha=1/6, type='norm', level=.95) +
  geom_point(aes(colour=site)) + 
  xlab('Recovery to 2005 drouht event') + 
  ylab('Recovery to 2012 drouht event') + 
  theme_bw()+
  geom_hline(yintercept = 1) +
  geom_vline(xintercept = 1) 
```

# EVI
```{r}
e2005 <- r.evi %>% 
  filter(disturb_year == 2005) %>% 
  rename(rs.2005 = rs, 
         rc.2005 = rc,
         rt.2005 = rt, 
         rrs.2005 = rrs) %>% 
  dplyr::select(-disturb_year)

e2012 <- r.evi %>% 
  filter(disturb_year == 2012) %>% 
  rename(rs.2012 = rs, 
         rc.2012 = rc,
         rt.2012 = rt, 
         rrs.2012 = rrs) %>% 
  dplyr::select(-disturb_year) %>% 
  dplyr::select(-c(long:site))

evi <- inner_join(e2005, e2012, by = 'iv_malla_modi_id')
```

# Resilience
```{r evi_rs}
evi_rs <- evi %>% ggplot(aes(x=rs.2005, y=rs.2012, fill=site)) + 
  stat_ellipse(aes(fill=site), geom = 'polygon', alpha=1/6, type='norm', level=.95) +
  geom_point(aes(colour=site)) + 
  xlab('Resilience to 2005 drouht event') + 
  ylab('Resilience to 2012 drouht event') + 
  theme_bw()+
  geom_hline(yintercept = 1) +
  geom_vline(xintercept = 1) 
```

```{r}
# Resistance
evi_rt <- evi %>% ggplot(aes(x=rt.2005, y=rt.2012, fill=site)) + 
  stat_ellipse(aes(fill=site), geom = 'polygon', alpha=1/6, type='norm', level=.95) +
  geom_point(aes(colour=site)) + 
  xlab('Resistance to 2005 drouht event') + 
  ylab('Resistance to 2012 drouht event') + 
  theme_bw()+
  geom_hline(yintercept = 1) +
  geom_vline(xintercept = 1) 
```

```{r}
# Recovery
evi_rc <- evi %>% ggplot(aes(x=rc.2005, y=rc.2012, fill=site)) + 
  stat_ellipse(aes(fill=site), geom = 'polygon', alpha=1/6, type='norm', level=.95) +
  geom_point(aes(colour=site)) + 
  xlab('Recovery to 2005 drouht event') + 
  ylab('Recovery to 2012 drouht event') + 
  theme_bw()+
  geom_hline(yintercept = 1) +
  geom_vline(xintercept = 1) 
```

```{r plot_resiliences_evi_ci, fig.width = 12, fig.height=12}
grid.arrange(bai_rc, bai_rt, bai_rs, ncol=3)
```

```{r, echo=FALSE}
pdf(paste0(di, '/out/mean_resilience/evi/interaction_plotsSE.pdf'), width=9, height = 9)
grid.arrange(plot_mdSE, plot_msSE, plot_mdsSE, ncol=3)
dev.off()

grid.arrange(plot_mdCI, plot_msCI, plot_mdsCI, ncol=3)