### 2017-12-12 man/ms_tables.Rmd
add tables

---
title: "ms_tables"
output:  
  md_document:
    toc: true
    variant: markdown_github
bibliography: ../references.bib
csl: ../ecology.csl
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, echo=FALSE}
library(tidyverse)
library(here)
library(pander)
library(kableExtra)
library(knitr)
library(dplyr)
```

### Table R4

```{r, echo=FALSE}
df <- read.csv(here("/out/anovas_resilience/huber_evi", "trimmed_anovas.csv"), header = TRUE)
names(df) <- c('factores', 'F', 'p', 'variables')

dff <- df %>% gather(temp, numeros, F, p) %>% 
  unite(temp1, variables, temp) %>% 
  spread(temp1, numeros) %>% 
  mutate(factores = recode(factores, 
                           "disturb_year" = "Disturb", 
                           "disturb_year:site" = "Disturb X Site", 
                           "site" = "Site")) %>% 
  arrange(match(factores, c("Disturb", "Site", "Disturb X Site")))

# knitr::kable(dff, 
#       col.names = c("", "F", "p", "F", "p", "F", "p", "F", "p"),
#       digits=c(0,2,3,2,3,2,3,2,3))

colnames(dff) <- c("", "rc\nF", "rc\np", "rrs\nF", "rc\np", "rs\nF", "rc\np", "rt\nF", "rc\np")
pander(dff,
             round=c(0,2,3,2,3,2,3,2,3),
             digits=c(0,5,4,5,4,5,4,5,4),
             keep.line.breaks = TRUE
             )
dff

```