### 2017-12-27 analysis/prepare_eobs_data.Rmd
add eobs data

---
title: "EOBS Climate data"
author: "AJ Perez-Luque (@ajpelu)"
date: "2017 Dec"
output:
  md_document:
    variant: markdown_github
bibliography: ../references.bib 
csl: ../ecology.csl
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, warning=FALSE, error=TRUE, message=FALSE)
```

```{r packages, warning=FALSE, message=FALSE, echo=FALSE}
library("tidyverse")
```

# Cargar los datos
```{r}
machine <- 'ajpelu'
# machine <- 'ajpeluLap'
di <- paste0('/Users/', machine, '/Dropbox/phd/phd_repos/qpyr_dendro/', sep = '')
```

```{r}
eobs <- read.csv(file=paste0(di, 'data_raw/eobs/eobs_data_robles.csv'), header=TRUE, sep=',')

eobs <- eobs %>% 
  mutate(date = as.Date(str_sub(startDate, 1, 11), format="%Y-%m-%d"), 
         year = lubridate::year(date), 
         month = lubridate::month(date))

eobs %>% filter(variable == 'rr') %>% ggplot(aes(x=date, y=value, colour=loc)) + geom_line()

```