### 2017-12-28 analysis/rwi/climate_rwi_relate.Rmd
update climate eobs relationships

---
title: "Climate-RWI relations"
author: "AJ Perez-Luque (@ajpelu)"
date: "2017 Dec"
output:
  md_document:
    variant: markdown_github
bibliography: ../references.bib 
csl: ../ecology.csl
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, warning=FALSE, error=TRUE, message=FALSE)
```

```{r packages, warning=FALSE, message=FALSE, echo=FALSE}
library("tidyverse")
library("stringr")
library("treeclim")
```

```{r}
machine <- 'ajpelu'
# machine <- 'ajpeluLap'
di <- paste0('/Users/', machine, '/Dropbox/phd/phd_repos/qpyr_dendro/', sep = '')
```

# Read and prepare data 
```{r}
eobs <- read.csv(file=paste0(di, "data/eobs/eobs_formatted.csv"), header=TRUE, sep=',')
crono <- read.csv(file=paste(di, "data/cronos_medias/cronos_sites_rwi.csv", sep=""), header=TRUE, sep=',')
```

* Utilizamos la crono residual no la estandar :red_circle: DUDA

```{r}
cro_sj <- crono %>% filter(site == 'sj') %>% dplyr::select(-site, -std) %>% column_to_rownames(var = "year") %>% na.omit()
cro_caL <- crono %>% filter(site == 'caL') %>% dplyr::select(-site, -std) %>% column_to_rownames(var = "year") %>% na.omit()
cro_caH <- crono %>% filter(site == 'caH') %>% dplyr::select(-site, -std) %>% column_to_rownames(var = "year") %>% na.omit()
```

# Create subsets data for eobs

```{r}
eobsN <- eobs %>% filter(year < 2017) %>% filter(loc == 'N')

prec_N <- eobsN %>% dplyr::select(year, month, prec) 
tmean_N <- eobsN %>% dplyr::select(year, month, tmean) 
tmin_N <- eobsN %>% dplyr::select(year, month, tmin) 
tmax_N <- eobsN %>% dplyr::select(year, month, tmax)

eobsS <- eobs %>% filter(year < 2017) %>% filter(loc == 'S')

prec_S <- eobsS %>% dplyr::select(year, month, prec) 
tmean_S <- eobsS %>% dplyr::select(year, month, tmean) 
tmin_S <- eobsS %>% dplyr::select(year, month, tmin) 
tmax_S <- eobsS %>% dplyr::select(year, month, tmax) 
```

# Precipitation
from previous June to

```{r}
rprec_SJ <- dcc(chrono = cro_sj, climate = prec_N, 
         method = "correlation",  boot = "std",
        .range(-3:9) + .sum(-9:8))

rprec_caL <- dcc(chrono = cro_caL, climate = prec_S, 
         method = "correlation",  boot = "std",
        .range(-3:9) + .sum(-9:8))

rprec_caH <- dcc(chrono = cro_caH, climate = prec_S, 
         method = "correlation",  boot = "std",
        .range(-3:9) + .sum(-9:8))

rprecs <- c('rprec_SJ', 'rprec_caH', 'rprec_caL')
rprec_all <- c()

for (i in rprecs){

  # Get object dcc
  aux <- get(i)

  # Get coeffs 
  aux_coef <- aux$coef

  # remove rownames
  row.names(aux_coef) <- NULL

  # Change name for aggregate variables 
  aux_coef <- aux_coef %>% 
    mutate(month = case_when(
      month == "sep...AUG" ~ "Hydrol", 
      TRUE ~ as.character(.$month)),
      sig = case_when(
      significant == "TRUE" ~ "*", 
      significant == "FALSE" ~ ""))

  # Add loc variable
  name_site <- str_sub(i, start=7)
  aux_coef$site <- name_site

  # rbind
  rprec_all <- rbind(rprec_all, aux_coef)
  } 
```

```{r}

rprec_all

rprec_all %>% ggplot(aes(x=id, y=coef, group = site)) +
  geom_bar(aes(fill=site),
           stat="identity", position = "dodge") +
  scale_x_continuous(breaks = rprec_all$id, labels = rprec_all$month) +
  theme_bw() + 
  theme(axis.title.x = element_blank()) +
  geom_text(aes(label = sig), 
            position = position_dodge(0.9), 
            vjust = 0.1, 
            size = 10) +
  ylab('Correlation')

```{r}

eobs_norht_no_prec <- eobs_norht %>% dplyr::select(-prec)

mc <- dcc(chrono = cro_sj, 
          climate = eobs_norht, 
          var_names = c("prec", "tmax", "tmean", "tmin"), 
          method = "correlation", boot = "stationary")

sea_sj <- seascorr(chrono = cro_sj, 
             climate = eobs_norht, 
             var_names = c("prec", "tmax", "tmean", "tmin"))

plot(sea_sj)

dc_sj

croN <- cro %>% filter(site == 'sj') %>% dplyr::select(-site) %>% 
  filter(year > 1949)
rownames(croN) <- croN$year

cl <- eobs %>% 
  filter(loc == 'N') %>% 
  filter(variable %in% c('rr', 'tg')) %>% 
  dplyr::select(value, variable, year, month) %>% 
  spread(variable, value)

cl <- cl %>% rename(prec = rr, temp = tg) %>% 
  filter(year < 2017)

test <- dcc(chrono = croN, climate = cl)

croN 
# http://czang.org/treeclim.html

```

The relationships between monthly climatic variables (mean maximum and minimum temperatures and total precipitation) and the chronologies of the ring-width indices were calculated using bootstrapped correlation functions in the treeclim package in R (Zang & Biondi, 2015).