### 2018-01-10 analysis/disturbance/explore_disturbance_chronos.Rmd

---
title: "disturbance chronologies"
author: "AJ Perez-Luque (@ajpelu)"
date: "2017 April"
output:  
  md_document:
    variant: markdown_github
bibliography: references.bib 
csl: ecology.csl
---

```{r, warning=FALSE, message=FALSE}
library("tidyverse")
library("dplR")
library("knitr")
library("rowr")
library("zoo")
```

# Read data

```{r, echo=FALSE}
machine <- 'ajpelu'
# machine <- 'ajpeluLap'
di <- paste('/Users/', machine, '/Dropbox/phd/phd_repos/qpyr_dendro/', sep='')

# sj 
sj <- read.rwl(fname=paste0(di, '/data_raw/dendro_ring/sn_sanjuan/sn_sanjuan.rwl'), format="tucson")

# canar 
ca <- read.rwl(fname=paste0(di, '/data_raw/dendro_ring/sn_canar/sn_canar.rwl'), format="tucson")
```

# Disturbance chronologies 
* Vamos a computar las disturbances chronologies 
...

```{r}
# Get summary ring-width series
sj_summ <- summary(sj)

# Get names cores with last year different to 2016 
id_cores_no_bark <- sj_summ %>% 
  filter(last != 2016) %>% select(series) %>% mutate(series = factor(series))

id_cores_no_bark <- unname(unlist(id_cores_no_bark))

# Subbet datasets 
sj_cor <- sj[which(! colnames(sj) %in% (id_cores_no_bark))]
sj_sin <- sj[which(colnames(sj) %in% (id_cores_no_bark))]
```

```{r}
rws <- data.frame(year = as.numeric(as.character(rownames(sj_cor))),
                  rw =sj_cor[,'SNA1102'])

# Some notes about lag, windows, etc. 
# https://cran.r-project.org/web/packages/dplyr/vignettes/window-functions.html

rws <- rws %>% 
  mutate(g2= rollapply(rw, width = 10, sum, align = 'left', partial=TRUE), 
         g1_nolag = rollapply(rw, width = 11, sum, align = 'right', partial=TRUE),
         rw_lag = lag(rw, 1),
         g1_previous10 = rollapply(rw_lag, width = 10, sum, align = 'right', partial=TRUE, na.rm=T),
         g1_mio = g1_nolag - rw,
         gc = ((g2 - g1_previous10) / g1_previous10)*100,
         gc_mio = ((g2 - g1_mio) / g1_mio)*100)

rws %>% 
  ggplot(aes(x=year, y=gc)) + geom_bar(stat='identity') +
  geom_hline(yintercept = c(-25,25), linetype = "dashed", colour="red") + 
  theme_bw()
```