### 2018-01-18 man/notes_EVI.Rmd
update ms

---
title: "notes EVI"
author: "AJ Perez-Luque (@ajpelu)"
date: "2017 nov"
output:  
  md_document:
    toc: true
    variant: markdown_github
bibliography: ../refs/references.bib
csl: ../refs/ecology.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Notas sobre EVI

## MODIS Vegetation Indices

***Aim***: To obtain vegetation indices for Quercus pyrenaica forests in Sierra Nevada from 2000 to 2016.

Technical info: 
* MODIS IV:

 * Collection `MODIS 006`
 * Temporal range: 2000 - 2016 
 * Pixel approach

* Distribution of Pyrenan oak forest in Sierra Nevada:

 * Derived from Ecosystems map of Sierra Nevada 
 * Datum: `epsg:4326`
 * See this [link](https://rawgit.com/ajpelu/qpyr_distribution/master/analysis/distribution_map_sn.html)

* Pixels covering Pyrenan oak forest in Sierra Nevada: `./github/ajpelu/qpyr_distribution/data_raw/geoinfo/iv_malla_modis_qp_centroid.shp` 
 * *n* pixels = 927 pixels 
 * See this [link](https://rawgit.com/ajpelu/qpyr_distribution/master/analysis/distribution_map_sn.html)

The script to download MODIS data is at: `./script/GEE/get_iv_modis_qp.js`

# Filtrado de datos

Para el filtrado de datos ver el siguiente link
[`./analysis/prepare_modis_qa.md`](/analysis/prepare_modis_qa.md)

# References