### 2018-01-18 man/report.Rmd

---
title: "ms_borrador"
author: "AJ Perez-Luque; G. Gea-Izquierdo; R. Zamora; FJ. Bonet"
date: "2017"
output:
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 7.5
    highlight: null
    reference_docx: template_ms.docx
bibliography: ../refs/references.bib
csl: ../refs/ecology.csl
---

```{r child = 'ms_0_intro.Rmd'}
```

```{r child = 'ms_1_material_methods.Rmd'}
```

```{r child = 'ms_2_results.Rmd'}
```

```{r child = 'ms_3_discussion.Rmd'}
```

```{r child = 'ms_tables.Rmd'}
```

```{r child = 'ms_figures.Rmd'}
```