### 2018-01-18 man/ms_supp.Rmd
create supplementary rmd

---
title: "ms_supplementary"
output:  
  md_document:
    toc: true
    variant: markdown_github
bibliography: ../refs/references.bib
csl: ../refs/ecology.csl
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```

```{r}
library("tidyverse")
library("here")
library("dplyr")
library("ggrepel")
library("scales")

library("grid")
library("gridExtra")
library("patchwork")

```

# Supplementary

## Supplementary 1

Listado de pkgs usados

## Suplementary 2

## Figure S2

![](../analysis/evi/compute_MannKendall_evimean_files/figure-markdown_github/rater_map_tau_EVI-1.png){width=150%}

Figure S2. Spatial distribution of the EVI-annual trends for the period 2000-2016. Trends ($\tau$) with the most positive values (green) indicate positive trend.

## Figure S3

```{r}
df <- read.csv(file='../out/anomalies/evi/percen_browing_greening.csv', header=TRUE)

df <- df %>% add_row(clu_pop2 = 'Northern slope',
                             type = 'greening',
                             count_clu = 0, per = 0.00, y = 2005)

per_pixels_green <- df %>% 
  ggplot(aes(x=type, y=per, fill=as.factor(y))) + 
  geom_bar(stat='identity', position = "dodge") + 
  geom_text(aes(label = paste0(per, " %")), position=position_dodge(width=0.9), vjust= - 0.75) + 
  facet_wrap(~clu_pop2) + 
  theme_bw() + xlab('') + ylab('Frequency of pixels (%)') +
  theme(strip.background = element_rect(colour = "black", fill = "white")
        ) + theme_minimal() + labs(fill = 'Drought event') + scale_fill_manual(values = c("#DAA520", "#104E8B"))

per_pixels_green
```

Figure S3. Percentage of pixels showing browning, greenning or no-changes during the 2005 and 2012 droguht events according to EVI standardized anomalies.