### 2018-01-22 man/ms_tables/t_huber_res_evi.Rmd

---
output:  
  word_document:
    reference_docx: "../../man/template_tables_portrait.docx"
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, echo=FALSE}
library(tidyverse)
library(here)
library(pander)
library(kableExtra)
library(knitr)
library(dplyr)
```

**Table S1.** Estimadores huber resiliences EVI

```{r, echo = FALSE, warning = FALSE}
mraw <- df <- read.csv(here("/out/anovas_resilience/huber_evi", "robust_mhuber_agg.csv"), header = TRUE)

m <- mraw %>% 
  filter(var != 'rrs') %>% 
  rename(variables = var)  %>% 
  dplyr::select(c(-n, -Letter)) %>% 
 unite("mh", c("M.Huber", "ci"), sep = "\n")

# Datos agrupados para todos los sitios 
mraw_a <- read.csv(here("/out/anovas_resilience/huber_evi", "robust_mhuber_agg_a.csv"), header = TRUE)

m_a <- mraw_a %>% 
  filter(var != 'rrs') %>% 
  rename(variables = var)  %>% 
  dplyr::select(c(-n, -Letter)) %>%  
  unite("mh", c("M.Huber", "ci"), sep = "\n") %>% 
  mutate(site = "z_all")

# Datos agrupados para todos los sitios 
mraw_b <- read.csv(here("/out/anovas_resilience/huber_evi", "robust_mhuber_agg_b.csv"), header = TRUE)

m_b <- mraw_b %>% 
  filter(var != 'rrs') %>% 
  rename(variables = var)  %>% 
  dplyr::select(c(-n, -Letter)) %>%  
  unite("mh", c("M.Huber", "ci"), sep = "\n") %>%
  mutate(disturb_year = 2099)

mm <- m %>% 
  bind_rows(m_a) %>%
  bind_rows(m_b) %>% 
  gather(temp, numeros, mh) %>% 
  unite(temp1, c(variables, disturb_year), temp) %>% 
  spread(temp1, numeros, fill = "")

mm <- mm %>% 
  dplyr::select(site, 
                rt_2005_mh, rc_2005_mh, rs_2005_mh,
                rt_2012_mh, rc_2012_mh, rs_2012_mh,
                rt_2099_mh, rc_2099_mh, rs_2099_mh)

colnames(mm) <- c(" \nsite", "2005\nRt", "2005\nRc", "2005\nRs", "2012\nRt", "2012\nRc", "2012\nRs", "Rt", "Rc", "Rs")
pander(mm, keep.line.breaks = TRUE, split.table = Inf)
```