### 2018-01-23 analysis/analysis_KW_stand_features.Rmd
include KW

---
title: "Comparison ANOVAS stand features"
author: "AJ Perez-Luque (@ajpelu)"
date: "2018 Jan"
output:  
  md_document:
    variant: markdown_github
bibliography: references.bib 
csl: ecology.csl
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library("tidyverse")
library("raster")
library("stringr")
library('rgdal')
library("sp")
library("knitr")
library("broom")
library("multcomp")
library("pander")
library("devtools")
source_gist("https://gist.github.com/ajpelu/194e721077ec045a2b864088908e7aca", filename = 'table_glht.r') # multicomp table
```

# Prepare Data
```{r, echo=FALSE}

machine <- 'ajpelu'
# machine <- 'ajpeluLap'
di <- paste0('/Users/', machine, '/Dropbox/phd/phd_repos/qpyr_dendro/', sep='')

# Read tree data
tree <- read.csv(file=paste0(di, '/data_raw/dendro_competence.csv'), header=TRUE, sep=',')
```

```{r}
# Compute diameter (mm)
tree <- tree %>% 
  mutate(dn_mm = (perim_mm / pi))

sj_code  <- paste0('A', str_pad(1:20, 2, pad='0'))
ca_lowcode <- c(paste0('B', str_pad(1:10, 2, pad='0')), paste0('B', 26:30))
ca_highcode <- paste0('B', 11:25)

tree <- tree %>% mutate(
  site = as.factor(case_when(
    id_focal %in% sj_code ~ "sj", 
    id_focal %in% ca_highcode ~ "caH", 
    id_focal %in% ca_lowcode ~ "caL")))

# Get only focal trees  
ft <- tree %>% 
  filter(sp=='Focal') %>% 
  filter(id_focal!='Fresno')

# Get only no focal trees 
nft <- tree %>% 
  filter(sp!='Focal')

# Select only variables to compare (ft)
ft_sel <- ft %>%
  mutate(dn = dn_mm / 10, 
         h = height_cm / 100) %>% 
  dplyr::select(dn, h, site)

# Select only variable to compare (all trees)
all <- nft %>% mutate(dn = dn_mm / 10, 
                      h = height_cm / 100) %>% 
  dplyr::select(dn, h, site) %>% 
  bind_rows(ft_sel)

```

# Competition data

* Read data 
```{r, echo=FALSE}
# Read competition
compe <- read.csv(file=paste(di, "/data/competence/competence_indices.csv", sep=""), header=TRUE, sep=',')

compe <- compe %>% 
  mutate(site = as.factor(case_when(
    id_focal %in% sj_code ~ "sj", 
    id_focal %in% ca_highcode ~ "caH", 
    id_focal %in% ca_lowcode ~ "caL")))

compe_sel <- compe %>% dplyr::select(ba, std, site)
```

# Kruskal-Wallis comparison

```{r}
comparaKW <- function(df, mivariable){ 
  require(PMCMR)
  require(multcompView)

  output <- list()

  # Model formulation
  myformula <- as.formula(paste0(mivariable, " ~ site"))

  # Kruskal Wallis  
  kt <- kruskal.test(myformula, data = df)

  # Summary AOV (broom style)
  tm <- broom::tidy(kt)
  tm$mi_variable <- mivariable

  # Dunn's test 
  object <- posthoc.kruskal.dunn.test(myformula, data=df, p.adjust.method = "bonferroni")

  # Get dataframe with letters and pvalues (#from summary.PMCMR)
  pval <- as.numeric(object$p.value)
  stat <- as.numeric(object$statistic)
  grp1 <- as.numeric(c(col(object$p.value)))
  cnam <- colnames(object$p.value)
  grp2 <- as.numeric(c(row(object$p.value)))
  rnam <- rownames(object$p.value)
  H0 <- paste(cnam[grp1], " = ", rnam[grp2])
  OK <- !is.na(pval)
  xdf <- data.frame(H0 = H0[OK], statistic = stat[OK], p.value = format.pval(pval[OK], 5))

  # Get letters (using multcompView) See viggnete PMCMR 
  dt_letters  <- multcompView::multcompLetters(get.pvalues(object), threshold = 0.05)

  output$KW <- tm
  output$post_hoc <- xdf
  output$letters <- dt_letters
  return(output)
  }
```

```{r, warning=FALSE}
# Focal Tree 
kw_ft_dn <- comparaKW(ft_sel, 'dn')
kw_ft_h <- comparaKW(ft_sel, 'h')

# All trees 
kw_all_dn <- comparaKW(all, 'dn') 
kw_all_h <- comparaKW(all, 'h')

# Competence 
kw_ba <- comparaKW(compe_sel, 'ba')
kw_std <- comparaKW(compe_sel, 'std')
```

# Export data to text files (see `/out/kw_stand_features/`)

```{r}

variables <- c('ft_dn', 'ft_h', 'all_dn', 'all_h', 'ba', 'std')

for (i in variables){

  out <- get(paste0('kw_', i))
  file_out <- file(paste0(di,'out/kw_stand_features/kw_', i, '.txt'), "w")
  sink(file_out)
  cat(paste0("VARIABLE = ",i," \n"))
  cat("\n")
  cat("KRUSKAL WALLIS \n")
  print(out$KW) 
  cat("\n")

  cat("POST HOC \n")
  print(out$post_hoc) 
  cat("\n")

  cat("Letters \n")
  print(out$letters)
  cat("\n")

  while (sink.number()>0) sink()
  # close(file_out)
}

while (sink.number()>0) sink()
```

```{r}

variables <- c('ft_dn', 'ft_h', 'all_dn', 'all_h', 'ba', 'std') 
output_letras <- c()

for (i in variables){

  out <- get(paste0('kw_', i))
  l <- as.data.frame(out$letters$Letters)
  l$site <- row.names(l)
  colnames(l) <- c('letter', 'site')
  l$mivariable <- i

  output_letras <- rbind(output_letras, l)
}

# Export data 
write.csv(output_letras, file=paste(di, "out/kw_stand_features/kw_letras.csv", sep=""), row.names = FALSE)
```