### 2018-02-13 analysis/drought/explore_prec_longterm.Rmd
compute 2sd precipitation

---
title: "LongTerm Hydrological Prec"
author: "AJ Perez-Luque (@ajpelu)"
output: 
  md_document:
    variant: markdown_github
bibliography: ../references.bib 
csl: ../ecology.csl
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, warning=FALSE, error=TRUE, message=FALSE)
```

```{r packages, warning=FALSE, message=FALSE, echo=FALSE}
library("here")
library("tidyverse")
library("lubridate")
library("ggplot2")
library("ggrepel")
```

```{r}
source(here::here('script/R', 'hydro_year.R'))
```

```{r}
ltdf <- read.csv(file=here::here('data_raw/meteo/', 'long_term_data_rediam_data.txt'), header=T, sep=';', dec=',')

ltdf <- ltdf %>% 
  dplyr::select(cod = INDICATIVO, VARIABLE, FECHA, VALOR) %>% 
  dplyr::mutate(fecha = as.Date(FECHA, format='%d/%m/%Y'),
         month = lubridate::month(fecha),
         hmonth = ifelse(month <= 8, month + 4, month -8), 
         hyear = hydro_year(fecha),
         hyear_f = paste0(hyear-1,'-', hyear))

st5514 <- ltdf %>% 
  filter(VARIABLE == 'PM1') %>% 
  filter(cod == '5514') %>% 
  filter(hyear >= 1950) %>% 
  dplyr::select(fecha, month, hmonth, hyear, hyear_f, pre=VALOR)

df <- st5514 %>% 
  group_by(hyear_f) %>% 
  dplyr::mutate(csum = cumsum(pre)) %>% 
  filter(hmonth == 12) %>% 
  dplyr::select(-month, -pre, -fecha) %>% as.data.frame()

meandf <- mean(df$csum)
sddf <- sd(df$csum)

prec_hydrol <- ggplot(df, aes(x=hyear, y=csum)) + 
  geom_point() + theme_bw() + 
  geom_hline(yintercept = meandf) + 
  geom_hline(yintercept = meandf + 2*sddf, color ='blue', linetype = 'dashed') +
  geom_hline(yintercept = meandf - 2*sddf, color ='red', linetype = 'dashed') +
    geom_hline(yintercept = meandf + 1*sddf, color ='blue', linetype = 'dotted') +
  geom_hline(yintercept = meandf  - 1*sddf, color ='red', linetype = 'dotted') +
  scale_x_continuous(breaks=seq(1950,2020, by=5)) +
  geom_text_repel(
    data = subset(df, csum < (meandf - 1*sddf)),
    aes(label = hyear)) +
  ylab('Cumulative prec (hydrological year)')

pdf(here::here('/out/precipitation_2sd', 'prec_hydrol.pdf'), width=12, height = 9)
prec_hydrol
dev.off()

df2sd <- df %>% filter(csum < (meandf - 2*sddf)) %>% as.data.frame()
pander(df2sd)
write.csv(df2sd, here::here("/data/sequias", "prec19502012_2sd.csv"), row.names = FALSE)

df1sd <- df %>% filter(csum < (meandf - 1*sddf)) %>% as.data.frame()
pander(df1sd)
write.csv(df1sd, here::here("/data/sequias", "prec19502012_1sd.csv"), row.names = FALSE)

```