### 2018-02-15 analysis/drought/explore_prec_longterm.Rmd
update

---
title: "LongTerm Hydrological Prec"
author: "AJ Perez-Luque (@ajpelu)"
output: 
  md_document:
    variant: markdown_github
bibliography: ../references.bib 
csl: ../ecology.csl
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, warning=FALSE, error=TRUE, message=FALSE)
```

```{r packages, warning=FALSE, message=FALSE, echo=FALSE}
library("here")
library("tidyverse")
library("lubridate")
library("ggplot2")
library("ggrepel")
```

```{r}
source(here::here('script/R', 'hydro_year.R'))
```

```{r}
ltdf <- read.csv(file=here::here('data_raw/meteo/', 'long_term_data_rediam_data.txt'), header=T, sep=';', dec=',')

ltdf <- ltdf %>% 
  dplyr::select(cod = INDICATIVO, VARIABLE, FECHA, VALOR) %>% 
  dplyr::mutate(fecha = as.Date(FECHA, format='%d/%m/%Y'),
         month = lubridate::month(fecha),
         hmonth = ifelse(month <= 8, month + 4, month -8), 
         hyear = hydro_year(fecha),
         hyear_f = paste0(hyear-1,'-', hyear))

st5514 <- ltdf %>% 
  filter(VARIABLE == 'PM1') %>% 
  filter(cod == '5514') %>% 
  filter(hyear >= 1950) %>% 
  dplyr::select(fecha, month, hmonth, hyear, hyear_f, pre=VALOR)

df <- st5514 %>% 
  group_by(hyear_f) %>% 
  dplyr::mutate(csum = cumsum(pre)) %>% 
  filter(hmonth == 12) %>% 
  dplyr::select(-month, -pre, -fecha) %>% as.data.frame()

meandf <- mean(df$csum)
sddf <- sd(df$csum)

prec_hydrol <- ggplot(df, aes(x=hyear, y=csum)) + 
  geom_point() + theme_bw() + 
  geom_hline(yintercept = meandf) + 
  geom_hline(yintercept = meandf + 2*sddf, color ='blue', linetype = 'dashed') +
  geom_hline(yintercept = meandf - 2*sddf, color ='red', linetype = 'dashed') +
    geom_hline(yintercept = meandf + 1*sddf, color ='blue', linetype = 'dotted') +
  geom_hline(yintercept = meandf  - 1*sddf, color ='red', linetype = 'dotted') +
  scale_x_continuous(breaks=seq(1950,2020, by=5)) +
  geom_text_repel(
    data = subset(df, csum < (meandf - 1*sddf)),
    aes(label = hyear)) +
  ylab('Cumulative prec (hydrological year)')

pdf(here::here('/out/precipitation_2sd', 'prec_hydrol.pdf'), width=12, height = 9)
prec_hydrol
dev.off()

df2sd <- df %>% filter(csum < (meandf - 2*sddf)) %>% as.data.frame()
pander(df2sd)
write.csv(df2sd, here::here("/data/sequias", "prec19502012_2sd.csv"), row.names = FALSE)

df1sd <- df %>% filter(csum < (meandf - 1*sddf)) %>% as.data.frame()
pander(df1sd)
write.csv(df1sd, here::here("/data/sequias", "prec19502012_1sd.csv"), row.names = FALSE)

```

## PREC + 2sd for more series

```{r}

prec_files <- list.files(here::here('data_raw/meteo/prec_subclima/'), pattern="*.csv")

dfpre <- c()

for (i in prec_files){

  auxdf <- read.csv(here::here('data_raw/meteo/prec_subclima', i), header = TRUE, sep=";")
  auxdf$code <- gsub(x=i, pattern = ".csv", replacement = "")

  dfpre <- rbind(dfpre, auxdf)
}

dfpre$date <- as.Date(dfpre$date, format = "%d/%m/%Y")

dfpre <- dfpre %>% 
  dplyr::mutate(fecha = as.Date(date, format='%d/%m/%Y'),
         month = lubridate::month(fecha),
         hmonth = ifelse(month <= 8, month + 4, month -8), 
         hyear = hydro_year(fecha),
         hyear_f = paste0(hyear-1,'-', hyear))

# sql to get lat/lon en Linaria
# SELECT 
#   cli_estaciones.id, 
#   cli_estaciones.nombre, 
#   cli_estaciones.codigo, 
#   cli_estaciones.punto_x, 
#   cli_estaciones.punto_y, 
#   cli_estaciones.geometria, 
#   ST_X(ST_Transform (geometria, 4326)) AS "Longitude",
#   ST_Y(ST_Transform (geometria, 4326)) AS "Latitude"
# FROM 
#   public.cli_estaciones 
# WHERE cli_estaciones.codigo IN ('5501','5502','5504','5504A','5505','5509','5509A','5511E',
# '5517','5518','6224','6234','6234E','6236','6242U','6243','6243E','6248','6251','6252','6253',
# '6257','6257I','6280U','6281E','6282','6307','6307X','6308');

stations  <- read.csv(here::here('data_raw/meteo/prec_subclima/mdestaciones', 'estaciones_latlong.csv'), header=TRUE, sep=";")

stations <- stations %>% dplyr::select(nombre, codigo, Longitude, Latitude)

aux <- dfpre %>% 
  group_by(hyear_f, code) %>% 
  dplyr::mutate(csum = cumsum(prec)) %>% 
  filter(hmonth == 12) %>% 
  dplyr::select(-month, -prec, -date, -fecha) %>% as.data.frame()

presn <- aux %>% group_by(hyear) %>% 
     dplyr::summarise(
     csum_mean = mean(csum, na.rm = TRUE),
     csum_sd = sd(csum, na.rm = TRUE), 
     csum_se = csum_sd/sqrt(length(csum_sd)), 
     n = n())

meandf <- mean(presn$csum_mean)
sddf <- sd(presn$csum_mean)

prec_hydrol_sn <- ggplot(presn, aes(x=hyear, y=csum_mean)) + 
  geom_hline(yintercept = meandf) + 
  geom_hline(yintercept = meandf + 2*sddf, color ='blue', linetype = 'dashed') +
  geom_hline(yintercept = meandf - 2*sddf, color ='red', linetype = 'dashed') +
  geom_hline(yintercept = meandf + 1*sddf, color ='blue', linetype = 'dotted') +
  geom_hline(yintercept = meandf  - 1*sddf, color ='red', linetype = 'dotted') +
  geom_errorbar(aes(ymin = csum_mean - csum_se, ymax= csum_mean + csum_se), width=.1, colour='gray') +
  geom_point(size=2) + theme_bw() +
  scale_x_continuous(breaks=seq(1950,2020, by=5)) +
  scale_y_continuous(breaks=seq(0,1500, by=250)) +
  geom_text_repel(
    data = subset(presn, csum_mean < (meandf - 1*sddf)),
    aes(label = hyear)) +
  theme(panel.grid.minor = element_blank()) +
  ylab('Cumulative Precipitation (hydrological year)') + xlab('year')
prec_hydrol_sn

pdf(here::here('/out/precipitation_2sd', 'prec_hydrol_sn.pdf'), width=12, height = 7)
prec_hydrol_sn
dev.off()

```