### 2018-02-22 analysis/notula/evi_resilience_fine_scale.Rmd
add analysis notula

---
title: "EVI-resilience fine-scale"
author: "AJ Perez-Luque (@ajpelu)"
output: 
  md_document:
    variant: markdown_github
---
```{r, echo=FALSE, message=FALSE}
require(knitr)
opts_chunk$set(fig.align='center', message = FALSE, warning = FALSE) 
```

```{r wd, echo=FALSE}
#---------------------------------
machine <- 'ajpelu'
#machine <- 'ajpeluLap'
di <- paste('/Users/', machine, '/Dropbox/phd/phd_repos/qpyr_dendro', sep='')
#---------------------------------
```

```{r packages, warning=FALSE, message=FALSE, echo=FALSE}
library("here")
library('dplyr')
library("sf")
library("mapview")
library('tidyverse')

library('broom')
library('pander')
library('lsmeans')
library('effects')
library('gridExtra')
library('WRS2')
library('car')
library('MASS')
library('rcompanion')
library('devtools')
# devtools::install_github("ajpelu/auxiliar")
```

# Resilience populations

```{r}
# Read data
raw_evires <- read.csv(file=here::here("/data/resilience", "resilience_evi.csv"), header = TRUE, sep = ',')

# add data of pop
anomalias <- read.csv(file=here::here("/data/anomalies", "anomalias_evimean.csv"), header = TRUE, sep = ',')

attr_iv_malla_modis_id <- anomalias %>% dplyr::select(iv_malla_modi_id,long,lat,pop) %>% unique()

raw_evires <- raw_evires %>% inner_join(attr_iv_malla_modis_id, by='iv_malla_modi_id')

# filter by pop and add new variable

# ¿Que poblaciones estamos evaluando? CA and SJ

evires <- raw_evires %>% 
  filter(pop %in% c(2,6)) %>% 
  mutate(
     clu_pop = as.factor(case_when(
      pop == 2 ~ "SJ",
      pop == 6 ~ "CA")),
     clu_pop2 = as.factor(case_when(
      pop == 2 ~ "N",
      pop == 6 ~ "S"))
     )

# Check spatial 
ev <- st_as_sf(x = evires, coords = c("long","lat"), 
               crs = "+proj=longlat +datum=WGS84")
mapview(ev)

# Change name of clu_pop2 and disturb_year para los analisis anovas
evires <- evires %>% dplyr::rename(site = clu_pop, site2 = clu_pop2) %>% 
  mutate(disturb_year = as.factor(disturb_year))
```

# Fine fine scale 
* Ver para la correspondencia para cada pixel de los árboles de muestreo (ver esto https://rawgit.com/ajpelu/qpyr_historic/master/analysis/historic_map.html)

```{r}
fine <- read.csv(file=here::here("/data_raw", "evi_dendro.csv"), header = TRUE, sep = ',')

# number of tree by pixels 
fineid <- fine %>% group_by(iv_malla_modi_id, site) %>% count() %>% as.data.frame()

pixel_selected <- fineid %>% dplyr::select(iv_malla_modi_id) %>% as.vector()

eviresf <- evires %>% filter(iv_malla_modi_id %in% pixel_selected$iv_malla_modi_id)
```

```{r}
# BAI resilience
baires <- read.csv(here::here("data/resilience", "resilience_bai.csv"), header = TRUE, sep=",")
```

# Prepare Data

```{r}
aux_bai <- baires %>% filter(disturb_year != 1995) %>% 
  mutate(vari = '2_BAI') %>% 
  dplyr::select(-tree) %>% 
  mutate(site = case_when(
    site == "caL" ~ "CA-L",
    site == "caH" ~ "CA-H",
    TRUE ~ as.character(site)))

aux_evi <- evires %>% 
  mutate(vari = '0_EVI_pop') %>% 
  dplyr::select(-iv_malla_modi_id, -long, -lat, -pop, -site2)

aux_evifine <- eviresf %>% 
  mutate(vari = '1_EVI_pix') %>%
  dplyr::select(-iv_malla_modi_id, -long, -lat, -pop, -site2)

df <- rbind(aux_bai, aux_evi, aux_evifine)

dfrs <- df %>% 
  dplyr::select(-rt, -rc, -rrs) %>% 
  group_by(vari, disturb_year, site) %>% 
  summarise(mean = mean(rs), 
                   sd = sd(rs),
                   se = sd/sqrt(length(rs)))

```

# Plot data

```{r}
pd <- position_dodge(width = 0.6)

aes_disturb <- aes(x = disturb_year, y = mean, 
                   color = site, group = site,  
                   fill = site, linetype=site,
                   shape = site, 
                   ymin = mean - se, ymax = mean + se)

variables  <- c("0_EVI_pop" = "EVI population scale", "1_EVI_pix" = "EVI fine scale","2_BAI" = "BAI")

rs <- ggplot(dfrs, aes_disturb) + 
  geom_errorbar(width=.1, position=pd) +
  geom_point(position = pd) + 
  facet_wrap(~vari, labeller = labeller(vari = variables)) + 
  theme_bw() + xlab('') + ylab('Resilience') + 
  theme(
    panel.grid.minor = element_blank(),
    strip.background = element_rect(
      colour = "black",
      fill = "white"),
    strip.text = element_text(face="bold", size = 14),
    axis.title.y = element_text(face="bold", size = 14),
    plot.title = element_text(hjust = -0.05, size = 16, margin = margin(t=-5, b=1))) +
  scale_shape_manual(
    values=c("SJ" = 19, "CA" = 21, "CA-H"=22, "CA-L" = 15)) + 
  scale_color_manual(
    values=c("SJ" = 'black',"CA" = "blue", "CA-H" = 'blue', "CA-L" = 'blue')) +
  scale_fill_manual(
    values=c("SJ" = 'black', "CA" = "white", "CA-H" = 'white', "CA-L" = 'blue')) +
  scale_linetype_manual(values=c("solid","solid","solid","solid")) +
  scale_size_manual(
    values=c("SJ" = 2, "CA" = 2, "CA-H"=2.5, "CA-L" = 2.5))

grid_arrange_shared_legend(rs, ncol = 1, nrow = 1, position='top')
```

```{r}

# dfcorrela <-

df %>% 
  dplyr::select(-rt, -rc, -rrs) %>% 
  gather(vari, value, -disturb_year, -site) %>% 
  unite(temp, rs, site) %>% 
  spread(temp, value)

```