### 2018-02-27 man/ms_0_intro.Rmd
update texts

---
title: "ms_0_intro"
output:  
  md_document:
    toc: true
    variant: markdown_github
bibliography: ../refs/references.bib
csl: ../refs/ecology.csl
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, echo=FALSE}
show_text <- FALSE
```

# Introduction

# 1) Sequías y Land use change

Drought is a natural phenomenon occurs when water availability is significantly below-normal levels [@Dai2011]. It is one of the most frequent and widespread climatic extremes affecting ecological systems across every climatic zone worldwide [@Mishra2010; @Dai2011; @Ummenhofer2017]. Drought causes significant changes in ecosystem productivity and water dynamics that can severely affect forest through changes in plant physiology, phenology and carbon allocation [@Allen2010; @Ummenhofer2017].

In the last decades there is an crescent interest in assessing the drought impacts on ecosystem, especially considering some studies that suggest an increase in severity, geographic extent and frequency of droughts as a result of climate change [@Dai2011; @Trenberth2014]. There are studies analyzing the response of vegetation to drought from global (*e.g*. @Allen2010; @VicenteSerrano2013; @Norman2016) to local scales (*e.g* @MartinezVilalta2016 for a revision). Some studies have focusing on the drought effects on Mediterranean species which representing their southernmost limit [@Pasho2011; @Camarero2011; @Herrero2014; @Linares2014; @SanchezSalguero2017; @Caminero2018].

Summer drought is a characteristic feature of the Mediterranean climate [@Lionello2012], but an increase in the frequency and severity of drought events have been recorded for the Mediterranean region [@Hoerling2011], particularly for southern Europe [@VicenteSerrano2014; @Spinoni2015; @Stagge2017] where a trend towards drier summers has been reported for the last three decades [@Spinoni2017b]. In addition, climate change projections suggest that extreme events, like droughts, will become more frequent and severe [@Trenberth2014; @IPCC2013] particularly for the Mediterranean area [@Spinoni2017a]. In the Iberian Peninsula, major drought episodies were recorded in 1943, 1981, 1995, 2000 and 2005 [@VicenteSerrano2014; @Guerreiro2017]. The 2004/2005 and 2011/2012 hydrological years are considered two of the worst drought periods recorded in the Iberian Peninsula, particularly in the southern sector [@GarciaHerrera2007; @Gouveia2015; @Trigo2013; @Guerreiro2017; @Pascoa2017]. These events were extreme in both its magnitude and spatial extent [@Gouveia2014].

Vegetation responses to drought are influenced by other drivers of environmental change [@Fischer2006; @Oliver2014; @Franklin2016; @Penuelas2017]. The impacts of drought on ecosystem must be evaluated in a context of global change considering the interactions with other drivers of environmental change [@Archaux2006; @Clavero2011], such us land-use change [@DoblasMiranda2017]. This is specially relevant for areas with a long history of landscape modification such us Mediterranean region where land-use change is a key driver of the global change [@Navarro2013; @Ameztegui2016]. In a recent review, @DoblasMiranda2017 found that land use changes and more especially drought, are crucial factors in the interactions among the drivers of global change for Mediterranean forests.

The impacts of drought are especially relevant for populations located in the retreating or rear edge of their distribution where climatic conditions often corresponds to the species-limits and the populations are likey to be more sensitive to small variations on climate [@Hampe2005]. The response of rear-edge populations to variations on climate could be useful to estimate the adaptation of the species to predicted climatic changes [@Jump2010]. It becomes even more important for relict populations driven by historical land-uses, which can harbour high levels of intraspecific genetic diversity [@Valbuena2010].

There are several works que han analizado la respuesta de la vegetación a la sequía en su rear edge, y en este contexto es interesante analizar no solo el efecto de la sequía en estas poblaciones sino también su resiliencia. -- Resilience is the capacity of an ecosystem, community or individual to restore pre-disturbance structure and function after a disturbance (analogous to ‘engineering resilience’, see Holling, 1996; see also Lloret et al., 2011). Thus, the assessment of forest resilience to climatic disturbances, such as extreme droughts, provides critical information about the capacity of forests to maintain their structure and to continue providing valuable ecosystem services. Furthermore, the analysis of resilience using multiple demographic rates allows us to detect key vulnerabilities to drought in tree population dynamics, such as reductions in recruitment or survival, which can prompt a decline in dominant species shaping forest structure.

Existen trabajos que han analizado los efectos de la sequía en especies situadas en su rear edge (Asier, por ejemplo; Matias) y algunos de ellos se han centrado en ver como responde el crecimiento de las especies en el rear edge utilizando dendro (Herrero Rigling.. // dorado-liñan // sanchez salguero

trabajos rear edge: 
* sanchez salguero Is drought the main decline factor at the rear edge of Europe? The case of southern Iberian pine plantations 
* sanchez salguero Assessing forest vulnerability to climate warming using a process-based model of tree growth: bad prospects for rear-edges
* dorado-liñán: Climate threats on growth of rear-edge European beech peripheral populations in Spain.
* herrero rigling: Varying climate sensitivity at the dry distribution edge of Pinus sylvestris and P. nigra
* dorado-liñan: Large-scale atmospheric circulation enhances the Mediterranean East-West tree growth contrast at rear-edge deciduous forests. 
buscar otros

Sin embargo el estudio de las poblaciones situadas en su rear edge requiere de una aproximación combinada donde se estudie la respuesta de las poblaciones al cambio climático y su resiliencia utiliznado diferentes aproximaciones metodológicas, por ejemplo combinando dendro y remote sensing [@Jump2010].

Remote sensing

These climatic alterations are likely to have important consequences for tree species dynamics at local and regional scales ( Peñuelas and Boada 2003 , Van Mantgem et al. 2009 , Matías and Jump 2015 ).

# 2) Combinación metodológica 
Analizar resiliencia de poblaciones utilizando RS y dendro.

Analizar el impacto de las sequías en los árboles es muy importante y prioriatrio...

* Analysing the impact of drought on tree growth is therefore a priority under the current climatic conditions, in which global warming is probably increasing the severity of droughts (Williams et al., 2013).

En este estudio combinamos el uso de datos de dendrocronologia con información de productividad procedente de satélite. Por un lado, sabemos que el annual radial growth increment se utiliza como proxy para estudiar la forest vitality [@Fritts1976; @Dobbertin2005]; por otro lado el EVI para productividad ...

Una vez dicho esto...

a) Como la dendro ayuda a evaluar las sequías (ver Gazol 2017)

* Gea-Izquierdo 2014: El análisis del inter-annual growth variability can be used to study changes in productiviyt and forest vitality to detect vulnerability to stress (Fritts 1976; Doberti)

* Tree ring width or annual radial growth increment is a widely used proxy for tree vitality (ver  [@Bhuyan2017] [@Dobbertin2005]

b) Remote sensing:

Remote sensing can be used for studying droughts impacts on ecosystems (*e.g.* @Deshayes2006; @Zhang2013; @AghaKouchak2015; @McDowell2015; @Norman2016). Several studies have analyzed drought impacts on vegetation using remote sensing at different scales [@VicenteSerrano2013; @Panisset2018], and for Mediterranean area there were studies focused on local scales (*e.g* @Lloret2007; @Gouveia2012, @Zribi2016) and recently other covering all Mediterranean region [@Gouveia2017].

sin embargo (aqui poner las limitaciones de una u otra metodología).

Jump et al habla de utilizar combinaciones de ambos métodos para estudiar esto, y esto es importante para las poblaciones del rear edge.

Existen algunos estudios que han combinado el uso de dendro y RS para estudiar el efecto de las sequías....

* @Dorman2013 --> interesante compara NDVI y dendro 
* Tree growth and vegetation activity at the ecosystem-scale in the eastern Mediterranean

Aqui nosotros evaluamos el efecto de la sequía ... The vegetation response to water deficit was assessed using remote sensing information and dendrocronologial data.

... ojo los remote sensing datos (aun con alta resolucion temporal y/o espacial) son insuficientes por si solos para demostrar 
 Remote sensing data, even high temporal or spatial resolution data, are insufficient alone to unequivocally demonstrate that drought is the causal agent of a particular change in reflectance values

-----------------------

### ojo un parrafito de resiliencia

Shifts in forest structure and composition in response to drought could alter ecosystem functions profoundly, affecting hydrological processes (Allen, 2007; Adams et al., 2012), nutrient cycling (Hughes et al., 2006; Xiong et al., 2011) and land-surface properties (Royer et al., 2011; Anderegg et al., 2013).

In this context, resilience is an important concept that is gaining unprecedented attention in ecology (e.g. Reyer et al., 2015). Resilience is the capacity of an ecosystem, community or individual to restore pre-disturbance structure and function after a disturbance (analogous to ‘engineering resilience’, see Holling, 1996; see also Lloret et al., 2011). Thus, the assessment of forest resilience to climatic disturbances, such as extreme droughts, provides critical information about the capacity of forests to maintain their structure and to continue providing valuable ecosystem services. Furthermore, the analysis of resilience using multiple demographic rates allows us to detect key vulnerabilities to drought in tree population dynamics, such as reductions in recruitment or survival, which can prompt a decline in dominant species shaping forest structure.

## 
```{r echo=FALSE, results='asis', eval=show_text}
cat("

### Sequías 
> notas sequia
> 
> * the drought events have been longest and most severe in the period 1991-2010 for mediterranean are of Southern Europe  
> * Seasonally, drought frequency is projected to increase everywhere in Europe for both scenarios in spring and summer, especially over southern Europe, and less intensely in autumn [@Spinoni2017a]
> Aumento de la frecuencia de sequías en spring and summer desde 1950 hasta 2014 [@Spinoni2017b]
> Both for frequency and severity, the evolution towards drier conditions is more relevant in the last three decades over Mediterranean area in summer, 
> an increase in the drought severity in the Iberian Peninsula has been observed in the last decades [@VicenteSerrano2014].

Althought several works have reported these two years as some of the worst drought events, we characterised the drought at several spatio-temporal scales in the study area. From a long-term perspective, we compare the accumulated monthly precipitation at a meteorological station (Granada, Base Aérea) during the hydrological years 2004-2005 and 2011-2012 with the average of accumulated monthly precipitation for the period 1950-2015.

### Extreme sequias y ecologia

No solo efectos aislados, sino también en conjunción con otros factores, sobre todo en el mediterráneo donde convergen muchos de los factores que pueden interaccionar:

* La sequía es un factor crucial a tener en cuenta, ya que es además de los efectos que puede tener de forma aislada, se ha visto que además presenta muchas interacciones con otros factores, siendo por tanto un factor crucial (`$IMPROVE$` ver [@DoblasMiranda2017]):
* Some interactions alter the effects of a single factor, as drought enhances or decreases the effects of atmospheric components on plant ecophysiology
* Drought and land use changes, among others, alter water resources and lead to land degradation, vegetation regeneration decline, and expansion of forest diseases.
* Climate change, and especially drought, emerges as a crucial factor in most of the reviewed interactions and therefore it should be considered when it comes to designing and applying international management policies
* Drought should be considered when designing and applying management policies.

Además la sequía se espera que tengan ...Droughts are most likely to have the largest and most long-lasting impacts globally due to large indirect and lagged impacts and long recovery especially for forest ecosystems (ver  18 en @Ummenhofer2017). Así por ejemplo se ha visto que todos los biomas presentan una vulnerabilidad similar ...
En una revision sobre la vulnerabilidad del sistema de trasnporte en plantas al embolismo inducido por sequía, ha mostrado que una convergencia de la vulnerabilidad de los bosques a la sequía, mostrando que todos los biomas son igualmente vulnerables a los fallos hidráulicos independientemente del régimen de precipitación (Choat et al. 2012, doi:10.1038/nature11688)

")
```

.. ver [@VicenteSerrano2013] impacto de las sequías en diferentes zonas (i.e montañas)

Objetivos del trabajo

### Aims
In this study we combined remote sensing information and dendroecological methods to evaluate the drought impacts in both greenees and growth of *Q. pyrenaica* forests in Sierra Nevada. Specifically,

The aims of this work were:

* To quantify how two extreme drought events influenced the greenness and radial growth of *Q. pyreancia* forests in their rear edge,

* to analyze the resilience of these forests to successive extreme drought events,

* and to explore differences in the resilience metrics between populations located in contrasting slopes within the rear edge of the distribution of this species.Do northern and southern populations of Pyrenean oak forests differ in their resistance, resilience or recovery to extreme drought events?

# Otras notas: 
* To guide sustainable forest management, forest researchers are asked to provide concrete answers about forest resilience in response to expected climatic trends, and extreme climatic events (Lindner et al., 2014) https://www.sciencedirect.com/science/article/pii/S030147971400379X?via%3Dihub

* Hablan de incorporar ademas de RS, el conocimiento ecológico (los principios ecológicos) para comprender las consecuencias de las disturbances analizadas con RS, y sobre todo para predecir futuro. (McDowell et al https://www.sciencedirect.com/science/article/pii/S1360138514002726)

----------------------------

* Proyeccciones de crecimeinto de para *Q. pyrenaica* sugieren un declive en el crecimiento en las proximas decadas a lo largo de su rango de distribución en la P. Iberica, en diferentes escenarios climáticos [@GeaIzquierdo2013]: Predictions suggest that QUPY productivity would decline in the next decades all along its distributional range in the Iberian Peninsula for all the climate scenarios studied. Este declive debería ser mas dramático en las baja altitudes de las zonas mas calientes del sur de su distribución

--> vulnerabilidad local (related to rear-edge)
La tendencia de reducción drástica sugerida por el modelo para el dry-edge de la distribución de la especie podría expresar una aumentada vulnerabilidad de los árboles al incremento del estrés hídrico forzado por un clima mas cálido [@GeaIzquierdo2013].

Negative trends over recent years and the greater response to moisture availability found at warmer and drier low elevations in the south suggest vulnerability to warming at the local low elevation dry edge of the species's range (e.g. QUPY9). Otros estudios en la zona Mediterranea reportaron un descenso similar en la productividad con un incremento en la vulterabilidad a la sequía ()

Muchos trabajos han analizado la respuesta de la vegetación a las sequías (eg. @Allen2010; @VicenteSerrano2013; @MartinezVilalta2016; @Norman2016), y algunos de ellos han mostrado el efecto de las sequías en especies Mediterráneas (ver por ejemplo Pasho et al. https://www.sciencedirect.com/science/article/pii/S016819231100253X#fig0020; @Camarero2011 combina uso y sequía; añadir alguno mas de GGI). Algunos Some of these species represent southernmost populations in the Mediterranean ambit, which explains their vulnerability to the warmer conditions (Andreu et al., 2007; Sánchez-Salguero et al., 2016).

Case studies that focus on multiple scales - including local scales - and also valuables (Dale et al. 2018 Frontiers) --> Se necesitan casos de estudio que se enfoquen en escalas multiples (incluyendo las locales) y que combinen varias metodologías (Jump et al. ...)

3) rear edge

- además (Vulnerabilidad de los bosques a los eventos climáticos extremos ...):

* Existen evidencias que sugieren que muchos bosques son vulnerables a eventos climáticos extremos ... (Zhang) y esto puede ser especialmente relevante para especies situadas en el rear edge (completar) ...

... Varios trabajos han apuntado la existencia de vulnerabilidad local al aumento de temperaturas en los sitios del sur de distribución de esta especie [@GeaIzquierdo2013; @GeaIzquierdo2014] así como una strong response to moisture availability...; Además teniendo en cuenta que en los últimos años estamos viendo un aumento en las sequías (menor disponibilidad de agua) --> por ello queremos analizar como están respondiendo esta especie, sobre todo en esas zonas dry-egde, y analizar su resiliencia...