### 2018-10-16 man/figures/0_figures.Rmd
last update

---
output:
  word_document:
    fig_caption: yes
    highlight: null
    reference_docx: ../../templates/template_pagebreak.docx
bibliography: ../../refs/references.bib
csl: ../../refs/ecology.csl
---

```{r child = 'location_map.Rmd'}
```

#####

```{r child = 'std_anomalies.Rmd'}
```