### 2018-11-08 man/figures/growth_changes.Rmd
ya casi lo tengo

---
output: 
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    highlight: null
    reference_docx: ../../templates/template_pagebreak.docx
bibliography: ../../refs/references.bib
csl: ../../refs/ecology.csl
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE, dpi = 200)
```

```{r}
library("tidyverse")
library("here")
library("dplyr")
library("ggrepel")
library("scales")
library("grid")
library("gridExtra")
library("patchwork")
library("ggcorrplot")
library("stringr")
library("pander")
library("lemon")
library("magrittr")
```

```{r, echo=FALSE}
show_text <- FALSE
```

**Figure 7.** Comparison of median growth change ($GC$) following @Nowacki1997 for *Q. pyrenaica* sites. Dashed black lines indicate a threshold of 50 % of GC (see material and methods).

```{r}
gc_todos <- read.csv(here("data/disturbance", "disturbance_gc.csv"), header = TRUE, sep = ',')

# gc_todos %<>% mutate(yearF = as.Date(as.character(year), format="%Y"))

# customize plot (corrected scale)
pos_neg_log <- scales::trans_new('signed_log', 
                                 transform = function(x) sign(x)*log(abs(x)),
                                 inverse = function(x) sign(x)*exp(abs(x)))

custom_ylab <- '% GC'
micolor <- c("#838B8B")

# solution for arrange
# http://felixfan.github.io/stacking-plots-same-x/

custom_corrected_gc <- list(
  geom_bar(stat='identity', fill=micolor, colour=micolor),
  theme_bw(), 
  geom_hline(yintercept=c(50, -50), linetype=2, colour='black'),
  theme(panel.grid.minor = element_blank(), 
        strip.background = element_rect(colour='black',fill='white'))
  )

gccaL <- gc_todos %>% 
  filter(site == 'gc_caL') %>% 
  ggplot(aes(x=year, y=gc_md)) + 
  geom_hline(yintercept = 0, colour = 'black', lwd = 0.3) + 
  geom_errorbar(aes(ymin = gc_md - gc_md_se, ymax = gc_md + gc_md_se), colour='gray', width=.3, size=.2) + 
  # scale_y_continuous(trans = pos_neg_log) +
  # facet_wrap(~site_sorted, nrow = 2,
  #            labeller=as_labeller(c("1_caH" = "CA-High", "2_caL" = "CA-Low"))) + 
    ylab(custom_ylab) + custom_corrected_gc

gccaH <- gc_todos %>% 
  filter(site == 'gc_caH') %>% 
  ggplot(aes(x=year, y=gc_md)) + 
  geom_hline(yintercept = 0, colour = 'black', lwd = 0.3) + 
  geom_errorbar(aes(ymin = gc_md - gc_md_se, ymax = gc_md + gc_md_se), colour='gray', width=.3, size=.2) + 
  # scale_y_continuous(trans = pos_neg_log) +
  # facet_wrap(~site_sorted, nrow = 2,
  #            labeller=as_labeller(c("1_caH" = "CA-High", "2_caL" = "CA-Low"))) + 
    ylab(custom_ylab) + custom_corrected_gc

ggplot_build(gccaH)$layout$panel_scales_y[[1]]$range$range
ggplot_build(gccaL)$layout$panel_scales_y[[1]]$range$range

##### -------------------------
# VALIDO 
# NEW
th <- list(

)

custom_corrected_gc <- list(
  geom_bar(stat='identity', fill=micolor, colour=micolor),
  theme_bw(), 
  geom_hline(yintercept=c(50, -50), linetype=2, colour='black'),
  theme(panel.grid.minor = element_blank(), 
        strip.background = element_rect(colour='black',fill='white'))
  )

# Function for plot

plotGC <- function(df, y, yse, x, my_ylab, 
                   micolor, ...){ 
  # Get error bar aesthetic
  yerrorbar <- aes_string(ymin = paste(y, yse, sep = '-'),
                         ymax = paste(y, yse, sep = '+'))

  p <- ggplot(df, aes_string(x = x, y = y)) +
    geom_errorbar(mapping = yerrorbar, colour = 'gray', width = .35, size = .3) + 
    geom_bar(stat = 'identity', fill = micolor, colour = micolor) +
    geom_hline(yintercept = 0, colour = 'black', lwd = 0.3) + 
    ylab(my_ylab) +
    # Add line to %GC = 50 
    geom_hline(yintercept = c(50, -50), linetype=2, colour='black') +
    theme_bw() +
    theme(panel.grid = element_blank()) 
    # Customize limits and breaks to min of year data 
  p
}

# Get min and max of year range and create vector of 10 years 
xrange <- seq(round(min(gc_todos$year) / 10) * 10, 
    round(max(gc_todos$year) / 10) * 10, by = 10)

    scale_x_continuous(limits = limitesx, 
                       breaks = xrange, 
                       position = pos) 
                       #sec.axis = dup_axis(name = NULL))

# Get automatically limits of y for caH anc caL
minimoY <- min((gc_todos %>% 
      dplyr::filter(site != 'gc_sj' & gc_md < 0) %>% 
      mutate(error = gc_md - gc_md_se))$error, na.rm = TRUE)

maximoY <- max((gc_todos %>% 
      dplyr::filter(site != 'gc_sj' & gc_md > 0) %>% 
      mutate(error = gc_md + gc_md_se))$error, na.rm = TRUE)

# Ojo ver limites de ylim 
# ggplot_build(pcaL)$layout$panel_params[[1]]$y.range
# ggplot_build(pcaH)$layout$panel_params[[1]]$y.range

# Establece rango de datos para y 
mismo_y <- scale_y_continuous(limits=c(minimoY, maximoY), 
                             breaks = c(-100, -50, 0, 50, 50, 100, 150))

mismo_x <- scale_x_continuous(limits=c(min(gc_todos$year), max(gc_todos$year)), 
                             breaks = xrange, position = 'bottom')

#### --> Hay un error con la localizacion de las barras de error en el eje x. REVISAR

pcaH <- gc_todos %>% dplyr::filter(site == 'gc_caH') %>% 
  plotGC(y = 'gc_md', yse = 'gc_md_se', x = 'year', 
         my_ylab = '% GC', micolor = "#838B8B") + mismo_y + mismo_x +
  annotate("text", x = 1990, y = 120, label = "CA-High", fontface = 2, size = 5)

pcaL <- gc_todos %>% dplyr::filter(site == 'gc_caL') %>% 
  plotGC(y = 'gc_md', yse = 'gc_md_se', x = 'year',
         my_ylab = '% GC', micolor = "#838B8B") +  
         # xrange=xrange, pos = 'bottom', limitesx = limitesx) + 
  ylab('') + mismo_y + mismo_x + 
  annotate("text", x = 1990, y = 120, label = "CA-Low", fontface = 2, size = 5)

psj <- gc_todos %>% dplyr::filter(site == 'gc_sj') %>% 
  plotGC(y = 'gc_md', yse = 'gc_md_se', x = 'year', 
         my_ylab = '% GC', micolor = "#838B8B") + 
         # xrange=xrange, pos = 'top', limitesx = limitesx) + 
  ylab('') + xlab('') + 
  scale_y_continuous(trans = pos_neg_log, 
                     breaks=c(-1000, -500, -100, 0, 100, 500)) + mismo_x + 
  geom_hline(yintercept = 1) +
  annotate("text", x = 1990, y = 500, label = "SJ", fontface = 2, size = 5)

c(-116, 150)

ggplot_build(pcaH)$layout$panel_scales_y[[1]]$range$range
ggplot_build(gccaL)$layout$panel_scales_y[[1]]$range$range

layer_scales(pcaH)$y$range$range

layer_scales(c)$x$range$range[2]) +

gc_todos %>% dplyr::filter(site == 'gc_caL') %>% 
  plotGC(y = 'gc_md', yse = 'gc_md_se', x = 'year', 
       my_ylab = NULL, micolor = "#838B8B", xrange=xrange)

gc_todos %>% dplyr::filter(site != 'gc_sj') %>% 
  plotGC(y = 'gc_md', yse = 'gc_md_se', x = 'year', 
       my_ylab = NULL, micolor = "#838B8B", xrange=xrange) + 
  facet_rep_wrap(~site, ncol = 1, strip.position = 'bottom') + 
  theme(panel.spacing = unit(0, "lines"),
        strip.placement = "outside", 
        strip.background = element_blank())

##### -------------------------
# END VALIDO

# Round to the nearest decade 
# https://stackoverflow.com/questions/35352914/floor-a-year-to-the-decade-in-r 
# options (lubridate::round_date(min(gc_todos$yearF), 'years')

custom_ylab <- '% GC'
micolor <- c("#838B8B")

gccaL <- gc_todos %>% 
  filter(site == 'gc_caL') %>% 
  ggplot(aes(x=year, y=gc_md)) +

gccaL <- gc_todos %>% 
  filter(site == 'gc_caL') %>% 
  ggplot(aes(x=year, y=gc_md)) + 
  geom_hline(yintercept = 0, colour = 'black', lwd = 0.3) + 
  geom_errorbar(aes(ymin = gc_md - gc_md_se, ymax = gc_md + gc_md_se), colour='gray', width=.3, size=.2) + 
  # scale_y_continuous(trans = pos_neg_log) +
  # facet_wrap(~site_sorted, nrow = 2,
  #            labeller=as_labeller(c("1_caH" = "CA-High", "2_caL" = "CA-Low"))) + 
    ylab(custom_ylab) + custom_corrected_gc

c <- gc_todos %>% 
  filter(site != 'gc_sj') %>% 
  ggplot(aes(x=year, y=gc_md)) + 
  geom_hline(yintercept = 0, colour = 'black', lwd = 0.3) + 
  geom_errorbar(aes(ymin = gc_md - gc_md_se, ymax = gc_md + gc_md_se), colour='gray', width=.3, size=.2) + 
  # scale_y_continuous(trans = pos_neg_log) +
  facet_wrap(~site_sorted, nrow = 2,
             labeller=as_labeller(c("1_caH" = "CA-High", "2_caL" = "CA-Low"))) + 
    ylab(custom_ylab) + custom_corrected_gc

d <- gc_todos %>% 
  filter(site == 'gc_sj') %>% 
  ggplot(aes(x=year, y=gc_md)) + 
  xlim(layer_scales(c)$x$range$range[1],layer_scales(c)$x$range$range[2]) +
  geom_hline(yintercept = 1, colour = 'gray', lwd = 0.3) + 
  geom_errorbar(aes(ymin = gc_md - gc_md_se, ymax = gc_md + gc_md_se), colour=micolor, width=.3, size=.2) +
  facet_wrap(~site_sorted, nrow = 2,
             labeller=as_labeller(c("0_sj" = "SJ"))) + 
  custom_corrected_gc + ylab('') + 
  scale_y_continuous(trans = pos_neg_log, 
                     breaks=c(-1000, -500, -100, 0, 100, 500)) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())

plot_medians_corregido <- d + c + plot_layout(ncol=1, heights = c(1, 2))
plot_medians_corregido
```

gc_todos <- read.csv(here::here("data/disturbance", "disturbance_gc.csv"), header = TRUE, sep = ',')

# customize plot (corrected scale)
pos_neg_log <- scales::trans_new('signed_log', 
                                 transform = function(x) sign(x)*log(abs(x)),
                                 inverse = function(x) sign(x)*exp(abs(x)))

custom_ylab <- '% GC'
micolor <- c("#838B8B")

# solution for arrange
# http://felixfan.github.io/stacking-plots-same-x/

custom_corrected_gc <- list(
  geom_bar(stat='identity', fill=micolor, colour=micolor),
  theme_bw(), 
  geom_hline(yintercept=c(50, -50), linetype=2, colour='black')

        # strip.background = element_rect(colour='black',fill='white'))
  )

# Function to return a desired sequence breaks along axes
# https://gist.github.com/valentinitnelav/ec7c253c80f4e045f4bd1c9b2920d7bd 
my.breaks <- function(my.vector, step){
    my.min <- floor(min(my.vector))
    my.max <- ceiling(max(my.vector))
    my.seq <- seq(from = round(my.min/step)*step, 
                  to   = round(my.max/step)*step,
                  by   = step)
    return(my.seq)
}

gc_todos %>% 
  filter(site == 'gc_caH') %>% 
  ggplot(aes(x=year, y=gc_md)) + 
  geom_hline(yintercept = 0, colour = 'black', lwd = 0.3) + 
  geom_errorbar(aes(ymin = gc_md - gc_md_se, ymax = gc_md + gc_md_se), colour='gray', width=.3, size=.2) +
  ylab(custom_ylab) +
  GCtheme + 
  ylim(-100, 150) +
  scale_x_continuous(name = "Year", breaks = my.breaks(gc_todos$year, step = 25))

GCtheme <- list(
  # BAR
  geom_bar(stat='identity', fill=micolor, colour=micolor),
  geom_hline(yintercept=c(50, -50), linetype=2, colour='black'),
  theme_bw(),
  theme(# AXIS ticks
        axis.ticks.length = unit(-1.4, "mm"),
        axis.text.x = element_text(size = 8, margin = unit(c(t = 2.5, r = 0, b = 0, l = 0), "mm")),
        panel.grid.minor = element_blank(),
        # Remove strip 
        strip.background = element_blank(),
        strip.text = element_blank())
)

d <-

  gc_todos %>% 
  # mutate(year = lubridate::ymd(year, truncated = 2L)) %>% 
  filter(site != 'gc_sj') %>% 
  ggplot(aes(x=year, y=gc_md)) + 
  geom_hline(yintercept = 0, colour = 'black', lwd = 0.3) + 
  geom_errorbar(aes(ymin = gc_md - gc_md_se, ymax = gc_md + gc_md_se), colour='gray', width=.3, size=.2) + 
  # scale_y_continuous(trans = pos_neg_log) +
  facet_wrap(~site_sorted, nrow = 2) +
             # labeller=as_labeller(c("1_caH" = "CA-High", "2_caL" = "CA-Low"))) + 
    ylab(custom_ylab) + custom_corrected_gc +
  scale_x_continuous(name = "Year", breaks = my.breaks(gc_todos$year, step = 25))

  scale_x_date(date_breaks = "25 year", date_labels = "%Y")

d <- gc_todos %>% 
  filter(site == 'gc_sj') %>% 
  ggplot(aes(x=year, y=gc_md)) + 
  xlim(layer_scales(c)$x$range$range[1],layer_scales(c)$x$range$range[2]) +
  geom_hline(yintercept = 1, colour = 'gray', lwd = 0.3) + 
  geom_errorbar(aes(ymin = gc_md - gc_md_se, ymax = gc_md + gc_md_se), colour=micolor, width=.3, size=.2) +
  facet_wrap(~site_sorted, nrow = 2,
             labeller=as_labeller(c("0_sj" = "SJ"))) + 
  custom_corrected_gc + ylab('') + 
  scale_y_continuous(trans = pos_neg_log, 
                     breaks=c(-1000, -500, -100, 0, 100, 500)) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())

plot_medians_corregido <- d + c + plot_layout(ncol=1, heights = c(1, 2))
plot_medians_corregido

```