### 2018-11-10 man/figures/gc_perc_trees.Rmd
add figure corrected

---
output: 
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    highlight: null
    reference_docx: ../../templates/template_pagebreak.docx
bibliography: ../../refs/references.bib
csl: ../../refs/ecology.csl
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE, dpi = 200)
```

```{r}
library("tidyverse")
library("here")
library("dplyr")
library("ggrepel")
library("scales")
library("grid")
library("gridExtra")
library("patchwork")
library("stringr")
library("pander")
library("lemon")
library("magrittr")
```

```{r, echo=FALSE}
show_text <- FALSE
```

##### 
**Figure 8**. Percentage of *Q. pyrenaica* trees affected by GC > 50 % by site. *Black* line shows number of trees (rigth-axis). Data for number of trees > 2 is shown.

```{r}
gc_treesMed <- read.csv(here("data/disturbance", "disturbance_trees_gc_treshold_median.csv"), header = TRUE, sep = ',')
gc_treesMed50 <- read.csv(here("data/disturbance", "disturbance_trees_gc_treshold_median50.csv"), header = TRUE, sep = ',')

colorpositivo <- c("#104E8B")
colornegativo <- c("#CD3333")

# Get min and max year and round to nearest decade
getBreaks <- function(df, v, by){ 
  x <- seq(round(min(df[v]) / 10)*10, 
           round(max(df[v]) / 10)*10, by= by) 
  x
  }

xrange <- getBreaks(gc_treesMed50, 'year', by=10)

label_site <- c("0_sj" = "SJ",
                "1_caH" = "CA-High",
                "2_caL" = "CA-Low")

pgc_median50 <- gc_treesMed50 %>% 
  filter(total_tree > 2) %>% 
  ggplot() +
  geom_bar(aes(x=year, y=per_pos), fill=colorpositivo, colour=colorpositivo, 
           stat = "identity", position = "dodge") + 
  geom_bar(aes(x=year, y=(-1)*per_neg), fill=colornegativo, colour=colornegativo, 
           stat = "identity", position = "dodge") +
  geom_line(aes(x=year, y=total_tree), colour = 'black') + 
  theme_bw() + ylab('% trees') +
  geom_hline(yintercept=0, alpha=.4) +
  geom_hline(yintercept=c(50, -50), linetype=2, alpha=.4) +
  scale_x_continuous(breaks = xrange) +
  scale_y_continuous(limits = c(-100, 100), breaks = seq(-100, 100, by = 50),
                     sec.axis = sec_axis(~ ., name = '# trees', breaks = c(0,15,30))) +
  facet_rep_wrap(~site_sorted, nrow = 3, 
                 repeat.tick.labels = TRUE, labeller = as_labeller(label_site)) +
  theme(panel.grid = element_blank(), 
        strip.background = element_blank(), 
        strip.text = element_blank(),
        axis.text.y = element_text(size=10),
        axis.text.x = element_text(size=9),
        axis.title = element_text(size = 12))

pgc_median50
```