### 2018-11-16 man/tables/t_robust_anovas.Rmd

---
output:  
  pdf_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, echo=FALSE}
library(tidyverse)
library(here)
library(pander)
library(kableExtra)
library(knitr)
library(dplyr)
```

**Table 3.** Robust two-way ANOVAs of the resilience metrics of greenness (EVI) and tree-growth (BAI) for the two factors drought events (in 2005 and 2012) and site.

```{r, echo=FALSE, include = FALSE}
### Table R1_evi
evi <- read.csv(here("/out/anovas_resilience/huber_evi", "trimmed_anovas.csv"), header = TRUE)
names(evi) <- c('factores', 'F', 'p', 'variables')

dff_evi <- evi %>% gather(temp, numeros, F, p) %>% 
  unite(temp1, variables, temp) %>% 
  spread(temp1, numeros) %>% 
  mutate(factores = recode(factores, 
                           "disturb_year" = "Drought event", 
                           "disturb_year:site" = "Drought event:Site", 
                           "site" = "Site")) %>% 
  arrange(match(factores, c("Drought event", "Site", "Drought event:Site"))) %>% 
  mutate_at(c(2,4,6,8), round, 1) %>% 
  mutate_at(c(3,5,7,9), round, 3) %>% 
  mutate(variable = 'EVI')

bai <- read.csv(here("/out/anovas_resilience/huber_bai", "trimmed_anovas.csv"), header = TRUE)
names(bai) <- c('factores', 'F', 'p', 'variables')

dff_bai <- bai %>% gather(temp, numeros, F, p) %>% 
  unite(temp1, variables, temp) %>% 
  spread(temp1, numeros) %>% 
  mutate(factores = recode(factores, 
                           "disturb_year" = "Drought event", 
                           "disturb_year:site" = "Drought event:Site", 
                           "site" = "Site")) %>% 
  arrange(match(factores, c("Drought event", "Site", "Drought event:Site"))) %>% 
  mutate_at(c(2,4,6,8), round, 1) %>% 
  mutate_at(c(3,5,7,9), round, 3) %>% 
  mutate(variable = 'BAI')

dff <- dff_evi %>% 
  bind_rows(dff_bai) %>% 
  dplyr::select(-rrs_F, -rrs_p) %>% 
  dplyr::select(variable, factores, rt_F, rt_p, rc_F, rc_p, rs_F, rs_p)
```

```{r, echo=FALSE}
# # https://haozhu233.github.io/kableExtra/use_kableExtra_with_formattable.html
# colnames(dff) <- c("variable", "factors", "F", "p", "F", "p", "F", "p")
# colnames(dff) <- c("variable", "factors", "rt\nF", "rt\np","rc\nF", "rc\np", "rs\nF", "rs\np")
# kable(dff, format = "latex") %>% add_header_above(c("", "", "Resistance" = 2, "Recovery" = 2, "Resilience" = 2))
#

# https://davidgohel.github.io/flextable/articles/examples.html

mt <- regulartable(
  data = dff, 
  col_keys = c("variable", "factores", "c3","rt_F", "rt_p", "c1", "rc_F", "rc_p","c2", "rs_F", "rs_p"),
)

mt <- theme_vanilla(mt)
mt <- autofit(mt)
mt <- empty_blanks(mt)
mt <- align(mt, j = 1,  align = "center")

mt <- merge_v(mt, ~variable)

mt <- set_header_labels(mt, variable = "", factores = "", rt_F = "Resistance", rt_p = "Resistance", rc_F = "Recovery", rc_p = "Recovery", rs_F = "Resilience", rs_p = "Resilience")

# # Add row of labels 
mt <- add_header(mt, variable = "Variable", factores = "Factor", rt_F = "F", rt_p = "p", rc_F = "F", rc_p = "p", rs_F = "F", rs_p = "p", top = FALSE)

# Format 
mt <- bold(mt, part = 'header') 
mt <- fontsize(mt, part = "header", size = 12)
mt <- fontsize(mt, part = "body", size = 11)
mt <- width(mt, j = 2, width = 2)

# mt <- align(mt, i = 1, j=2:3, align = "left")

mt <- italic(mt, i = ~ rt_p < 0.01, j = ~ rt_p)
mt <- italic(mt, i = ~ rc_p < 0.01, j = ~ rc_p)
mt <- italic(mt, i = ~ rs_p < 0.01, j = ~ rs_p)
mt

mt <- set_formatter(mt, 
                    rt_F = function(x) format(round(x, 1)),
                    rc_F = function(x) format(round(x, 1)),
                    rs_F = function(x) format(round(x, 1))
                    )

mt

```