### 2018-11-16 man/tables/t_robust_anovas.Rmd

---
output:
  word_document:
    fig_caption: yes
    highlight: null
    reference_docx: ../../templates/template_tables_portrait.docx
bibliography: ../../refs/references.bib
csl: ../../refs/ecology.csl
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
library(here)
library(pander)
library(kableExtra)
library(knitr)
library(dplyr)
library(flextable)
library(officer)
```

**Table 3.** Robust two-way ANOVAs of the resilience metrics of greenness (EVI) and tree-growth (BAI) for the two factors drought events (in 2005 and 2012) and site.

```{r}
### Table R1_evi
evi <- read.csv(here("/out/anovas_resilience/huber_evi", "trimmed_anovas.csv"), header = TRUE)
names(evi) <- c('factores', 'F', 'p', 'variables')

dff_evi <- evi %>% gather(temp, numeros, F, p) %>% 
  unite(temp1, variables, temp) %>% 
  spread(temp1, numeros) %>% 
  mutate(factores = recode(factores, 
                           "disturb_year" = "Drought event", 
                           "disturb_year:site" = "Drought event:Site", 
                           "site" = "Site")) %>% 
  arrange(match(factores, c("Drought event", "Site", "Drought event:Site"))) %>% 
  mutate_at(c(2,4,6,8), round, 1) %>% 
  mutate_at(c(3,5,7,9), round, 3) %>% 
  mutate(variable = 'EVI')

bai <- read.csv(here("/out/anovas_resilience/huber_bai", "trimmed_anovas.csv"), header = TRUE)
names(bai) <- c('factores', 'F', 'p', 'variables')

dff_bai <- bai %>% gather(temp, numeros, F, p) %>% 
  unite(temp1, variables, temp) %>% 
  spread(temp1, numeros) %>% 
  mutate(factores = recode(factores, 
                           "disturb_year" = "Drought event", 
                           "disturb_year:site" = "Drought event:Site", 
                           "site" = "Site")) %>% 
  arrange(match(factores, c("Drought event", "Site", "Drought event:Site"))) %>% 
  mutate_at(c(2,4,6,8), round, 1) %>% 
  mutate_at(c(3,5,7,9), round, 3) %>% 
  mutate(variable = 'BAI')

dff <- dff_evi %>% 
  bind_rows(dff_bai) %>% 
  dplyr::select(-rrs_F, -rrs_p) %>% 
  dplyr::select(variable, factores, rt_F, rt_p, rc_F, rc_p, rs_F, rs_p)
```

```{r}
mt <- regulartable(
  data = dff, 
  col_keys = c("variable", "factores", "c3","rt_F", "rt_p", "c1", "rc_F", "rc_p","c2", "rs_F", "rs_p"),
)

mti <- mt %>% 
  autofit() %>% 
  empty_blanks() %>% 
  align(j = 1,  align = "center") %>% 
  merge_v(~variable) %>% 
  set_header_labels(variable = "", factores = "", rt_F = "Resistance", rt_p = "Resistance", rc_F = "Recovery", rc_p = "Recovery", rs_F = "Resilience", rs_p = "Resilience") %>% 
  add_header(variable = "Variable", factores = "Factor", rt_F = "F", rt_p = "p", rc_F = "F", rc_p = "p", rs_F = "F", rs_p = "p", top = FALSE) %>% 
  bold(part = 'header') %>% 
  fontsize(part = "header", size = 12) %>%
  fontsize(part = "body", size = 11) %>% 
  width(j = 2, width = 2) %>% 
  italic(i = ~ rt_p < 0.01, j = ~ rt_p) %>% 
  italic(i = ~ rc_p < 0.01, j = ~ rc_p) %>% 
  italic(i = ~ rs_p < 0.01, j = ~ rs_p)
mti

# mt <- set_formatter(mt, 
#                     rt_F = function(x) format(round(x, 1)),
#                     rc_F = function(x) format(round(x, 1)),
#                     rs_F = function(x) format(round(x, 1))
#                     ) 
# 
# mt

```