### 2018-11-23 man/figures/growth_changes.Rmd
adjust sizes

---
output:
  word_document
bibliography: ../../refs/references.bib
csl: ../../refs/ecology.csl
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```

```{r load_pkg}
library("tidyverse")
library("here")
library("scales")
library("patchwork")
```

**Figure 5.** Comparison of median growth change ($GC$) following @Nowacki1997 for *Q. pyrenaica* sites. Dashed black lines indicate a threshold of 50 % of GC (see material and methods). Note that y-axes do not correspond in all of the three panels for the sake of clarity.

```{r data}
gc_todos <- read.csv(here::here("data/disturbance", "disturbance_gc.csv"), header = TRUE, sep = ',')

# customize plot (corrected scale)
pos_neg_log <- scales::trans_new('signed_log', 
                                 transform = function(x) sign(x)*log(abs(x)),
                                 inverse = function(x) sign(x)*exp(abs(x)))
```

```{r plot_function}
# Function for plot

plotGC <- function(df, y, yse, x, my_ylab, 
                   micolor, limitesX, ...){ 
  # Get error bar aesthetic
  yerrorbar <- aes_string(ymin = paste(y, yse, sep = '-'),
                         ymax = paste(y, yse, sep = '+'))

  p <- ggplot(df, aes_string(x = x, y = y)) +
    geom_errorbar(mapping = yerrorbar, colour = 'gray', width = .35, size = .3, 
                  na.rm = TRUE) + 
    geom_bar(stat = 'identity', fill = micolor, colour = micolor) +
    geom_hline(yintercept = 0, colour = 'black', lwd = 0.3) + 
    ylab(my_ylab) +
    # Add line to %GC = 50 
    geom_hline(yintercept = c(50, -50), linetype=2, colour='black') +
    theme_bw() +
    theme(panel.grid = element_blank(),
          axis.text.y = element_text(size = 5.5),
          axis.text.x = element_text(size = 6)) +
    coord_cartesian(xlim=limitesX)
    # Customize limits and breaks to min of year data 
  p
}

```

```{r plot_configure}
# ---------- Min and Max axis        
# Get min and max of year range and create vector of 10 years for labelling
xrange <- seq(round(min(gc_todos$year) / 10) * 10, 
    round(max(gc_todos$year) / 10) * 10, by = 10)

limitesX <- c(min(gc_todos$year), max(gc_todos$year))

# Get automatically limits of y for caH anc caL
minimoY <- min((gc_todos %>% 
      dplyr::filter(site != 'gc_sj' & gc_md < 0) %>% 
      mutate(error = gc_md - gc_md_se))$error, na.rm = TRUE)

maximoY <- max((gc_todos %>% 
      dplyr::filter(site != 'gc_sj' & gc_md > 0) %>% 
      mutate(error = gc_md + gc_md_se))$error, na.rm = TRUE)

# Establece rango de datos para y 
mismo_y <- scale_y_continuous(limits=c(minimoY, maximoY), 
                            breaks = c(-100, -50, 0, 50, 50, 100, 150))
```

```{r plot_caH}
pcaH <- gc_todos %>% dplyr::filter(site == 'gc_caH') %>% 
  plotGC(y = 'gc_md', yse = 'gc_md_se', x = 'year', 
         my_ylab = '% GC', micolor = "#838B8B", limitesX = limitesX) + mismo_y +
  scale_x_continuous(breaks = xrange, position = 'bottom') + xlab('') + 
  annotate("text", x = 1990, y = 120, label = "CA-High", fontface = 2, size = 3)
```

```{r plot_caL}
pcaL <- gc_todos %>% dplyr::filter(site == 'gc_caL') %>% 
  plotGC(y = 'gc_md', yse = 'gc_md_se', x = 'year',
         my_ylab = NULL, micolor = "#838B8B", limitesX = limitesX) +  mismo_y +
  scale_x_continuous(breaks = xrange, position = 'bottom') + xlab('Year') + 
  annotate("text", x = 1990, y = 120, label = "CA-Low", fontface = 2, size = 3) 
```

```{r plot_SJ}
psj <- gc_todos %>% dplyr::filter(site == 'gc_sj') %>% 
  plotGC(y = 'gc_md', yse = 'gc_md_se', x = 'year', 
         my_ylab = NULL, micolor = "#838B8B", limitesX = limitesX) + 
  xlab('') + 
  scale_x_continuous(breaks = xrange, position = 'bottom') + 
  scale_y_continuous(trans = pos_neg_log, 
                     breaks=c(-1000, -500, -100, 1, 100, 500)) + 
  geom_hline(yintercept = 1)+
  annotate("text", x = 1990, y = 440, label = "SJ", fontface = 2, size = 3)

```

```{r plot_join }
gg <- psj + pcaH + pcaL + 
  plot_layout(ncol=1) & 
  # theme(plot.tag = element_text(size = 12, face = 2)) & 
  theme(plot.margin = margin(t=-1,r=3,b=-1,l=3, "pt"))
```

```{r out_plot, fig.height=5, dpi = 300, out.width = 0.8}
gg

# gg + plot_annotation(tag_levels = c('a','b', 'c'))  
```