### 2019-04-04 man/figures/nls_long_term_resilience.Rmd
add pvalue and rsquared

---
output:
  word_document
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```

```{r load_pkg}
library("tidyverse")
library("here")
library("scales")
library("patchwork")
library("ggrepel")
```

```{r data}
df <- read.csv(file=here::here('/out/bai_resiliences_severes' ,'resiliences_bai_droughts.csv'), header=TRUE)

# peores sequias 
dy <- read.csv(file=here::here('data/sequias', 'severe_spei12.csv'), header=TRUE)
dy <- dy %>% arrange(desc(d_duration))
# solamente aquellos mayor de 2 meses 
dys <- dy %>% filter(d_duration > 2) %>% filter(maxyear > 1940 ) %>% as.data.frame() 
# dys <- dys[,'maxyear']

sdf <- df %>% 
  filter(samps >= 10) %>% 
  group_by(disturb_year) %>% 
  summarise_at(c("rt_mean", "rs_mean", "rc_mean", "rrs_mean"), 
               funs(mean(., na.rm = TRUE), sd(., na.rm=TRUE),se=sd(., na.rm=TRUE)/sqrt(n()))) %>% 
  inner_join(dy, by=c("disturb_year" = "maxyear"))

```

TE Quedaste por aquí. No acabaste de encontrar un NLS mejor que la LM

```{r nls}
# subset de datos para Non linear squares
d <- sdf %>% 
  dplyr::select(d_severity, rs_mean_mean) %>% 
  rename(sev = d_severity, rs = rs_mean_mean) %>%  as.data.frame()

library("minpack.lm")
library("nls2")
library("nlshelper")
library("nlstools")

d$ly <- log(d$rs)
d$lx <- log(d$sev)

summary(lm(ly~lx, data=d))

d$y <- log(d$rs-0.6)

# Catalogo de funciones en Catalog of Curves for Curve Fitting
# V. Sit and M. Poulin-Costello 1994 Biometrics Information Handbook 
# https://www.for.gov.bc.ca/hfd/pubs/docs/bio/bio04.htm

set.seed(1467)
# Exponential type I

pF <- function(x,a,b,c) {
  I(a*x^b + c)
}

fe1 <- as.formula(rs ~ I(a*sev^b + c))
m1 <- nls(fe1, data = d, start = list(a=2.755, b=.00014, c=.85), 
          control = list(maxiter = 500), trace = TRUE)
summary(m1)
abline(lm(rs~sev, data=d), col="blue")

summary(lm(rs~sev, data=d))

grid.start <- expand.grid(list(
  a = seq(0.001, 0.03, by=0.001),
  b = seq(1.8, 2.2, by=0.01),
  c= c(0.8)))

m2 <- nls2(rs~pF(sev, a,b,c), data = d, 
          start = grid.start,
          algorithm = "brute-force")

preview(fe1, data = d, start = list(a= 0.5, b = 0.23, c=0.8))

e1 <- nls(fe1, data = d, start = list(a= 0.5, b = 0.18, c=0.8),
          trace = TRUE, control = list(maxiter = 100))
summary(e1)

set.seed(1467)

te <- nls(rs ~ I(exp(1)^(a+b*sev)), data = d, start = list(a=0.8, b = 0.03), trace = TRUE)
summary(te)

f <- as.formula(d$rs ~ d$sev + d$sev^2)
lf <- lm(f, data=d)

alpha <- 0.05
df.new <- data.frame(rs=seq(2,24, 2))
conf.dist <- predict(lf, newdata = df.new, interval="confidence", level=1-alpha) 
pred.dist <- predict(lf, newdata = df.new, interval="prediction", level=1-alpha) 
df.new[c("fit","lwr.conf", "upr.conf")] <- conf.dist
df.new[c("lwr.pred", "upr.pred")] <- pred.dist[,2:3]
p +   
  geom_ribbon(data=df.new, aes(x=rs, ymin=lwr.pred, ymax=upr.pred), alpha=0.1, inherit.aes=F, fill="blue") +
    geom_line(data=df.new, aes(x=rs, y=fit), colour="#339900", size=1)

+ 
  geom_ribbon(data=df.new, aes(x=speed, ymin=lwr.conf, ymax=upr.conf), alpha=0.2, inherit.aes=F, fill="#339900") +

p <- ggplot(d) + geom_point(aes(x=sev, y=rs))

pl <- ggplot(cars) + geom_point(aes(x=speed, y=dist), size=2, colour="#993399") + 
  xlab("Speed (mph)") + ylab("Stopping distance (ft)")

offset <- round(min(d$rs),2)

d$lrs <- log(d$rs + 0.5)
d$lsev <- log(d$sev)

l1 <- lm(d$rs~d$lsev, data=d)
plot(d$rs~d$lsev) 
abline(l1)
summary(l1)

R2 
RSSp <- sum(residuals(e1)^2)

TSS <- sum((d$rs - mean(d$rs))^2)

1 - RSSp/TSS

# R2

r2s <- function(modelo, dataframe, y){

  RSS <- sum(residuals(modelo)^2)
  TSS <- sum(dataframe[,y] - mean())

  out <- data.frame(RSS, TSS)
  return(out)

}

r2s(e1, d, "rs")

RSS

# Adjusted-R2

n <- dim(d)[1]
p <- length(coefficients(e1))

r2adj_pi <- 1 - ((n -1))/(n-p)) * (1- r2.pi)

library("mgcv")
library("rcompanion")
library("gam")

# Modelo 1 gam 
mg <- mgcv::gam(rs~s(sev), data=d, trace = TRUE)
summary(mg)

po <- nls(rs ~ exp.eq(sev, a, b), data = d, start = list(a = .02, b = -.07))

p <- nls(rs ~ a*exp(b*sev), data=d, start = list(a = 0.81, b = 0.021))

plot(d$sev, d$rs, pch=19)
x <- seq(5,25, 0.1)
y <- predict(p, list(sev=x))
lines(x, y)
abline(lm(d$rs~d$sev), col="red")

library("nlstools")

# exploring
plot(d$sev, d$rs, pch=19, xlim=c(0,25), ylim=c(0.2, 1.6))

# try exponential 
f <- as.formula(rs ~ a*exp(b*sev))
preview(f, data = d, start = list(a = 0.8, b=0.019))

nls(f, data = d, start = list(a = 0.75, b=-0.019))

summary(mo)
nd <- expand.grid(rs=seq(0.4,1.6, by=0.1),
                  sev=seq(5,25, by=1))

lines(d$sev, predict(m, nd))

lines(nd$sev, predict(m, nd))

library("FlexParamCurve")

#modpar(posneg.data$age,posneg.data$mass,pn.options = "myoptions.1")

modpar(d$sev, d$rs, pn.options = "myoptions.1")

modcompare <- pn.mod.compare(d$sev, d$rs, d$id, pn.options = "myoptions.1")
modstep <-  pn.modselect.step(d$sev, d$rs, d$id,existing = TRUE, pn.options = "myoptions.1")

posneg.modstep <- pn.modselect.step(posneg.data$age, posneg.data$mass, posneg.data$id, existing = TRUE, pn.options = "myoptions.1")
  #existing = TRUE used to avoid refitting existing nlsList models

f <- as.formula(rs ~ a-b*exp(-c*sev))

preview(f, data=d, start=list(a=-1.6, b=-0.2, c=0.4))

# install nls2
devtools::install_github("ggrothendieck/nls2")

fo <- rs ~ Const + B * (sev ^ A)
st1 <- expand.grid(Const = seq(-2, 2, len = 10),
B = seq(-2,2, len = 10), A = seq(-1, 1, len = 10))
mod1 <- nls2(fo, start = st1, data=d, algorithm = "brute-force")
mod1
# use nls object mod1 just calculated as starting value for
# nls optimization. Same as: nls(fo, start = coef(mod1))
nls2(fo, start = mod1)

summary(mod1)

g <- gam(rs~s(sev), data=d)

plot(g)

library

```

```{r plot_functions}
#https://gist.github.com/ottadini/6882677 
lm_eqn = function(m) {
  # Displays regression line equation and R^2 value on plot
  # Usage:
  # p + annotate("text", x=25, y=300, label=lm_eqn(lm(y ~ x, df)), parse=TRUE)

  l <- list(#a = format(coef(m)[1], digits = 2),
            # b = format(abs(coef(m)[2]), digits = 2),
            r2 = format(summary(m)$r.squared, digits = 3));

  if (coef(m)[2] >= 0)  {
    # eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,l)
    eq <- substitute(~~italic(r)^2~"="~r2,l)
  } else {
    # eq <- substitute(italic(y) == a - b %.% italic(x)*","~~italic(r)^2~"="~r2,l) 
    eq <- substitute(~~italic(r)^2~"="~r2,l)
  }

  as.character(as.expression(eq));                 
}

plotrel <- function(df, v, x, label, coordX_eq, coordY_eq, titulo, ...){
  require(ggplot2)
  require(ggrepel)

  formulita <- as.formula(paste0(v, "_mean_mean ~", x))

  yerrbar <- aes_string(ymin = paste0(v, "_mean_mean - ",  v, "_mean_se"), 
                        ymax = paste0(v, "_mean_mean + ",  v, "_mean_se"))

  ggplot(df, aes_string(y=paste0(v, "_mean_mean"), x=x, label=label)) + 
    geom_smooth(method='lm', se=FALSE, colour="gray") +
    geom_errorbar(mapping = yerrbar) + 
    geom_point(size=1.5, pch = 21, colour="black", fill="white") + 
    theme_bw() + 
    geom_text_repel(nudge_x = 0.5, size = 3) + 
    annotate("text", x=coordX_eq, y = coordY_eq, 
             label=lm_eqn(lm(formulita, df)), parse=TRUE, 
             size = 4.5, fontface = "bold") +
    ggtitle(titulo) + 
    theme(plot.title = element_text(size = 11, face = "bold.italic"),
          panel.grid = element_blank(), 
          axis.text.y = element_text(size=9),
          axis.text.x = element_text(size=9),
          axis.title = element_text(size = 10))
  }

```

```{r plot_rt}
sevRt <- plotrel(sdf, v = "rt", x = 'd_severity', label = 'disturb_year',
                 coordX_eq = 10, coordY_eq = 0.5, titulo = "Rt") +
  xlab('') + ylab('') +
  theme(plot.margin = margin(r=-0.7, l=0, t=0, b=0, "pt"))
```

```{r plot_rs}
sevRs <- plotrel(sdf, v = "rs", x = 'd_severity', label = 'disturb_year',
                 coordX_eq = 10, coordY_eq = 0.5, titulo = "Rs") + 
  xlab('') + ylab('') +
  theme(plot.margin = margin(r=0, l=-0.7, t=0, b=0, "pt"))
```

```{r plot_rc}
sevRc <- plotrel(sdf, v = "rc", x = 'd_severity', label = 'disturb_year',
                 coordX_eq = 10, coordY_eq = 0.5, titulo = "Rc") +
  xlab("Drought Severity") + ylab('') +
  theme(plot.margin = margin(r=-0.7, l=-0.7, t=0, b=0, "pt"))
```

```{r plot_join}
res_long_term <- sevRt + sevRc + sevRs + 
  plot_layout(nrow = 1) 
```

**Figure 6.** Resilience metrics of the tree-growth for the most severe drought events (as from Table S3). *Left*: Resistance (*Rt*); *Center*: Recovery (*Rc*); *Right*: Resilience (*Rs*). Points indicate average of resilience metrics for all populations. Error bar corresponds standard error. Resilience metrics were computed for each population (sample depth > 10) and drought event.

```{r plot_out, fig.width=6, dpi = 300, out.width = 0.8}
res_long_term
```