### 2019-04-10 man/figures/plot_spei12.Rmd
update plot spei12

---
output: 
  word_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```

```{r load_pkg}
library("tidyverse")
library("here")
```

```{r data}
# Get file names
myfiles <- list.files(path = here::here("/data_raw/spei_all/"), pattern = "\\.csv$")

mydf <- data.frame()

# Loop to read files and prepare data
for (j in myfiles){ 
  aux <- read.csv(file=paste0(here::here("/data_raw/spei_all/"), j), header = TRUE, sep = ',')

  # Remove csv and get name 
  name_aux <- str_replace(j, ".csv", "")

  # Get lat long
  latlong <- str_replace(name_aux, "SPEI_", "")

  mylat <- as.numeric(str_split_fixed(latlong, "_", 2))[1]
  mylong <- as.numeric(str_split_fixed(latlong, "_", 2))[2]

  aux$lat <- mylat
  aux$long <- mylong

  # Split time 
  # Get year 
  aux$year <- as.numeric(str_extract(aux$DATA, "[0-9]+"))

  # Get month 
  aux$month <- str_extract(aux$DATA, "[aA-zZ]+")

  # Variable month in number 
  aux$months <- match(aux$month, month.abb)

  # assign(name_aux, aux)

  mydf <- rbind(mydf, aux)
}

# Select SPEI 12
spei <- mydf %>% select(SPEI_12, year, months) %>% 
  group_by(year,months) %>% 
  summarise(spei = mean(SPEI_12),
            spei_sd = sd(SPEI_12)) %>% 
  mutate(date = lubridate::ymd(paste(year, months, "01")))

spei <- spei %>% 
  mutate(sign = case_when(
    spei > 0 ~ "pos",
    TRUE ~ "neg"))
```

```{r plot_spei}
plot_spei <- spei %>%   
  ggplot(aes(x=date, y=spei, fill = sign)) + 
  geom_bar(stat = "identity", colour = "transparent") + 
  scale_fill_manual(values = c("pos" = "#696969", "neg" = "brown4")) +
  theme_bw() + 
  geom_hline(yintercept = 0, colour = 'gray') +
  labs(x = 'Year', y ='SPEI (12-month scale)') +
  scale_x_date(date_breaks = "5 year", date_labels = "%Y", 
               limits=as.Date(c("1949-01-01", "2015-12-01"))) +
  theme(strip.background = element_rect(fill = "white"),
        legend.position = "none", 
        panel.grid = element_blank(), 
        axis.text.y = element_text(size=10),
        axis.text.x = element_text(size=9),
        axis.title.y = element_text(size = 12)) +
  geom_vline(xintercept = as.Date(c("2005-01-01", "2012-01-01")), colour = "grey", alpha = 0.8)

```

```{r plot_out, fig.width=6, dpi = 300, out.width = 0.8} 
# plot_spei
```

```{r}

# DATOS desde 1901

# Get file names
myfiles <- list.files(path = here("data_raw/spei/spei_longterm"), pattern = "\\.csv$")

mydf <- data.frame()

for (j in myfiles){ 
  aux <- read.csv(file=here("data_raw/spei/spei_longterm", j), header = TRUE, sep = ',')

  # Remove csv and get name 
  name_aux <- str_replace(j, ".csv", "")

  # Get lat long
  latlong <- str_replace(name_aux, "spei_", "")
  mylong <- as.numeric(str_split_fixed(latlong, "_", 2))[1]
  mylat <- as.numeric(str_split_fixed(latlong, "_", 2))[2]

  # Store lat and long 
  aux$lat <- mylat
  aux$long <- mylong

  # year and month 
  aux$month <- rep(1:12, 115)

  auxy <- c()
  years <- c() 
  for (j in 1901:2015) { 
    auxy <- as.data.frame(rep(j,12))
    years <- rbind(years, auxy)
    } 
  names(years) <- 'year'

  aux <- cbind(aux,years)

  # join all df 
  mydf <- rbind(mydf, aux)
}

# Convert to date  
mydf <- mydf %>% mutate(date = sprintf("%d-%02d", year, month),
                        date = as.Date(paste0(date, '-01'), format="%Y-%m-%d"))

speidf <- mydf %>% dplyr::select(SPEI06,SPEI12,SPEI24,month, year, date) %>% 
  group_by(year,month, date) %>%
  summarise_at(vars(SPEI06,SPEI12,SPEI24), funs(mean,sd,se=sd(.)/sqrt(n()))) %>% 
  as.data.frame() 
```

```{r}
speiPlot <- function(df, vi, mytitle, year_scaleX){
  require(dplyr)
  require(scales)

  # Select variable of interest 
  dfs <- df %>% dplyr::select(year, month, date, starts_with(vi))

  # Rename variables 
  dfs <- dfs %>% rename_all(~sub(paste0(vi,'_'), '',.x))
  # Create variable pos and neg 
  dfs <- dfs %>% mutate(signo = ifelse(mean > 0, "pos", "neg"))

  # Plot  
  p <- ggplot(dfs, aes(x = date, y = mean, fill = signo)) + 
    geom_rect(aes(xmin=as.Date("2005-01-01"), xmax = as.Date("2005-12-01"), ymin=-Inf, ymax=Inf),
              fill="grey", alpha=.9) + 
    geom_rect(aes(xmin=as.Date("2012-01-01"), xmax = as.Date("2012-12-01"), ymin=-Inf, ymax=Inf),
              fill="grey", alpha=.9) + 
    geom_bar(stat = "identity") +
    scale_fill_manual(values = c("pos" = "darkblue", "neg" = "red")) +
    theme_bw()  + 
    labs(title= mytitle, x= '', 
         y = paste0(substr(vi, 1, 4), ' (', gsub("[^0-9]","",vi),'-month scale)')) +
    theme(strip.background = element_rect(fill = "white"),
        legend.position = "none", 
        panel.grid = element_blank(), 
        axis.text.y = element_text(size=10),
        axis.text.x = element_text(size=9),
        axis.title.y = element_text(size = 12))

  # Date Scale in X 
  if (missing(year_scaleX)){
    p
    } else { 
    year_interval <- paste0(year_scaleX," year")
    p <- p + scale_x_date(date_breaks = year_interval, date_labels = "%Y",
                          limits=as.Date(c("1901-01-01", "2015-12-01")))
    }

  return(p)
}

s <- speiPlot(speidf, vi = 'SPEI12', mytitle = "", year_scaleX = 10)
```

**Figure S2**. Drought severity in the Sierra Nevada for the 1901-2016 period based on the Standardised Precipitation-Evapotranspiration Index (SPEI). Data from Global SPEI database (http://spei.csic.es/database.html). We obtanied the SPEI data for a 12 month scale and for all 0.5º grid cells covering Sierra Nevada. Horizontal gray bars indicate 2005 and 2012 year.

```{r plot_out1900, fig.width=6, dpi = 300, out.width = 0.8} 
s
```