### 2019-07-10 man/tables/t_forest_history.Rmd

---
output:
  word_document:
    reference_docx: ../../templates/template_tables_portrait.docx
bibliography: ../../refs/references.bib
csl: ../../refs/ecosystems.csl
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
library(here)
library(pander)
library(kableExtra)
library(knitr)
library(dplyr)
library(flextable)
library(officer)
library(tidyverse)
```

**Table S4.** Review of the forest and management history of the samplin site. An exhaustive review of historical documents was done to compile information about socio-economical activities affecting forest. Several documentary sources were reviewed: historical documents and maps [*e.g.* @Titos1990]; detailed mining reports [*e.g.* @Maestre1858]; official information about recent wildfires events and forest management practices [@Bonet2016obsnev_forest]; livestock farming [*e.g.* @MorenoLlorca2016]; traditional irrigation ditches [@RuizRuiz2017] and other studies reviewing the socioeconomic dynamics of forest of Sierra Nevada at different scales [@JimenezOlivencia2015; @MorenoLlorca2016].

[@Catastro1752; @Lopez1776; @Madoz1846; @Maestre1852; @Maestre1858; @Cruz1991; @Titos1990; @Titos1997; @Ferrer1999; @MesaTorres2009; @Bonet2014_conama; @Navarro2014; @MartinCivantos2014; @MartinMontanes2015; @JimenezOlivencia2015; @2015Zoido; @Wing2015; @Bonet2016obsnev_forest; @MorenoLlorca2016; @RuizRuiz2017; @CMA2018; @Calatrava2019; J. Navarro and F.J. Cano-Manuel *personal communications*]

@Catastro1752; @MesaTorres2009; @Bonet2014_conama
@Bonet2014_conama; @MorenoLlorca2016; @CMA2018
@Bonet2016obsnev_forest; @MorenoLlorca2016; J. Navarro and F.J. Cano-Manuel *personal communications*
@Cruz1991; @Wing2015
@JimenezOlivencia2015; @2015Zoido; @MorenoLlorca2016; @Calatrava2019
@Maestre1852; @Maestre1858; @Titos1990; @MesaTorres2009
@Navarro2014
@MartinCivantos2014; @MartinMontanes2015; @RuizRuiz2017
@Catastro1752; @Lopez1776; @Madoz1846; @Titos1997; @Ferrer1999; @MesaTorres2009; @Bonet2014_conama

```{r}
u_raw <- read_delim(here::here("/data_raw/review_usos.tsv"), delim = "\t") %>% 
  dplyr::select(-ref) %>% spread(Site, item)

u <- regulartable(u_raw, 
                   col_keys = c("activity", "CA", "SJ", "refall"))

tabla <- u %>% 
  set_header_labels(activity = "", CA = "Cáñar (CA sites)", SJ = "Güejar-Sierra (SJ site)", refall = "References") %>% 
  bold(part = "header") %>% 
  bold(j = 1) %>% 
  border_outer(part = "all", border = fp_border(color = "black", width = 1.5)) %>% 
  border_inner_h(part="all", border = fp_border(color = "gray")) %>% 
  border_inner_v(part="all", border = fp_border(color = "gray")) %>% 
  fontsize(part = "header", size = 10) %>% 
  fontsize(part = "body", size = 9) %>% 
  autofit(add_w = 0, add_h = 0)

tabla 
```