### 2019-08-14 man/figures/resilience_metrics.Rmd
update new version

---
output: 
  word_document
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```

```{r load_pkg}
library("tidyverse")
library("here")
library("ggrepel")
library("grid")
library("gridExtra")
library("patchwork")
library("lemon")
library("bookdown")
source(here::here('/script/R/custom_functions.R'))
```

```{r data}
huber_evi <- 
  read.csv(here::here("/out/anovas_resilience/huber_evi", "robust_mhuber.csv"), header = TRUE) %>% 
  filter(var != "rrs") %>% 
  mutate(site_sorted = as.factor(
  case_when(
  site == "N" ~ "0_N",
  site == "S" ~ "1_S"
)))

huber_b_evi <- 
  read.csv(here::here("/out/anovas_resilience/huber_evi", "robust_mhuber_b.csv"), header = TRUE) %>%
  filter(var != "rrs") %>% 
  mutate(var_sorted = as.factor(
    case_when(
      var == "rt" ~ "0_rt",
      var == "rc" ~ "1_rc",
      var == "rs" ~ "2_rs",
))) %>% 
mutate(site_sorted = as.factor(
  case_when(
  site == "N" ~ "0_N",
  site == "S" ~ "1_S"
)))

huber_bai <- read.csv(here::here("/out/anovas_resilience/huber_bai", "robust_mhuber.csv"), header = TRUE) %>%
  filter(var != "rrs") %>% 
  mutate(site_sorted = as.factor(
  case_when(
  site == "SJ" ~ "0_SJ",
  site == "caH" ~ "2_CA-High",
  site == "caL" ~ "1_CA-Low"
)))

huber_b_bai <- 
  read.csv(here::here("/out/anovas_resilience/huber_bai", "robust_mhuber_b.csv"), header = TRUE) %>%
  filter(var != "rrs") %>% 
  mutate(var_sorted = as.factor(
    case_when(
      var == "rt" ~ "0_rt",
      var == "rc" ~ "1_rc",
      var == "rs" ~ "2_rs",
))) %>% 
  mutate(site_sorted = as.factor(
  case_when(
  site == "SJ" ~ "0_SJ",
  site == "caH" ~ "2_CA-High",
  site == "caL" ~ "1_CA-Low"
)))
```

```{r}
# plot_params 
pd <- position_dodge(width = 1)
aes_main <- aes(x=site_sorted, y = M.Huber, 
                ymin = lower.ci, ymax = upper.ci,
                fill = site_sorted, colour = site_sorted)
aes_aux <- aes(shape = as.factor(disturb_year), 
               label = Letter)
labelita <- as_labeller(c("0_rt" = "Resistance",
                                      "1_rc" = "Recovery", 
                                      "2_rs" = "Resilience"))

item_NS <- c("0_N" = "N", "1_S" = "S")
item_site <- c("0_SJ" = "SJ", "2_CA-High" = "CA-High", "1_CA-Low" = "CA-Low")

theme_res <- list(
  geom_errorbar(width=.1, position = pd),
  geom_point(size=7, shape=21),
  geom_text(aes(label = str_to_upper(Letter)), fontface = "bold", colour = "white"), 
  theme_bw(), xlab(""), ylab (""), 
  theme(panel.grid = element_blank(), 
        axis.text = element_text(size=9),
        legend.title = element_blank(),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(size = 11, face = "bold.italic"), 
        plot.title = element_text(size = 12, face = "bold")),
  scale_shape_manual(values = c(24,22)),
  facet_wrap(~var_sorted, scales = "free_y", labeller = labelita))

# PLOT EVI 
pevi <- ggplot(huber_b_evi, aes_main) + 
  # by disturb year 
  geom_errorbar(data=huber_evi, aes_aux, width=.1, position = pd) +
  geom_point(data=huber_evi, aes_aux, position=pd, size = 4.5, fill="white") + 
  geom_text(data = huber_evi, aes_aux, position = pd, size = 2.5) +
  scale_x_discrete(labels = item_NS) +
  scale_fill_manual(values = colores_den[3:2], labels = item_NS) +
  scale_colour_manual(values = colores_den[3:2], labels = item_NS) + theme_res +
  theme(legend.position = "top") + ggtitle ("EVI")

# PLOT BAI
pbai <- ggplot(huber_b_bai, aes_main) + 
  # by disturb year 
  geom_errorbar(data=huber_bai, aes_aux, width=.1, position = pd) +
  geom_point(data=huber_bai, aes_aux, position=pd, size = 4.5, fill="white") + 
  geom_text(data = huber_bai, aes_aux, position = pd, size = 2.5) +
  scale_x_discrete(labels = item_site) + 
  scale_fill_manual(values = rev(colores_den), labels = item_site) +
  scale_colour_manual(values = rev(colores_den), labels = item_site) + 
  theme_res + theme(legend.position = "bottom") + 
  ggtitle("BAI")

plot_res <- pevi + pbai + plot_layout(ncol=1, tag_level = "keep") + plot_annotation(tag_levels = "a") & theme(plot.tag = element_text(size = 12, face = 2)) & theme(plot.margin = margin(t=0,r=0,b=-2,l=0, "pt"))
```

**Figure 7.** Comparison of the response of  *Q. pyrenaica* forests to drought in terms of resistance, recovery and resilience of greenness (a) and tree-growth (b). For EVI we compared northern populations (*red circle*) with southern ones (*green circle*). For BAI we compared the more xeric northern population (San Juan, SJ; *red circle*) with the two southerns populations: Cáñar-High (CA-High; *blue circle*) and Cáñar-Low (CA-Low; *green circle*). Different letters indicate significant *post hoc* differences between groups (see material and methods for details).

```{r plot_out, fig.height=5, dpi = 300, out.width = 0.8} 
plot_res 
```

```{r}
# SAVE as eps 
saveImage(plot_res, path="./man/figures/svg/", dpi=300, w=22, h=25, fig= "fig_7", u = "cm") 
```

```{r, eval=FALSE, echo = FALSE}
# old plots

mhuberbai <- read.csv(here::here("/out/anovas_resilience/huber_bai", "robust_mhuber.csv"), header = TRUE)

# NO rrs
mhuber_filter <- mhuber %>% filter(var != 'rrs') %>% mutate(disturb_year = as.factor(disturb_year))
mhuberbai_filter <- mhuberbai %>% filter(var != 'rrs') %>% mutate(disturb_year = as.factor(disturb_year))

aux <- mhuber_filter %>% mutate(vari = '0_EVI')
aux2 <- mhuberbai_filter %>% mutate(vari = '1_BAI')
m <- bind_rows(aux2, aux)

pd <- position_dodge(width = 0.6)

aes_disturb <- aes(x = disturb_year, y = M.Huber, 
                   color = site_sorted, group = site_sorted,  
                   fill = site_sorted, linetype=site_sorted,
                   shape = site_sorted, 
                   ymin = lower.ci, ymax = upper.ci)

variables  <- c("0_EVI" = "EVI", "1_BAI" = "BAI")

# orden <- c("N", "S", "SJ", "CA-High","CA-Low")
item_site <- c("0_N" = "N", "1_S" = "S", "2_SJ" = "SJ", "3_CA-High" = "CA-High", "4_CA-Low" = "CA-Low")

tl <- guide_legend(title = "Site")

themeR <- list(
  geom_errorbar(width=.1, position = pd),
  geom_point(position = pd),
  theme_bw(),
  facet_wrap(~vari, labeller = labeller(vari = variables), scales="free"),
  xlab(''), 
  theme(
    panel.grid.minor = element_blank(),
    strip.background = element_rect(
      colour = "black",
      fill = "white"),
    strip.text = element_text(face="bold", size = 10),
    axis.title.y = element_text(face="bold", size = 10),
    plot.title = element_text(hjust = -0.05, size = 10, margin = margin(t=-5, b=1)),
    axis.text = element_text(size = 9),
    panel.spacing.x = unit(1, "lines")),
  scale_shape_manual(labels = item_site, guide = guide_legend(title = "Site"),
    values=c("0_N" = 19, "1_S" = 21, "3_CA-High" = 22, "4_CA-Low" = 15, "2_SJ" = 17)), 
  scale_color_manual(labels = item_site, guide = tl,
    values=c("0_N" = 'black', "1_S" = "blue", "3_CA-High" = 'blue', "4_CA-Low" = 'blue', "2_SJ" = 'black')),
  scale_fill_manual(labels = item_site, guide = tl,
    values=c("0_N" = 'black', "1_S" = "white", "3_CA-High" = 'white', "4_CA-Low" = 'blue', "2_SJ" = 'black')),
  scale_linetype_manual(labels = item_site, guide = tl,
                        values=c("solid","solid","solid","solid", "solid")),
  scale_size_manual(labels = item_site, guide = tl,
    values=c("0_N" = "2", "1_S" = "2", "3_CA-High" = 2.5, "4_CA-Low" = 2.5, "2_SJ" = 2.5)),

  geom_text(aes(label=Letter, y=postext), 
            position=pd, 
            show.legend = FALSE, size = 3)
  # , position_jitterdodge(dodge.width = 0.6, jitter.width = .2)
  # geom_text_repel( 
  #                 nudge_x = - 0.5,
  #                 nudge_y = 0.01,
  #                 direction="both", 
  #                 size = 3.5, show.legend = FALSE,
  #                 segment.size = 0.5, 
  #                 segment.alpha = 0.5) 
)

m <- m %>% mutate(postext = case_when(
  vari == "0_EVI" ~
    case_when(var_sorted == "0_rt" ~ 0.97,
              var_sorted == "1_rc" ~ 1.19,
              var_sorted == "2_rs" ~ 1.02),
  vari == "1_BAI" ~
     case_when(var_sorted == "0_rt" ~ 1.1,
              var_sorted == "1_rc" ~ 1.7,
              var_sorted == "2_rs" ~ 1.2)))

m <- m %>% mutate(site = as.factor(
  case_when(
  site == "caH" ~ "CA-High",
  site == "caL" ~ "CA-Low",
  TRUE ~ site
)))

m <- m %>% mutate(site_sorted = as.factor(
  case_when(
  site == "N" ~ "0_N",
  site == "S" ~ "1_S",
  site == "SJ" ~ "2_SJ",
  site == "CA-High" ~ "3_CA-High",
  site == "CA-Low" ~ "4_CA-Low"
)))

rt <- m %>% filter(var == "rt") %>% 
  ggplot(aes_disturb) + themeR + 
  ylab("Resistance") +
  theme(plot.margin = margin(r=0, l=0, t=0, b=-2, "pt"))
# + geom_text(aes(label=Letter), position=position_jitterdodge(jitter.width = .2, dodge.width = 0.6), show.legend = FALSE)

rc <- m %>% filter(var == "rc") %>% 
  ggplot(aes_disturb) + themeR + 
  ylab("Recovery") +
  theme(plot.margin = margin(r=0, l=0, t=0, b=-2, "pt"))
# + geom_text(aes(label=Letter, y = 1.65), position=pd, show.legend = FALSE)

rs <- m %>% filter(var == "rs") %>% 
  ggplot(aes_disturb) + themeR + 
  ylab("Resilience") + xlab('Drought event') +
    theme(plot.margin = margin(r=0, l=0, t=0, b=-2, "pt"))
# + geom_text(aes(label=Letter, y = 1.2), position=pd, show.legend = FALSE)

# ```{r plot_out, fig.height=5, dpi = 300, out.width = 0.8} 
# grid_arrange_shared_legend(rt, rc, rs, ncol = 1, nrow = 3, position ='top')

# ```

```

```{r old1, eval=FALSE}

theme_de <- list(
  scale_color_manual(values=c('black','red')), 
  scale_shape_manual(values=c(19, 15)),
  scale_fill_manual(values=c('black','red')),  
  scale_linetype_manual(values=c("solid", "solid")),
  labs(color = 'Drougth event', fill='Drougth event', linetype='Drougth event', shape='Drougth event')

)

# EVI_site
res_evi_site <-
  ggplot(mhuber_filter, aes_site) + 
  theme_res + theme_de + theme(legend.position='top') + 
  ggtitle("a) EVI")

# BAI_site
res_bai_site <-
  ggplot(mhuberbai_filter, aes_site) + 
  theme_res + theme_de + theme(legend.position='bottom') + 
  xlab('Site') + 
  ggtitle("b) BAI")

res_evi_site  + res_bai_site + plot_layout(ncol=1)

# **Figure .** Response of *Q. pyrenaica* forests to 2005 (*black*) and 2012 (*red*) drought events in terms of resistance, recovery and resilience of greenness (EVI; a) and tree radial growth (BAI; b) by site. Different letters above bars indicate significant post hoc differences between groups (see material and methods). 
```

```{r old2, eval=FALSE, echo=FALSE, fig.height=7}
mhuber <- read.csv(here("/out/anovas_resilience/huber_evi", "robust_mhuber.csv"), header = TRUE)
mhuberbai <- read.csv(here("/out/anovas_resilience/huber_bai", "robust_mhuber.csv"), header = TRUE)

# NO plot rrs 
mhuber_filter <- mhuber %>% filter(var != 'rrs') %>% mutate(disturb_year = as.factor(disturb_year))
mhuberbai_filter <- mhuberbai %>% filter(var != 'rrs') %>% mutate(disturb_year = as.factor(disturb_year))

pd <- position_dodge(.2)
line_lwd <- .8

aes_disturb <- aes(x = disturb_year, y = M.Huber, 
                   color = site, group = site, 
                   fill = site, linetype=site,
                   shape = site, 
                   ymin = lower.ci, ymax = upper.ci)
aes_site <- aes(x = site, y = M.Huber, 
                color = disturb_year, group=disturb_year, 
                fill=disturb_year, linetype=disturb_year,
                shape = disturb_year, 
                ymin = lower.ci, ymax = upper.ci)

theme_res <- list(
  geom_errorbar(width=.2, position=pd, lwd=line_lwd), 
  geom_line(position = pd, lwd=line_lwd),
  geom_point(size = 2.5, position = pd), 
  facet_wrap(
    ~var_sorted, nrow = 1, scales="free_y",
    labeller = as_labeller(c(
      "0_rt" = "Resistance",
      "1_rc" = "Recovery",
      "2_rs" = "Resilience"))),
  geom_text_repel(aes(label=Letter), 
                  nudge_x = - 0.5,
                  nudge_y = 0.01,
                  direction="both", 
                  size = 3.5, show.legend = FALSE,
                  segment.size = 0.5, 
                  segment.alpha = 0.5),
  theme_bw(), 
  theme(
    panel.grid.minor = element_blank(),
    strip.background = element_rect(
      colour = "black",
      fill = "white"),
    plot.title = element_text(hjust = -0.05, size = 16, margin = margin(t=-5, b=1))),  
  ylab("Response value"), 
  xlab("")
)

theme_evi <-  list(
  scale_color_manual(values = c("black", "blue")),
  scale_shape_manual(values=c(19, 22)),
  scale_fill_manual(values = c("black", "blue")),
  scale_linetype_manual(values = c("solid", "solid"))
)

theme_bai <- list(
  scale_color_manual(values=c('black','blue','blue')), 
  scale_shape_manual(values=c(19, 22, 15)),
  scale_fill_manual(values=c('black','white','blue')),  
  scale_linetype_manual(values=c("solid", "solid", 'solid'))
)

# EVI_drought 
res_evi_drought <-
  ggplot(mhuber_filter, aes_disturb) + 
  theme_res + theme_evi + theme(legend.position='top') + 
  ggtitle("a) EVI")

# BAI_drought 
res_bai_drought <-
  ggplot(mhuberbai_filter, aes_disturb) + 
  theme_res + theme_bai + theme(legend.position='bottom') + 
  xlab('Drought event') + 
  ggtitle("b) BAI")

res_evi_drought  + res_bai_drought + plot_layout(ncol=1)
```