### 2020-05-24 man/ecosystems/2_revision2/fig/aux/bai_vs_rwi.Rmd
update figures

---
output: 
  word_document
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```

```{r load_pkg}
library("tidyverse")
library("here")
library("ggrepel")
library("grid")
library("gridExtra")
library("patchwork")
library("lemon")
library("bookdown")
source(here::here('/script/R/custom_functions.R'))
```

```{r data}
huber_bai <- read.csv(here::here("/out/anovas_resilience/huber_bai", "robust_mhuber.csv"), header = TRUE) %>%
  filter(var != "rrs") %>% 
  mutate(site_sorted = as.factor(
  case_when(
  site == "SJ" ~ "0_SJ",
  site == "caH" ~ "2_CA-High",
  site == "caL" ~ "1_CA-Low"
)))

huber_b_bai <- 
  read.csv(here::here("/out/anovas_resilience/huber_bai", "robust_mhuber_b.csv"), header = TRUE) %>%
  filter(var != "rrs") %>% 
  mutate(var_sorted = as.factor(
    case_when(
      var == "rt" ~ "0_rt",
      var == "rc" ~ "1_rc",
      var == "rs" ~ "2_rs",
))) %>% 
  mutate(site_sorted = as.factor(
  case_when(
  site == "SJ" ~ "0_SJ",
  site == "caH" ~ "2_CA-High",
  site == "caL" ~ "1_CA-Low"
)))

huber_rwi <- read.csv(here::here("/out/anovas_resilience/huber_rwi", "robust_mhuber.csv"), header = TRUE) %>%
  filter(var != "rrs") %>% 
  mutate(site_sorted = as.factor(
  case_when(
  site == "SJ" ~ "0_SJ",
  site == "caH" ~ "2_CA-High",
  site == "caL" ~ "1_CA-Low"
)))

huber_b_rwi <- 
  read.csv(here::here("/out/anovas_resilience/huber_rwi", "robust_mhuber_b.csv"), header = TRUE) %>%
  filter(var != "rrs") %>% 
  mutate(var_sorted = as.factor(
    case_when(
      var == "rt" ~ "0_rt",
      var == "rc" ~ "1_rc",
      var == "rs" ~ "2_rs",
))) %>% 
  mutate(site_sorted = as.factor(
  case_when(
  site == "SJ" ~ "0_SJ",
  site == "caH" ~ "2_CA-High",
  site == "caL" ~ "1_CA-Low"
)))
```

```{r}
df <- huber_bai %>% mutate(growth_var = "bai") %>% 
  bind_rows(huber_rwi %>% mutate(growth_var = "rwi"))
```

```{r}
# plot_params 
labelita <- c("0_rt" = "Resistance",
                                      "1_rc" = "Recovery", 
                                      "2_rs" = "Resilience")
item_site <- c("0_SJ" = "SJ", "2_CA-High" = "CA-High", "1_CA-Low" = "CA-Low")

pd <- position_dodge(width = .8)

ggplot(df, aes(x=site_sorted, y = M.Huber, 
               colour=as.factor(growth_var), 
               fill=as.factor(growth_var),
               ymin = lower.ci, 
               ymax = upper.ci)) + 
  geom_point(size=4, position = pd) + 
  facet_grid(disturb_year~var_sorted, 
             labeller = labeller(var_sorted = labelita)) +
  scale_fill_manual(values = c("black", "gray")) + 
  scale_color_manual(values = c("black", "gray")) + 
  theme(panel.grid = element_blank(), 
        axis.text = element_text(size=9),
        legend.title = element_blank(),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(size = 11, face = "bold.italic"), 
        plot.title = element_text(size = 12, face = "bold")) +
   theme_bw() + 
  geom_errorbar(width=.1, position = pd) + 
  xlab("") + ylab ("") + 
  theme(panel.grid = element_blank(), 
        axis.text = element_text(size=9),
        legend.title = element_blank(),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(size = 11, face = "bold.italic"), 
        plot.title = element_text(size = 12, face = "bold")) +
  scale_x_discrete(labels = item_site)

```

```{r}

dfres <- df %>% filter(var == "rs")

ggplot(dfres, aes(x=site_sorted, y = M.Huber, 
               colour=as.factor(growth_var), 
               fill=as.factor(growth_var),
               ymin = lower.ci, 
               ymax = upper.ci)) + 
  geom_point(size=4, position = pd) + 
  facet_wrap(~disturb_year) +
  scale_fill_manual(values = c("black", "gray")) + 
  scale_color_manual(values = c("black", "gray")) + 
  theme(panel.grid = element_blank(), 
        axis.text = element_text(size=9),
        legend.title = element_blank(),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(size = 11, face = "bold.italic"), 
        plot.title = element_text(size = 12, face = "bold")) +
   theme_bw() + 
  geom_errorbar(width=.1, position = pd) + 
  xlab("") + ylab ("Resilience") + 
  theme(panel.grid = element_blank(), 
        axis.text = element_text(size=9),
        legend.title = element_blank(),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(size = 11, face = "bold.italic"), 
        plot.title = element_text(size = 12, face = "bold")) +
  scale_x_discrete(labels = item_site)
```