### 2020-05-24 man/figures/std_anomalies.Rmd

---
output:
  word_document
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```

```{r load_pkg}
library("tidyverse")
library("here")
library("scales")
library("patchwork")
source(here::here('/script/R/custom_functions.R'))
```

**Figure 2.** EVI standardized anomaly during the period 2000-2016 for northern and southern populations. Error bars show standard error. See main text for details on EVI calculation.

```{r data}
# From explore_anomalies.Rmd 
# Read data 
anomalias_evimean <- read.csv(here("data/anomalies", "anomalias_evimean.csv"),header = TRUE, sep = ',')

anomalias_evimean <- anomalias_evimean %>% 
  mutate(
    clu_pop = as.factor(case_when(
      pop == 1 ~ "Camarate",
      pop %in% c(2,3,4,5) ~ 'Northern slope',
      pop %in% c(6,7,8) ~ 'Southern slope',
      pop == 9 ~ 'out')),
    clu_pop2 = as.factor(case_when(
      pop %in% c(1,2,3,4,5) ~ 'Northern slope',
      pop %in% c(6,7,8) ~ 'Southern slope',
      pop == 9 ~ 'out'))) %>% 
  filter(clu_pop != 'out')

# Standardized Anomalies 
avg_sa_clu <- anomalias_evimean %>% 
  group_by(clu_pop2, y) %>% 
  dplyr::summarise(mean=mean(sa, na.rm=T),
            sd = sd(sa, na.rm=T),
            se = sd/sqrt(length(sa))) %>% 
 mutate(signo = ifelse(mean >= 0, 'pos', 'neg'))
```

```{r plot}
# colours 
color_neg <- '#a63603'
color_pos <- '#006d2c'

myylab <- 'EVI Standardized Anomaly'

plot_sa_clu  <- avg_sa_clu %>%  
  ggplot(aes(x=y, y=mean, fill=signo)) + 
  geom_hline(yintercept = 0, size=0.3, colour = "gray") + 
  geom_bar(stat = "identity") + 
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se, colour=signo), width=.5) +
  facet_wrap(~clu_pop2, nrow = 2) + 
  scale_fill_manual(values = c("pos" = color_pos, "neg" =  color_neg)) +
  scale_color_manual(values = c("pos" = color_pos, "neg" =  color_neg)) +
  scale_x_continuous(breaks = seq(2000, 2016, by=2)) +
  ylab(myylab) + xlab('Year') +
  theme_bw() +
  theme(axis.text = element_text(size=8),
        legend.position = "none",
        panel.grid = element_blank(),
        strip.background = element_rect(fill = "transparent", colour = "black", size = 0.2),
        plot.title = element_text(hjust = -0.05, size = 9, 
                                    margin = margin(t=-5, b=1)))
```

```{r out_plot, fig.width= 4, dpi = 300, out.width = 0.8}
plot_sa_clu 
```

```{r}
saveImage(plot_sa_clu, path="./man/figures/svg/", dpi=300, w=14, h=14, fig= "fig_2", u = "cm") 
```

```{r}
## Reviewer 1. Is it necessary/correct to remove year i from the calculation of EVImean, ref?

# From explore_anomalies.Rmd 
# Read data 
anomalias_evimeanc <- read.csv(here("data/anomalies", "anomalias_evimean_corrected.csv"),header = TRUE, sep = ',')

anomalias_evimeanc <- anomalias_evimeanc %>% 
  mutate(
    clu_pop = as.factor(case_when(
      pop == 1 ~ "Camarate",
      pop %in% c(2,3,4,5) ~ 'Northern slope',
      pop %in% c(6,7,8) ~ 'Southern slope',
      pop == 9 ~ 'out')),
    clu_pop2 = as.factor(case_when(
      pop %in% c(1,2,3,4,5) ~ 'Northern slope',
      pop %in% c(6,7,8) ~ 'Southern slope',
      pop == 9 ~ 'out'))) %>% 
  filter(clu_pop != 'out')

# Standardized Anomalies 
avg_sa_cluc <- anomalias_evimeanc %>% 
  group_by(clu_pop2, y) %>% 
  dplyr::summarise(mean=mean(sa, na.rm=T),
            sd = sd(sa, na.rm=T),
            se = sd/sqrt(length(sa))) %>% 
 mutate(signo = ifelse(mean >= 0, 'pos', 'neg'))

a <- avg_sa_clu %>% 
  mutate(method = "original") %>% 
  bind_rows(avg_sa_cluc %>% mutate(method = "no remove")) %>% 
  unite("f", signo:method, remove = FALSE)

comparative_anomalies <- a %>% ggplot(aes(x=y, y=mean, fill=f)) + 
  geom_hline(yintercept = 0, size=0.3, colour = "gray") + 
  geom_bar(position = "dodge", stat = "identity", width = 0.8) + geom_errorbar(aes(ymin = mean - se, ymax = mean + se, colour = f), 
                width=.5, position = position_dodge(0.9)) +
  facet_wrap(~clu_pop2, nrow = 2) + 
  scale_fill_manual(values = c("neg_original" =  "#a63603",
                               "neg_no remove" =  "#FFCCCC",
                               "pos_original" = "#006d2c", 
                               "pos_no remove" = "#CCFFCC"
                                 )) +
  scale_colour_manual(values = c("neg_original" =  "#a63603",
                               "neg_no remove" =  "#FFCCCC",
                               "pos_original" = "#006d2c", 
                               "pos_no remove" = "#CCFFCC"
                                 )) +
  scale_x_continuous(breaks = seq(2000, 2016, by=2)) +
  ylab('EVI Standardized Anomaly') + xlab('Year') +
  theme_bw() +
  theme(axis.text = element_text(size=8),
        legend.position = "none",
        panel.grid = element_blank(),
        strip.background = element_rect(fill = "transparent", colour = "black", size = 0.2),
        plot.title = element_text(hjust = -0.05, size = 9, 
                                    margin = margin(t=-5, b=1)))
```

```{r oldFigure, eval=FALSE, fig.height=7}
## See profile_EVI.Rmd 
# grid.arrange(profile_compara, plot_sa_clu)

a_profile_compara <- profile_compara + labs(title = "a")
b_plot_sa_clu <- plot_sa_clu + labs(title = "b")
a_profile_compara + b_plot_sa_clu + plot_layout(ncol=1)
```