### 2020-06-07 man/ecosystems/2_revision2/fig/aux/comparison_sites.Rmd

---
title: "spei_locations"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(here)
library(SPEI)
library(lubridate)
source(here::here('script/R', 'hydro_year.R'))
source(here::here('/script/R/custom_functions.R'))
```

```{r}
# SJ

prec_files <- list.files(here::here('data_raw/meteo/prec_subclima/'), pattern="*.csv")

dfpre <- c()

for (i in prec_files){

  auxdf <- read.csv(here::here('data_raw/meteo/prec_subclima', i), header = TRUE, sep=";")
  auxdf$code <- gsub(x=i, pattern = ".csv", replacement = "")

  dfpre <- rbind(dfpre, auxdf)
}

dfpre$date <- as.Date(dfpre$date, format = "%d/%m/%Y")

dfpre <- dfpre %>% 
  dplyr::mutate(fecha = as.Date(date, format='%d/%m/%Y'),
         month = lubridate::month(fecha),
         hmonth = ifelse(month <= 8, month + 4, month -8), 
         hyear = hydro_year(fecha),
         hyear_f = paste0(hyear-1,'-', hyear))

aux <- dfpre %>% 
  group_by(hyear_f, code) %>% 
  dplyr::mutate(csum = cumsum(prec)) %>% 
  filter(hmonth == 12) %>% 
  dplyr::select(-month, -prec, -date, -fecha) %>% as.data.frame()

aux_sj <- aux %>% 
  filter(code %in% c("5501","5502","5504","5504A","5505","5509","5509A")) %>% 
  group_by(hyear) %>% 
     dplyr::summarise(
     csum_mean = mean(csum, na.rm = TRUE),
     csum_sd = sd(csum, na.rm = TRUE), 
     csum_se = csum_sd/sqrt(length(csum_sd)), 
     n = n())

## 
guejar <- read_csv("/Users/ajpelu/Downloads/cn_5501_72.csv") %>% 
  filter(variable_id == 29) %>% 
  group_by(mensual=floor_date(time, "month")) %>%
  summarize(prec=sum(value)) %>% 
    mutate(site = "Guejar")

quentar <- read_csv("/Users/ajpelu/Downloads/cn_E42_235.csv") %>% 
  filter(variable_id == 49) %>% 
  group_by(mensual=floor_date(time, "month")) %>%
  summarize(prec=sum(value)) %>% 
  mutate(site = "Quentar")

quentar_pantano <- read_csv("/Users/ajpelu/Downloads/cn_5506_100.csv") %>% 
  filter(variable_id == 29) %>% 
  group_by(mensual=floor_date(time, "month")) %>%
  summarize(prec=sum(value)) %>% 
  mutate(site = "Quentar Pantano")

df <- bind_rows(guejar, quentar, quentar_pantano) %>% 
  filter(prec >= 0) %>% 
  mutate(month = lubridate::month(mensual),
         hmonth = ifelse(month <= 8, month + 4, month -8), 
         hyear = hydro_year(mensual),
         hyear_f = paste0(hyear-1,'-', hyear)) %>% 
  group_by(hyear_f, site) %>% 
  dplyr::mutate(csum = cumsum(prec)) %>% 
  filter(hmonth == 12) %>% 
  dplyr::select(-month, -prec, -mensual) %>% 
  group_by(hyear) %>% 
     dplyr::summarise(
     csum_mean = mean(csum, na.rm = TRUE),
     csum_sd = sd(csum, na.rm = TRUE), 
     csum_se = csum_sd/sqrt(length(csum_sd)), 
     n = n())

# Compare df con aux_sj

# aux_sj <- aux_sj %>% mutate(tipo = "old") 
# df <- df %>% mutate(tipo = "new")
# 
# s <- bind_rows(aux_sj, df)
# 
# 
# s %>% filter(hyear > 1990) %>% 
#   ggplot(aes(x=hyear, y=csum_mean, fill=tipo, colour = tipo)) + 
#   geom_line(position = pos) +
#   scale_x_continuous(breaks=seq(1990,2020, by=5)) +
#   scale_y_continuous(breaks=seq(0,1500, by=250)) +
#   geom_errorbar(aes(ymin = csum_mean - csum_se, ymax= csum_mean + csum_se), width=.1, position = pos) + geom_point(size=2, position = pos) + theme_bw()

# utilizamos el old 
```

# Pra CA sites usamos

```{r}

#6236
ca <- aux %>% 
  filter(code %in% c("6257I","6242U","6236")) %>% 
  group_by(hyear) %>% 
     dplyr::summarise(
     csum_mean = mean(csum, na.rm = TRUE),
     csum_sd = sd(csum, na.rm = TRUE), 
     csum_se = csum_sd/sqrt(length(csum_sd)), 
     n = n()) %>% 
  mutate(site = "CA")

sj <- aux %>% 
  filter(code %in% c("5501", "5502","5504", "5504A", "5505")) %>% 
  group_by(hyear) %>% 
     dplyr::summarise(
     csum_mean = mean(csum, na.rm = TRUE),
     csum_sd = sd(csum, na.rm = TRUE), 
     csum_se = csum_sd/sqrt(length(csum_sd)), 
     n = n()) %>% 
    mutate(site = "SJ")

s <- bind_rows(sj, ca)

# Get mean and sd 
ca_mean <- mean(ca$csum_mean)
ca_sd <- sd(ca$csum_mean)
sj_mean <- mean(sj$csum_mean)
sj_sd <- sd(sj$csum_mean)

prec_sites_plot <- s %>% filter(hyear > 1990) %>% 
  ggplot(aes(x=hyear, y=csum_mean, fill=site, colour = site)) + 
  geom_line(position = pos) +
  scale_x_continuous(breaks=seq(1990,2020, by=2)) +
  scale_y_continuous(breaks=seq(0,1500, by=250)) +
  geom_errorbar(aes(ymin = csum_mean - csum_se, ymax= csum_mean + csum_se), width=.1, position = pos) + geom_point(size=2, position = pos) + theme_bw() +
  ylab('Cumulative Precipitation (mm)') + xlab('Year') +
  scale_fill_manual(values = colores_den[3:2]) +
  scale_colour_manual(values = colores_den[3:2]) + theme(panel.grid = element_blank())

```

Temporal evolution of cumulative precipitation (hydrological year) during the period 1990-2017. Points represent the mean, and error bars the standard error. Data from meteorological stations distributed around northern (SJ, n=4) and southern (CA, n=4) sites respectively.