### 2020-06-07 man/figures/schematic_plots.Rmd

---
output:
  word_document
bibliography: ../../refs/references.bib
csl: ../../refs/ecology.csl
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```

```{r load_pkg}
library("tidyverse")
library("here")
library("ggrepel")
library("scales")
library("grid")
library("gridExtra")
library("patchwork")
library("stringr")
library("pander")
library("lemon")
library("magrittr")
source(here::here('/script/R/custom_functions.R'))
```

```{r}
# elementos comunes

# co <- "#678cab"
co <- "#6e7561"
# co <- colores_den[3]

# co_evi <- "#6e7561"
co_evi <- "#678cab"

limitesX <- c(1925, 2020)

eje_x <- scale_x_continuous(breaks = seq(1925, 2020, by=10), limits=c(1925,2020))
```

```{r}
# BAI 
cro_sites <- read.csv(here("data/cronos_medias", "cronos_sites.csv"), header = TRUE, sep = ',')

cro_sites <- cro_sites %>% mutate(site = 
  case_when(
    site == "CA_Low" ~ "CA-Low", 
    site == "CA_High" ~ "CA-High",
    site == "SJ" ~ "SJ",
  )
)

pbai <- cro_sites %>% 
  filter(nseries > 5) %>% 
  filter(site == "SJ") %>% 
  ggplot(aes(x=year, y=bai_mean/100, colour=site)) + 
  theme_bw() + ylab('BAI') + xlab("") + 
  geom_ribbon(aes(ymin = (bai_mean - bai_se)/100,
                  ymax = (bai_mean + bai_se)/100,
                  fill=site), alpha=.4, colour=NA) +
  geom_line() + 
  geom_point(size=.8) + 
  scale_colour_manual(values = co) +
  scale_fill_manual(values = co) +
  ylab(expression(paste("BAI (", cm^2, year^-1, ")", sep=''))) +
  theme(panel.grid = element_blank(),
        legend.position = "none", #c(0.2,.8),
        strip.background = element_blank(),
        axis.ticks.length.x.top = unit(-3.5, "pt"),
        axis.title.x.top = element_blank(), 
        axis.text.x.top = element_blank(), 
        panel.border = element_blank(),
        axis.line.y = element_line(), 
        axis.ticks.x = element_blank(), 
        axis.text.x = element_blank()) + 
  eje_x

```

```{r}
# GC
gc_todos <- read.csv(here::here("data/disturbance", "disturbance_gc.csv"), header = TRUE, sep = ',')

# customize plot (corrected scale)
pos_neg_log <- scales::trans_new('signed_log', 
                                 transform = function(x) sign(x)*log(abs(x)),
                                 inverse = function(x) sign(x)*exp(abs(x)))

plotGC <- function(df, y, yse, x, my_ylab, 
                   micolor, limitesX, ...){ 
  # Get error bar aesthetic
  yerrorbar <- aes_string(ymin = paste(y, yse, sep = '-'),
                         ymax = paste(y, yse, sep = '+'))

  p <- ggplot(df, aes_string(x = x, y = y)) +
    geom_errorbar(mapping = yerrorbar, colour = micolor, width = .35, size = .2, 
                  na.rm = TRUE) + 
    geom_bar(stat = 'identity', fill = micolor, colour = micolor) +
    geom_hline(yintercept = 0, colour = 'black', lwd = 0.2) + 
    ylab(my_ylab) +
    # Add line to %GC = 50 
    geom_hline(yintercept = c(50, -50), linetype=2, colour='black', lwd=0.2) +
    theme_bw() +
    theme(panel.grid = element_blank(),
          axis.text.y = element_text(size = 5.5),
          axis.text.x = element_text(size = 6)) +
    coord_cartesian(xlim=limitesX)
    # Customize limits and breaks to min of year data 
  p
}

```

```{r plot_SJ}
gcsj <- gc_todos %>% dplyr::filter(site == 'gc_sj') %>% 
  plotGC(y = 'gc_md', yse = 'gc_md_se', x = 'year', 
         my_ylab = '% GC', micolor = co, limitesX = limitesX) + 
  xlab('') + xlab('Time') + 
  eje_x + 
  # scale_x_continuous(breaks = xrange, position = 'bottom') + 
  scale_y_continuous(trans = pos_neg_log, 
                     breaks=c(-50, 1, 50)) + 
  geom_hline(yintercept = 1,  lwd=0.2)  +
  theme(
    panel.border = element_blank(),
        axis.line.y = element_line(), 
        axis.ticks.x = element_blank(),
    axis.line.x = element_line(colour = "grey")
  )
```

```{r}
cro_rwi <- read.csv(here("data/cronos_medias", "cronos_sites_rwi.csv"), header = TRUE, sep = ',')
year_cutoffs <- read.csv(here("data/cronos_medias", "cronos_sites_rwi_year_cutoffs.csv"), header = TRUE, sep = ',')
year_sampledepth <- read.csv(here("data/cronos_medias", "cronos_sites_rwi_year_sampledepth.csv"), header = TRUE, sep = ',')

# Reorder 
cro_rwi <- cro_rwi %>% mutate(site_sorted = 
                                recode_factor(site,
                                              "caH" = "0_caH", 
                                              "caL" = "1_caL", "sj" = "2_sj"))

year_cutoffs <- year_cutoffs %>% mutate(site_sorted = 
                                recode_factor(site,
                                              "caH" = "0_caH", 
                                              "caL" = "1_caL", "sj" = "2_sj"))

year_sampledepth <- year_sampledepth %>% mutate(site_sorted = 
                                   recode_factor(site,
                                              "caH" = "0_caH", 
                                              "caL" = "1_caL", "sj" = "2_sj"))

# peores sequias 
dy <- read.csv(file=here::here('data/sequias', 'severe_spei12.csv'), header=TRUE)
dy <- dy %>% arrange(desc(d_duration))
# solamente aquellos mayor de 2 meses 
dys <- dy %>% filter(d_duration > 2) %>% as.data.frame() 
dys <- dys[,'maxyear']

rwi_sj <- cro_rwi %>% filter(site =="sj") %>% 
  ggplot(aes(x=year, y=res, colour=site)) +
  theme_bw() +
  theme(panel.grid = element_blank(), 
        axis.text = element_text(size=8), 
        legend.position = "none", 
        panel.border = element_blank(),
        axis.line.y = element_line(), 
        axis.ticks.x = element_blank(), 
        axis.text.x = element_blank()) + 
  geom_hline(yintercept = 1, colour = "grey") + 
  geom_line(size=.8) + scale_y_continuous(breaks = seq(0.4,1.8, 0.2)) + 
  ylab("RWI") + xlab("") + 
  geom_vline(xintercept = c(1995, 2005, 2012), linetype = "dotted") + 
  eje_x +
  scale_colour_manual(values = co)

```

```{r}
# Datos de iv 
iv <- read.csv(here::here("data/evi/iv_composite.csv"), header = TRUE, sep = ',')

ivv_mean <- iv %>% 
   mutate(
    clu_pop = as.factor(case_when(
      pop %in% c(1,2,3,4,5) ~ 'N',
      pop %in% c(6,7,8) ~ 'S',
      pop == 9 ~ 'out'))) %>% 
  filter(clu_pop == 'N') %>% 
  mutate(date = as.Date(date, format = "%Y-%m-%d")) %>% 
  group_by(date, clu_pop) %>% 
  summarise(iv_avg = mean(evi, na.rm = TRUE), 
            iv_sd = sd(evi, na.rm = TRUE), 
            iv_se = iv_sd/sqrt(length(evi)))

# iv_plot <- ivv_mean %>% ggplot(aes(x=date, y=iv_avg)) +
#   scale_x_date(limits= as.Date(c("1920-01-01","2020-01-01"))) + 
#   # geom_rect(aes(xmin=as.Date("2012-01-01"), 
#   #               xmax = as.Date("2012-12-31"), 
#   #               ymin=-Inf, ymax=Inf),
#   #           fill="grey", alpha=.9) + 
#   # geom_rect(aes(xmin=as.Date("2005-01-01"), 
#   #               xmax = as.Date("2005-12-31"), 
#   #               ymin=-Inf, ymax=Inf),
#   #           fill="grey", alpha=.9) + 
#   geom_line(colour=co_evi) +
#   theme_bw() + ylab("EVI") + xlab("") + 
#   theme(panel.grid = element_blank(), 
#         panel.border = element_blank(),
#         axis.line.y = element_line(), 
#         axis.ticks.x = element_blank(), 
#         axis.text.x = element_blank())

ivv_mean_year <- iv %>% 
   mutate(
    clu_pop = as.factor(case_when(
      pop %in% c(1,2,3,4,5) ~ 'N',
      pop %in% c(6,7,8) ~ 'S',
      pop == 9 ~ 'out'))) %>% 
  filter(clu_pop == 'S') %>% 
  mutate(date = as.Date(date, format = "%Y-%m-%d"), 
         year = lubridate::year(date)) %>% 
  group_by(year, clu_pop) %>% 
  summarise(iv_avg = mean(evi, na.rm = TRUE), 
            iv_sd = sd(evi, na.rm = TRUE), 
            iv_se = iv_sd/sqrt(length(evi))) %>% 
  mutate(yr = as.Date(as.character(year), format = "%Y"))

iv_plot <- ivv_mean_year %>% 
  ggplot(aes(x=yr, y=iv_avg)) + 
  geom_line(data=ivv_mean, aes(x=date, y=iv_avg), colour = "gray", alpha=.4) +
  geom_errorbar(aes(ymin = iv_avg - 5*iv_se, ymax = iv_avg + 5*iv_se), width=.3, colour = co_evi) +
  geom_point(size=1, shape=21, fill=co_evi, colour = co_evi) + 
  theme_bw() + ylab("EVI") + xlab("") + 
  theme(panel.grid = element_blank(), 
        panel.border = element_blank(),
        axis.line.y = element_line(), 
        axis.ticks.x = element_blank(), 
        axis.text.x = element_blank()) + 
  scale_x_date(limits = c(as.Date("1920-01-01"), as.Date("2020-12-31")))

```

```{r}
# From explore_anomalies.Rmd 
# Read data 
anomalias_evimean <- read.csv(here("data/anomalies", "anomalias_evimean.csv"),header = TRUE, sep = ',')

anomalias_evimean <- anomalias_evimean %>% 
  mutate(
    clu_pop = as.factor(case_when(
      pop == 1 ~ "Camarate",
      pop %in% c(2,3,4,5) ~ 'Northern slope',
      pop %in% c(6,7,8) ~ 'Southern slope',
      pop == 9 ~ 'out')),
    clu_pop2 = as.factor(case_when(
      pop %in% c(1,2,3,4,5) ~ 'Northern slope',
      pop %in% c(6,7,8) ~ 'Southern slope',
      pop == 9 ~ 'out'))) %>% 
  filter(clu_pop != 'out')

# Standardized Anomalies 
avg_sa_clu <- anomalias_evimean %>% 
  group_by(clu_pop2, y) %>% 
  dplyr::summarise(mean=mean(sa, na.rm=T),
            sd = sd(sa, na.rm=T),
            se = sd/sqrt(length(sa))) %>% 
 mutate(signo = ifelse(mean >= 0, 'pos', 'neg'),
        yr = as.Date(as.character(y), format = "%Y"))

sa <- avg_sa_clu %>%
  filter(clu_pop2 =="Northern slope") %>% 
  ggplot(aes(x=yr, y=mean)) + 
  geom_hline(yintercept = 0, size=0.3, colour = "gray") + 
  geom_bar(stat = "identity", fill=co_evi, colour="white", size=.1) + 
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width=.5, colour=co_evi) +
  # facet_wrap(~clu_pop2, nrow = 2) + 
  # scale_fill_manual(values = c("pos" = color_pos, "neg" =  color_neg)) +
  # scale_color_manual(values = c("pos" = color_pos, "neg" =  color_neg)) +
  scale_x_date(limits = c(as.Date("1920-01-01"), as.Date("2020-12-31"))) +  
  ylab('EVI std. Anomaly') + xlab("") +
  theme_bw() +
  theme(panel.grid = element_blank(), 
        panel.border = element_blank(),
        axis.line.y = element_line(), 
        axis.ticks.x = element_blank(), 
        axis.text.x = element_blank())
```

# SPEI

```{r}

# DATOS desde 1901

# Get file names
myfiles <- list.files(path = here("data_raw/spei/spei_longterm"), pattern = "\\.csv$")

mydf <- data.frame()

for (j in myfiles){ 
  aux <- read.csv(file=here("data_raw/spei/spei_longterm", j), header = TRUE, sep = ',')

  # Remove csv and get name 
  name_aux <- str_replace(j, ".csv", "")

  # Get lat long
  latlong <- str_replace(name_aux, "spei_", "")
  mylong <- as.numeric(str_split_fixed(latlong, "_", 2))[1]
  mylat <- as.numeric(str_split_fixed(latlong, "_", 2))[2]

  # Store lat and long 
  aux$lat <- mylat
  aux$long <- mylong

  # year and month 
  aux$month <- rep(1:12, 115)

  auxy <- c()
  years <- c() 
  for (j in 1901:2015) { 
    auxy <- as.data.frame(rep(j,12))
    years <- rbind(years, auxy)
    } 
  names(years) <- 'year'

  aux <- cbind(aux,years)

  # join all df 
  mydf <- rbind(mydf, aux)
}

# Convert to date  
mydf <- mydf %>% mutate(date = sprintf("%d-%02d", year, month),
                        date = as.Date(paste0(date, '-01'), format="%Y-%m-%d"))

speidf <- mydf %>% dplyr::select(SPEI06,SPEI12,SPEI24,month, year, date) %>% 
  group_by(year,month, date) %>%
  summarise_at(vars(SPEI06,SPEI12,SPEI24), funs(mean,sd,se=sd(.)/sqrt(n()))) %>% 
  as.data.frame() 
```

```{r}

vi <- "SPEI12"

df <- speidf %>% dplyr::select(date, starts_with("SPEI12")) %>% 
  select_all(~gsub("SPEI12_", "", .)) %>% 
  mutate(signo = ifelse(mean > 0, "pos", "neg"))

p_spei <- df %>% 
  ggplot(aes(x = date, y = mean)) + 
    geom_bar(stat = "identity", colour="grey", fill="grey") + 
  ylab("SPEI") + xlab("") + 
  theme_bw() +
  theme(panel.grid = element_blank(), 
        panel.border = element_blank(),
        axis.line = element_line()) +
  scale_x_date(limits = c(as.Date("1901-01-01"), as.Date("2016-12-31")), date_labels = "%Y", date_breaks ="5 years")
```

```{r}

a <- rwi_sj + theme(plot.margin = unit(c(0,0,-5,0), "pt")) 
b <- pbai + theme(plot.margin = unit(c(0,0,-5,0), "pt"))
c <- gcsj +  theme(plot.margin = unit(c(0,0,-4,0), "pt"))

m <- p_spei + iv_plot + sa + a+b+c+plot_layout(ncol = 1)
```

```{r}
saveImage(m, path="./man/figures/svg/", dpi=300, w=25, h=15, fig= "plot_metrics", u = "cm") 
```